### 2025-02-01 products/manuscript/Manuscript.qmd
Part 1

---
title: "Learning Influenza Forecasting"
author: "Annalise Cramer"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

I have two potential ideas for this project. The first one would be more challenging for me, but the second I have more skills for currently, please give me feedback on what you think is reasonable.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(tidyverse)
library(survey)
```

**Authors**

* Annalise Cramer$^{1,2 \land}$ (ORCID: 0009-0002-8718-3593)

**Author affiliations**

1. College of Public Health, University of Georgia, Athens, GA, USA.
2. Center for Ecology of Infectious Diseases, Athens, GA, USA.

$\land$ Corresponding author: atc10260@uga.edu

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect those of the University of Georgia.

{{< pagebreak >}}

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._

## Questions/Hypotheses to be addressed
Option 1: Can I make a flu forecast for this season? When I compare my work to the actual data, will it hold up?

Option 2:

What is the prevalence of poor mental health? Is poor mental health associated with a positive covid test?

Factors studied will include hours of sleep, reported depressed mood, reported trouble sleeping, recent exercise, and reported social/emotional support.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
Option 1:

Data for this project would be using positive infleunza case counts, which are listed weekly from CDC's [FluView](https://www.cdc.gov/fluview/index.html). This data is updated weekly, and extends years backwards. Other useful measures are available, such as hospitalizations. I would need to obtain number from each prior week going backawards, I'm not sure if I'd have to do this manually or if there's a consolidated form somewhere. Several people in my lab do forecasting, and I'm hoping to learn enough that I can be caught up to speed on their reserach projects. I plan to use and ARIMA model and a weighted interval score to evaluate my work. I'm currently taking Time Series in the statistics department, so I'm hoping as the semester progresses I will learn more about this modeling technique.

Option 2:

Data for this potential project would be using survey data from the National Health Interview Survey, produced by IPUMS through the University of Minnesota. This data was collected among United States residents of all ages and genders from 2019 - 2022 in the United States. There's 151406 observations and 47 vairables. I would select variables I need, including creating new ones, and using the survey design variables. I would create models (basic and adjusted) to study association (generating prevelance ratios with CIs) between mental health factors (hours of sleep, feeling depressed mood, amount of social and emotional support, trouble sleeping, recent exercise) with a postiive covid test result.

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

```{r}
data2_original <- readRDS("~/MADA/CRAMER-MADA-Project/data/NHIS_COVID.rds")
data2 <- labelled::remove_val_labels(data_original)
summary(data2)

data_using <- data2 %>%
  select(STRATA, PSU, PERNUM, SAMPWEIGHT, MOD10DMIN, VIG10DTP, PA18AER, PA18AERSTR, HRSLEEP, DEPFREQ, PHQSLEEP, CMPSUPPORT, CVDTESTRSLT)
```

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

### 2025-02-06 products/manuscript/Manuscript.qmd
part 1 fix
Fix loading issue commit

@@
 library(survey)
+here()
 ```
@@
 ## Questions/Hypotheses to be addressed
-Option 1: Can I make a flu forecast for this season? When I compare my work to the actual data, will it hold up?
+Option 1:
+
+Can I make a flu forecast for this season? When I compare my work to the actual data, will it hold up?
 
@@
 I would create models (basic and adjusted) to study association (generating prevelance ratios with CIs) between mental health factors (hours of sleep, feeling depressed mood, amount of social and emotional support, trouble sleeping, recent exercise) with a postiive covid test result.
+I will generate tables and plots for all results.
 
@@
 ```{r}
-data2_original <- readRDS("~/MADA/CRAMER-MADA-Project/data/NHIS_COVID.rds")
+data_original <- readRDS(here("data", "NHIS_COVID.rds"))
 data2 <- labelled::remove_val_labels(data_original)

### 2025-02-21 code/processing-code/data-cleaning-project.qmd
Part 2

---
title: "R Coding Exercise"
---

To start, I'll load packages. As I need packages along the way, I'll return here and add them.
```{r}
library(tidyverse)
library(gt)
library(webshot2)
library(here)
```

Now, loading data using the here package.
```{r}
here() #set working directory
data_original <- readRDS(here("data","raw-data", "NHIS_COVID.rds"))
names(data_original)
```

Selecting variables we need will make the dataset smaller and quicker to work with. Note it is renamed at each step for reproducibility. 
```{r}
data1 <- data_original %>%
  select(STRATA, PSU, PERNUM, SAMPWEIGHT, AGE, PA18AERSTR, HRSLEEP, DEPFREQ, PHQSLEEP, CMPSUPPORT, CVDTESTRSLT)
```

In the following code chunk, exploration and definition of each of these variables will be completed. These definitions and unique values help me to know how to define my binary vairables in the next step. Several of these vairable have codes for unknown, don't know, etc. I will need to convert these to NA.
```{r}
unique(data1$AGE)
unique(data1$PA18AERSTR)
unique(data1$HRSLEEP)
unique(data1$DEPFREQ)
unique(data1$PHQSLEEP)
unique(data1$CMPSUPPORT)
unique(data1$CVDTESTRSLT)
```

We will create binary vairables in a new dataset. For each variable, those who answered NA, 0, 996, 997, 998, 999 will be converted to NA for the remainder of that variable for the continuation of the analysis.
```{r}
unique(data1$CVDTESTRSLT)
data2 <- data1 %>%
  mutate(
    PA18AERSTR = ifelse(PA18AERSTR %in% c(0, 7, 8, 9), NA, PA18AERSTR), #put NA, 0, 6, 7, 8, 9 to NA
    EXERCISE = if_else(PA18AERSTR %in% c(2,3,4), 1, 0, missing = NA_real_) #create EXERCISE vairable, 1 if guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met
    ) %>%
    mutate(
    AGE = ifelse(AGE %in% c(997, 999), NA, AGE),  #invalid ages to NA
    SLEEP = case_when( #Based on age, amount of sleep required from https://www.nhlbi.nih.gov/health/sleep/how-much-sleep#:~:text=Experts%20recommend%20that%20adults%20sleep,or%20more%20hours%20a%20night.
      AGE < 1  ~ if_else(HRSLEEP >= 12 & HRSLEEP <= 16, 1, 0, missing = NA_real_),
      AGE >= 1 & AGE < 3       ~ if_else(HRSLEEP >= 11 & HRSLEEP <= 14, 1, 0, missing = NA_real_),
      AGE >= 3 & AGE < 6       ~ if_else(HRSLEEP >= 10 & HRSLEEP <= 13, 1, 0, missing = NA_real_),
      AGE >= 6 & AGE < 13      ~ if_else(HRSLEEP >= 9 & HRSLEEP <= 12, 1, 0, missing = NA_real_),
      AGE >= 13 & AGE <= 18    ~ if_else(HRSLEEP >= 8 & HRSLEEP <= 10, 1, 0, missing = NA_real_),
      TRUE                     ~ NA_real_  #default NA for unspecified ages
    )) %>%
  mutate(
    DEPRESSED = case_when(
      DEPFREQ %in% c(1, 2) ~ 1, #daily and weekly is considered depressed
      DEPFREQ %in% c(3, 4, 5)    ~ 0, #monthly and a few times a year and never are considered not depressed
      DEPFREQ %in% c(0, 7, 8, 9) ~ NA_real_,#NIU & Unknowns are NA
      TRUE ~ NA_real_ #default NA for unexpected values
    )) %>%
  mutate(
    TROUBLE_SLEEPING = case_when(
      PHQSLEEP %in% c(1, 2, 3) ~ 1, #several days, more than half days, nearly every day are trouble sleeping
      PHQSLEEP == 0            ~ 0,  #not at all is no trouble sleeping
      PHQSLEEP %in% c(6, 7, 8, 9) ~ NA_real_,  #NIU & Unknowns are NA
      TRUE ~ NA_real_  #default NA for unexpected values
    )) %>%
  mutate(
    SOCIAL = case_when(
      CMPSUPPORT == 1 ~ 0,   #more social and emotional support
      CMPSUPPORT == 2 ~ 1,  #less social and emotional support
      CMPSUPPORT == 3 ~ 0,   #about the same
      CMPSUPPORT %in% c(0, 7, 8, 9) ~ NA_real_,  # NIU & Unknowns → NA
      TRUE ~ NA_real_  #default NA for unexpected values
    )) %>%
  mutate(
    COVID = case_when(
      CVDTESTRSLT == 2 ~ 1,  #yes, had COVID
      CVDTESTRSLT == 1 ~ 0,  #no, did not have COVID
      CVDTESTRSLT %in% c(0, 3, 7, 8, 9) ~ NA_real_,  #NIU, did not receive results, unknowns, and NA are NA
      TRUE ~ NA_real_  #default NA for unexpected values
    ))

```

Now, I will make plots to explore how these look to ensure there's no problems. I used ChatGPT to help make the table. Each row doesn't add up to 1, this is because the yes/no adds up to 1, and the NA and not NA adds up to 1. Notably, some of these vairables have high NA porportions, but I can't remove them from the set because that would affect the survey design variables, but I will keep it in mind for the rest of the analysis should there be any issues such as singleton strata varibales etc.
```{r}
hist(data2$SLEEP) #histograms of 0 and 1 for each variable
hist(data2$EXERCISE)
hist(data2$DEPRESSED)
hist(data2$TROUBLE_SLEEPING)
hist(data2$SOCIAL)
hist(data2$COVID)

#summarize EXERCISE
summary_exercise <- data2 %>%
  summarise(
    Variable = "EXERCISE",
    Prop_1 = mean(EXERCISE == 1, na.rm = TRUE),
    Prop_0 = mean(EXERCISE == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(EXERCISE))
  )

#summarize SLEEP
summary_sleep <- data2 %>%
  summarise(
    Variable = "SLEEP",
    Prop_1 = mean(SLEEP == 1, na.rm = TRUE),
    Prop_0 = mean(SLEEP == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(SLEEP))
  )

#summarize DEPRESSED
summary_depressed <- data2 %>%
  summarise(
    Variable = "DEPRESSED",
    Prop_1 = mean(DEPRESSED == 1, na.rm = TRUE),
    Prop_0 = mean(DEPRESSED == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(DEPRESSED))
  )

#summarize TROUBLE SLEEPING
summary_trouble_sleeping <- data2 %>%
  summarise(
    Variable = "TROUBLE SLEEPING",
    Prop_1 = mean(TROUBLE_SLEEPING == 1, na.rm = TRUE),
    Prop_0 = mean(TROUBLE_SLEEPING == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(TROUBLE_SLEEPING))
  )

#summarize SOCIAL
summary_social <- data2 %>%
  summarise(
    Variable = "SOCIAL",
    Prop_1 = mean(SOCIAL == 1, na.rm = TRUE),
    Prop_0 = mean(SOCIAL == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(SOCIAL))
  )

#summarize COVID
summary_covid <- data2 %>%
  summarise(
    Variable = "COVID",
    Prop_1 = mean(COVID == 1, na.rm = TRUE),
    Prop_0 = mean(COVID == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(COVID))
  )

#combine all summaries and convert to gt table
summary_table <- bind_rows(
  summary_exercise, 
  summary_sleep, 
  summary_depressed, 
  summary_trouble_sleeping, 
  summary_social, 
  summary_covid
) %>%
  gt() %>%
  tab_header(
    title = "Proportions of 1, 0, and NA for Various Variables"
  )

#print the combined table
summary_table

#save the table as a PNG image
gtsave(summary_table, here("data", "processed-data", "figures", "summary_table.png"))
```

For our final step, we will save this data to the processed_data folder.
```{r}
#save data2 as data3 in the processed-data folder
here()
write_rds(data2, here("data","processed-data", "data3.rds"))
```

### 2025-02-21 products/manuscript/Manuscript.qmd
Part 2

@@
 ---
-
-I have two potential ideas for this project. The first one would be more challenging for me, but the second I have more skills for currently, please give me feedback on what you think is reasonable.
 
@@
 ## Questions/Hypotheses to be addressed
-Option 1:
 
-Can I make a flu forecast for this season? When I compare my work to the actual data, will it hold up?
-
-Option 2:
-
-What is the prevalence of poor mental health? Is poor mental health associated with a positive covid test?
+1.) What is the prevalence of poor mental health?
+2.) Is poor mental health associated with a positive covid test?
 
@@
 _Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
-Option 1:
 
-Data for this project would be using positive infleunza case counts, which are listed weekly from CDC's [FluView](https://www.cdc.gov/fluview/index.html).
-This data is updated weekly, and extends years backwards.
-Other useful measures are available, such as hospitalizations.
-I would need to obtain number from each prior week going backawards, I'm not sure if I'd have to do this manually or if there's a consolidated form somewhere.
-Several people in my lab do forecasting, and I'm hoping to learn enough that I can be caught up to speed on their reserach projects.
-I plan to use and ARIMA model and a weighted interval score to evaluate my work.
-I'm currently taking Time Series in the statistics department, so I'm hoping as the semester progresses I will learn more about this modeling technique.
-
-Option 2:
-
-Data for this potential project would be using survey data from the National Health Interview Survey, produced by IPUMS through the University of Minnesota.
+Data for this project is survey data collected from the National Health Interview Survey, produced by IPUMS through the University of Minnesota.
 This data was collected among United States residents of all ages and genders from 2019 - 2022 in the United States.
 There's 151406 observations and 47 vairables.
-I would select variables I need, including creating new ones, and using the survey design variables.
-I would create models (basic and adjusted) to study association (generating prevelance ratios with CIs) between mental health factors (hours of sleep, feeling depressed mood, amount of social and emotional support, trouble sleeping, recent exercise) with a postiive covid test result.
+
+Survey design variables were used and new binary variables were created from other variables.
+
+I will create models (basic and adjusted) to study association (generating prevelance ratios with CIs) between mental health factors (hours of sleep, feeling depressed mood, amount of social and emotional support, trouble sleeping, recent exercise benchmarks) with a postiive covid test result.
 I will generate tables and plots for all results.
@@
 ## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
+This data was obtained from IPUMS, used with permission.
 
@@
 
+Raw data called NHIS_COVID.rds, sits with the data folder, in a a subfolder called raw data.
+The processed data, called data3.rds, is within the data folder in a subfolder called processed-data.
+The code to clean and explore and process this dadta is within the code folder, in the processing-code subfolder, called data-cleaning-project.qmd.
+
+The raw data was loaded, and necessary variables and survery design variables were selected. After exploring each variable, new binary variables were created.
+
+For now, let us load the processed data.
+Here's a histogram of the processed outcome vairable, which represents whether or not an indivudal has ever tested postive for COVID.
+Histograms of all vairables are available in the data-cleaning-project.qmd file, but will not be shown here for the sake of space.
 ```{r}
-data_original <- readRDS(here("data", "NHIS_COVID.rds"))
-data2 <- labelled::remove_val_labels(data_original)
-summary(data2)
-
-data_using <- data2 %>%
-  select(STRATA, PSU, PERNUM, SAMPWEIGHT, MOD10DMIN, VIG10DTP, PA18AER, PA18AERSTR, HRSLEEP, DEPFREQ, PHQSLEEP, CMPSUPPORT, CVDTESTRSLT)
+library(here)
+data <- read_rds(here("data", "processed-data", "data3.rds"))
+hist(data$COVID)
 ```
@@
 #| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
+#| tbl-cap: "Data summary table, proportions of 0 compared to 1, and proportion NA to not NA."
 #| echo: FALSE
-resulttable=readRDS("../../results/tables/summarytable.rds")
-knitr::kable(resulttable)
+here()
+knitr::include_graphics(here("data","processed-data", "figures", "summary_table.png"))
 ```

### 2025-03-15 code/analysis-code/analysis.qmd
Part 3 commit, complete eda and start analysis

---
title: "Analysis"
---

To start, I'll load packages. As I need packages along the way, I'll return here and add them.
```{r}
library(tidyverse)
library(survey)
library(tableone)
library(gt)
library(here)
```

Now, loading data using the here package.
```{r}
here() #set working directory
data <- readRDS(here("data","processed-data", "data3.rds"))
data <- labelled::remove_val_labels(data) #removes labels from haven package which cause problems
```

As a reminder, we are going to use the following variables, and stratify by covid test result.
EXERCISE: Whether or not individual met guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met, 1 if any were met
SLEEP: If individual reported getting the proper amount of sleep for their resspective age group
DEPRESSED: If individual felt depressed weekly+ or not
TROUBLE_SLEEPING: If individual reported having trouble sleeping over past several days, more than half days, or nearly every day or not at all
SOCIAL: If person felt like they were reciveing less social support or not
COVID: If they've tested positive for covid or not

We create estimates and CIs using complex survey design variables
```{r}
#list variables
varstable1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
             "SOCIAL", "COVID")

#list categories
cat_vartab1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
             "SOCIAL", "COVID")

survey = svydesign(
  ids = ~PSU, 
  strata = ~STRATA, 
  weights = ~SAMPWEIGHT, 
  nest = TRUE, 
  data = data)

exercise <- svyby(~as.factor(EXERCISE), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
sleep <- svyby(~as.factor(SLEEP), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
depressed <- svyby(~as.factor(DEPRESSED), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
trouble_sleeping <- svyby(~as.factor(TROUBLE_SLEEPING), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
social <- svyby(~as.factor(SOCIAL), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
str(exercise)
```

Let's rearrange these exercise numbers into a table. I used ChatGPT and Claude to help with making the table. 
```{r}
# Convert the svyby object to a data frame (if not already)
exercise_df <- as.data.frame(exercise)

# Clean up column names to make them more readable
colnames(exercise_df) <- c("COVID", "EXERCISE_0", "EXERCISE_1", "CI_L_EXERCISE_0", "CI_L_EXERCISE_1", "CI_U_EXERCISE_0", "CI_U_EXERCISE_1")

# Use a different approach to reshape the data
exercise_long <- exercise_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and exercise level
  mutate(
    Type = case_when(
      grepl("^EXERCISE", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    Exercise = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, Exercise),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
exercise_with_ci <- exercise_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, Exercise, Mean, CI)

# Now create the final table format
exercise_final <- exercise_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = Exercise,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
exercise_table <- exercise_final %>%
  gt() %>%
  tab_header(
    title = "Exercise Data by COVID Status"
  ) %>%
  cols_label(
    Exercise = "Exercise Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = Exercise == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Exercise Level: 0 = No regular exercise, 1 = Meets requirements for regular exercise according to American Heart Association",
    locations = cells_column_labels(columns = Exercise)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = Exercise),
    fn = function(x) {
      ifelse(x == "0", "No Regular Exercise", "Regular Exercise")
    }
  )

# Print the table
exercise_table
```

Next, we will add the sleep variable. I used ChatGPT for this.
```{r}
# Convert the svyby object to a data frame (if not already)
sleep_df <- as.data.frame(sleep)  # Assuming 'sleep' is the dataset

# Clean up column names to make them more readable
colnames(sleep_df) <- c("COVID", "SLEEP_0", "SLEEP_1", "CI_L_SLEEP_0", "CI_L_SLEEP_1", "CI_U_SLEEP_0", "CI_U_SLEEP_1")

# Use a different approach to reshape the data
sleep_long <- sleep_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and sleep level
  mutate(
    Type = case_when(
      grepl("^SLEEP", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    Sleep = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, Sleep),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
sleep_with_ci <- sleep_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, Sleep, Mean, CI)

# Now create the final table format
sleep_final <- sleep_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = Sleep,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
sleep_table <- sleep_final %>%
  gt() %>%
  tab_header(
    title = "Sleep Data by COVID Status"
  ) %>%
  cols_label(
    Sleep = "Sleep Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = Sleep == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Sleep Level: 0 = Less than recommended sleep, 1 = Met recommendations for sleep by age groups according to NIH",
    locations = cells_column_labels(columns = Sleep)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = Sleep),
    fn = function(x) {
      ifelse(x == "0", "Not enough sleep", "Enough sleep")
    }
  )

# Print the table
sleep_table

#################################################################################

# Extract the data from the gt tables
exercise_table_df <- as.data.frame(exercise_table)
sleep_table_df <- as.data.frame(sleep_table)

# Ensure that the column names are the same across both tables
colnames(sleep_table_df) <- colnames(exercise_table_df)

# Combine the tables row by row
combined_df <- bind_rows(exercise_table_df, sleep_table_df)

# Convert the combined data frame back to a gt table
combined_table <- combined_df %>%
  gt() %>%
  tab_header(
    title = "Combined Exercise and Sleep Data by COVID Status"
  )

# Print the combined gt table
combined_table
```

Create the same tables with depressed, trouble_sleeping, and social. This was completed in BBEdit.
```{r}
#########################################################################################
# depressed variable

# Convert the svyby object to a data frame (if not already)
depressed_df <- as.data.frame(depressed)

# Clean up column names to make them more readable
colnames(depressed_df) <- c("COVID", "depressed_0", "depressed_1", "CI_L_depressed_0", "CI_L_depressed_1", "CI_U_depressed_0", "CI_U_depressed_1")

# Use a different approach to reshape the data
depressed_long <- depressed_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and depressed level
  mutate(
    Type = case_when(
      grepl("^depressed", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    depressed = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, depressed),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
depressed_with_ci <- depressed_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, depressed, Mean, CI)

# Now create the final table format
depressed_final <- depressed_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = depressed,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
depressed_table <- depressed_final %>%
  gt() %>%
  tab_header(
    title = "depressed Data by COVID Status"
  ) %>%
  cols_label(
    depressed = "depressed Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = depressed == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Depressed level: 0 = Feeling depressed less than weekly, 1 = Feeling depressed weekly or more",
    locations = cells_column_labels(columns = depressed)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = depressed),
    fn = function(x) {
      ifelse(x == "0", "Not regularly depressed", "Frequent depressed feeling")
    }
  )

# Print the table
depressed_table

#########################################################################################
# trouble_sleeping variable

# Convert the svyby object to a data frame (if not already)
trouble_sleeping_df <- as.data.frame(trouble_sleeping)

# Clean up column names to make them more readable
colnames(trouble_sleeping_df) <- c("COVID", "trouble_sleeping_0", "trouble_sleeping_1", "CI_L_trouble_sleeping_0", "CI_L_trouble_sleeping_1", "CI_U_trouble_sleeping_0", "CI_U_trouble_sleeping_1")

# Use a different approach to reshape the data
trouble_sleeping_long <- trouble_sleeping_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and trouble_sleeping level
  mutate(
    Type = case_when(
      grepl("^trouble_sleeping", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    trouble_sleeping = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, trouble_sleeping),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
trouble_sleeping_with_ci <- trouble_sleeping_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, trouble_sleeping, Mean, CI)

# Now create the final table format
trouble_sleeping_final <- trouble_sleeping_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = trouble_sleeping,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
trouble_sleeping_table <- trouble_sleeping_final %>%
  gt() %>%
  tab_header(
    title = "trouble_sleeping Data by COVID Status"
  ) %>%
  cols_label(
    trouble_sleeping = "trouble_sleeping Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = trouble_sleeping == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "trouble_sleeping level: 0 = No trouble sleeping, 1 = Reported trouble sleeping",
    locations = cells_column_labels(columns = trouble_sleeping)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = trouble_sleeping),
    fn = function(x) {
      ifelse(x == "0", "No reported trouble sleeping", "Regular trouble sleeping")
    }
  )

# Print the table
trouble_sleeping_table

#########################################################################################
# social variable

# Convert the svyby object to a data frame (if not already)
social_df <- as.data.frame(social)

# Clean up column names to make them more readable
colnames(social_df) <- c("COVID", "social_0", "social_1", "CI_L_social_0", "CI_L_social_1", "CI_U_social_0", "CI_U_social_1")

# Use a different approach to reshape the data
social_long <- social_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and social level
  mutate(
    Type = case_when(
      grepl("^social", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    social = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, social),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
social_with_ci <- social_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, social, Mean, CI)

# Now create the final table format
social_final <- social_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = social,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
social_table <- social_final %>%
  gt() %>%
  tab_header(
    title = "social Data by COVID Status"
  ) %>%
  cols_label(
    social = "social Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = social == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "social Level: 0 = Regular social support, 1 = Less social support",
    locations = cells_column_labels(columns = social)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = social),
    fn = function(x) {
      ifelse(x == "0", "Regular social support", "Less social support")
    }
  )

# Print the table
social_table

```

Now, we will create the combined table of exercise, sleep, depressed, trouble_sleeping, and social. ChatGPT helped with many tweaks.
```{r}
# Extract the data from the other gt tables
depressed_table_df <- as.data.frame(depressed_table)
trouble_sleeping_table_df <- as.data.frame(trouble_sleeping_table)
social_table_df <- as.data.frame(social_table)

# Ensure the column names match across all tables
colnames(depressed_table_df) <- colnames(combined_df)
colnames(trouble_sleeping_table_df) <- colnames(combined_df)
colnames(social_table_df) <- colnames(combined_df)

# Combine all the tables row by row
combined_all_df <- bind_rows(combined_df, depressed_table_df, trouble_sleeping_table_df, social_table_df)

# Add a row_id column to the data frame
combined_all_df <- combined_all_df %>%
  mutate(row_id = row_number())

# Convert the combined data frame back to a gt table
combined_all_table <- combined_all_df %>%
  gt() %>%
  tab_header(
    title = "Mental Health Factors by SARS-CoV-2 Test Status"
  ) %>%
  # Spanners for COVID Negative and COVID Positive
  tab_spanner(
    label = "Never has tested positive for Covid",
    columns = starts_with("Estimate") | starts_with("CI 95%")
  ) %>%
  tab_spanner(
    label = "Has tested positive for Covid",
    columns = starts_with("Estimate") | starts_with("CI 95%")
  ) %>%
  # Remove the word 'Exercise' from the first row
  cols_label(
    Exercise = ""  # This removes the "Exercise" label
  ) %>%
  # Format the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Exercise requirements determined according to American Heart Association. Sleep levels determined by NIH recommendations by age group.",
    locations = cells_column_labels(columns = Exercise)
  ) %>%
  # Add alternating row colors using row_id
  tab_style(
    style = cell_fill(color = "#f5f5f5"),  # Gray color for odd rows
    locations = cells_body(rows = row_id %% 2 == 1)  # Apply to odd rows
  ) %>%
  tab_style(
    style = cell_fill(color = "white"),  # White color for even rows
    locations = cells_body(rows = row_id %% 2 == 0)  # Apply to even rows
  ) %>%
  # Add solid vertical line between row names and Mean_0 columns
  tab_style(
    style = cell_borders(
      sides = c("left"),
      color = "grey",
      weight = px(1)
    ),
    locations = cells_body(
      columns = starts_with("Mean_0")
    )
  ) %>%
  # Add solid vertical line between COVID Negative and COVID Positive columns
  tab_style(
    style = cell_borders(
      sides = c("left"),
      color = "grey",
      weight = px(1)
    ),
    locations = cells_body(
      columns = starts_with("Mean_1")
    )
  ) %>%
  # Hide the row_id column
  cols_hide(columns = "row_id")

# Print the combined gt table
combined_all_table

#footnote:
#"Exercise requirements determined according to American Heart Association. Sleep levels determined by NIH recommendations by age group."

# Save the table as a PNG
here()
gtsave(combined_all_table, here("results", "figures", "result_table1.png"))
```

### 2025-03-15 code/eda-code/eda.qmd
Part 3 commit, complete eda and start analysis

---
title: "Data-Cleaning"
---

To start, I'll load packages. As I need packages along the way, I'll return here and add them.
```{r}
library(tidyverse)
library(gt)
library(webshot2)
library(here)
```

Now, loading data using the here package.
```{r}
here() #set working directory
data <- readRDS(here("data","processed-data", "data3.rds"))
```

Now, I will make plots to explore how these look to ensure there's no problems. I used ChatGPT to help make the table. Each row doesn't add up to 1, this is because the yes/no adds up to 1, and the NA and not NA adds up to 1. Notably, some of these vairables have high NA porportions, but I can't remove them from the set because that would affect the survey design variables, but I will keep it in mind for the rest of the analysis should there be any issues such as singleton strata variables, lower than expected estimates, etc.
```{r}
hist(data$SLEEP) #histograms of 0 and 1 for each variable
hist(data$EXERCISE)
hist(data$DEPRESSED)
hist(data$TROUBLE_SLEEPING)
hist(data$SOCIAL)
hist(data$COVID)
```

```{r}
#summarize EXERCISE
summary_exercise <- data %>%
  summarise(
    Variable = "EXERCISE",
    Prop_1 = mean(EXERCISE == 1, na.rm = TRUE),
    Prop_0 = mean(EXERCISE == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(EXERCISE))
  )

#summarize SLEEP
summary_sleep <- data %>%
  summarise(
    Variable = "SLEEP",
    Prop_1 = mean(SLEEP == 1, na.rm = TRUE),
    Prop_0 = mean(SLEEP == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(SLEEP))
  )

#summarize DEPRESSED
summary_depressed <- data %>%
  summarise(
    Variable = "DEPRESSED",
    Prop_1 = mean(DEPRESSED == 1, na.rm = TRUE),
    Prop_0 = mean(DEPRESSED == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(DEPRESSED))
  )

#summarize TROUBLE SLEEPING
summary_trouble_sleeping <- data %>%
  summarise(
    Variable = "TROUBLE SLEEPING",
    Prop_1 = mean(TROUBLE_SLEEPING == 1, na.rm = TRUE),
    Prop_0 = mean(TROUBLE_SLEEPING == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(TROUBLE_SLEEPING))
  )

#summarize SOCIAL
summary_social <- data %>%
  summarise(
    Variable = "SOCIAL",
    Prop_1 = mean(SOCIAL == 1, na.rm = TRUE),
    Prop_0 = mean(SOCIAL == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(SOCIAL))
  )

#summarize COVID
summary_covid <- data %>%
  summarise(
    Variable = "COVID",
    Prop_1 = mean(COVID == 1, na.rm = TRUE),
    Prop_0 = mean(COVID == 0, na.rm = TRUE),
    Prop_NA = mean(is.na(COVID))
  )

#combine all summaries and convert to gt table
summary_table <- bind_rows(
  summary_exercise, 
  summary_sleep, 
  summary_depressed, 
  summary_trouble_sleeping, 
  summary_social, 
  summary_covid
) %>%
  gt() %>%
  tab_header(
    title = "Proportions of 1, 0, and NA for Various Variables"
  )

#print the combined table
summary_table

#save the table as a PNG image
gtsave(summary_table, here("results", "figures", "summary_table_with_nas.png"))
```
Now, let's make a table of the proportions of answers for each of our variables. 
```{r}
#list variables
varstable1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
             "SOCIAL", "COVID")

#list categories
cat_vartab1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
             "SOCIAL", "COVID")

#use function to make table
Tab1 = CreateTableOne(vars = varstable1, factorVars = cat_vartab1, data = data)

#print and summarize
print(Tab1, showAllLevels = TRUE)
summary(Tab1)

# Convert the TableOne object to a data frame
tab1_df <- as.data.frame(print(Tab1, showAllLevels = TRUE, quote = FALSE, noSpaces = TRUE))

# Clean up row names
tab1_df <- tibble::rownames_to_column(tab1_df, "Variable")
```

That's a helpful table, let's turn that into a figure. I used ClaudeAI to help with this code.
```{r}
# More advanced formatting
tab1_gt <- gt(tab1_df) %>%
  # Add a title
  tab_header(
    title = md("**COVID test status and Mental Health Factors**"),
    subtitle = "Survey Results Summary"
  ) %>%
  # Format the columns
  fmt_percent(
    columns = contains("%"),
    decimals = 1
  ) %>%
  # Group the variables
  tab_row_group(
    label = "COVID Status",
    rows = contains("COVID")
  ) %>%
  tab_row_group(
    label = "Mental Health Indicators",
    rows = contains(c("DEPRESSED", "TROUBLE_SLEEPING"))
  ) %>%
  tab_row_group(
    label = "Lifestyle Factors",
    rows = contains(c("EXERCISE", "SLEEP", "SOCIAL"))
  ) %>%
  # Style the table
  tab_style(
    style = cell_fill(color = "#e6f3ff"),
    locations = cells_row_groups()
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_row_groups()
  ) %>%
  # Add footnotes if needed
  tab_footnote(
    footnote = "Percentages may not sum to 100% due to rounding",
    locations = cells_column_labels(columns = contains("%"))
  ) %>%
  # Adjust column widths
  cols_width(
    Variable ~ px(250),
    everything() ~ px(120)
  )

tab1_gt

# Save the table as a PNG
here()
gtsave(tab1_gt, here("results", "figures", "table1.png"))
```

### 2025-03-15 code/processing-code/processing.qmd
Part 3 commit, complete eda and start analysis

--- 2025-02-21 code/processing-code/data-cleaning-project.qmd
@@
 ---
-title: "R Coding Exercise"
+title: "Processing Code"
 ---
@@
 data_original <- readRDS(here("data","raw-data", "NHIS_COVID.rds"))
-names(data_original)
+data1 <- labelled::remove_val_labels(data_original) #removes labels from haven package which cause problems
 ```
@@
 ```{r}
+names(data_original) #view names of variables
 data1 <- data_original %>%
@@
 ```{r}
-unique(data1$CVDTESTRSLT)
 data2 <- data1 %>%
@@
     PA18AERSTR = ifelse(PA18AERSTR %in% c(0, 7, 8, 9), NA, PA18AERSTR), #put NA, 0, 6, 7, 8, 9 to NA
-    EXERCISE = if_else(PA18AERSTR %in% c(2,3,4), 1, 0, missing = NA_real_) #create EXERCISE vairable, 1 if guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met
+    EXERCISE = if_else(PA18AERSTR %in% c(2,3,4), 1, 0, missing = NA_real_) #create EXERCISE variable, 1 if guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met
     ) %>%
@@
 
-Now, I will make plots to explore how these look to ensure there's no problems.
-I used ChatGPT to help make the table.
-Each row doesn't add up to 1, this is because the yes/no adds up to 1, and the NA and not NA adds up to 1.
-Notably, some of these vairables have high NA porportions, but I can't remove them from the set because that would affect the survey design variables, but I will keep it in mind for the rest of the analysis should there be any issues such as singleton strata varibales etc.
-```{r}
-hist(data2$SLEEP) #histograms of 0 and 1 for each variable
-hist(data2$EXERCISE)
-hist(data2$DEPRESSED)
-hist(data2$TROUBLE_SLEEPING)
-hist(data2$SOCIAL)
-hist(data2$COVID)
-
-#summarize EXERCISE
-summary_exercise <- data2 %>%
-  summarise(
-    Variable = "EXERCISE",
-    Prop_1 = mean(EXERCISE == 1, na.rm = TRUE),
-    Prop_0 = mean(EXERCISE == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(EXERCISE))
-  )
-
-#summarize SLEEP
-summary_sleep <- data2 %>%
-  summarise(
-    Variable = "SLEEP",
-    Prop_1 = mean(SLEEP == 1, na.rm = TRUE),
-    Prop_0 = mean(SLEEP == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(SLEEP))
-  )
-
-#summarize DEPRESSED
-summary_depressed <- data2 %>%
-  summarise(
-    Variable = "DEPRESSED",
-    Prop_1 = mean(DEPRESSED == 1, na.rm = TRUE),
-    Prop_0 = mean(DEPRESSED == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(DEPRESSED))
-  )
-
-#summarize TROUBLE SLEEPING
-summary_trouble_sleeping <- data2 %>%
-  summarise(
-    Variable = "TROUBLE SLEEPING",
-    Prop_1 = mean(TROUBLE_SLEEPING == 1, na.rm = TRUE),
-    Prop_0 = mean(TROUBLE_SLEEPING == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(TROUBLE_SLEEPING))
-  )
-
-#summarize SOCIAL
-summary_social <- data2 %>%
-  summarise(
-    Variable = "SOCIAL",
-    Prop_1 = mean(SOCIAL == 1, na.rm = TRUE),
-    Prop_0 = mean(SOCIAL == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(SOCIAL))
-  )
-
-#summarize COVID
-summary_covid <- data2 %>%
-  summarise(
-    Variable = "COVID",
-    Prop_1 = mean(COVID == 1, na.rm = TRUE),
-    Prop_0 = mean(COVID == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(COVID))
-  )
-
-#combine all summaries and convert to gt table
-summary_table <- bind_rows(
-  summary_exercise, 
-  summary_sleep, 
-  summary_depressed, 
-  summary_trouble_sleeping, 
-  summary_social, 
-  summary_covid
-) %>%
-  gt() %>%
-  tab_header(
-    title = "Proportions of 1, 0, and NA for Various Variables"
-  )
-
-#print the combined table
-summary_table
-
-#save the table as a PNG image
-gtsave(summary_table, here("data", "processed-data", "figures", "summary_table.png"))
-```
-
 For our final step, we will save this data to the processed_data folder.

### 2025-03-15 products/manuscript/Manuscript.qmd
Part 3 commit, complete eda and start analysis

@@
 ---
-title: "Learning Influenza Forecasting"
+title: "Mental Health Factors and Covid"
 author: "Annalise Cramer"
@@
     highlight-style: github
-bibliography: ../../assets/dataanalysis-references.bib
-csl: ../../assets/american-journal-of-epidemiology.csl
+#bibliography: ../../assets/dataanalysis-references.bib
+#csl: ../../assets/american-journal-of-epidemiology.csl
 ---
@@
 library(tidyverse)
+library(tableone)
 library(survey)
+library(gt)
 here()
@@
 ## General Background Information
-_Provide enough background on your topic that others can understand the why and how of your analysis_
+
+During the SARS-CoV pandemic, methods implemented to control the transmission of the disease had broader societal impacts.
+A common complaint surrounding social distancing and quarantine policies is the effects on mental health 1.
+This study aims to better understand the associations between mental health factors and contracting covid.
+Using data from the National Health Interview Surveys collected during the SARS-CoV-2 pandemic, simple and multiple regression models are created.
+
+Cite this properly later:
+1.) https://www.tandfonline.com/doi/full/10.1080/09638237.2020.1757052
 
 ## Description of data and data source
-_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
+
+Data for this project is survey data collected from the National Health Interview Survey, produced by IPUMS through the University of Minnesota.
+This data was collected among United States residents of all ages and genders from 2019 - 2022 in the United States.
 
@@
 
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].
-
 {{< pagebreak >}}
@@
 
-Data for this project is survey data collected from the National Health Interview Survey, produced by IPUMS through the University of Minnesota.
-This data was collected among United States residents of all ages and genders from 2019 - 2022 in the United States.
-There's 151406 observations and 47 vairables.
+There's 151406 observations and 47 variables.
 
-Survey design variables were used and new binary variables were created from other variables.
+Survey design variables were used and new binary variables were created from other variables. Using complex survey design variables, a descriptive table was created.
 
@@
 Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
+fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
 We store those figures in the `assets` folder.
@@
 #| echo: FALSE
-knitr::include_graphics(here("assets","antigen-recognition.png"))
+#knitr::include_graphics(here("assets","antigen-recognition.png"))
 ```
@@
 ## Data aquisition
+
 This data was obtained from IPUMS, used with permission.
@@
 ## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
 
-Raw data called NHIS_COVID.rds, sits with the data folder, in a a subfolder called raw data.
-The processed data, called data3.rds, is within the data folder in a subfolder called processed-data.
-The code to clean and explore and process this dadta is within the code folder, in the processing-code subfolder, called data-cleaning-project.qmd.
+The following code details the data cleaning process.
+Labels are removed from the data to make it easier to work with to create models later in the process.
+Variables needed are selected, then a series of binary varibales are created.
+Additional details are available in the processing folder.
+```{r}
+here() #set working directory
+data_original <- readRDS(here("data","raw-data", "NHIS_COVID.rds"))
+data1 <- labelled::remove_val_labels(data_original) #removes labels from haven package which cause problems
 
-The raw data was loaded, and necessary variables and survery design variables were selected. After exploring each variable, new binary variables were created.
+data1 <- data_original %>%
+  select(STRATA, PSU, PERNUM, SAMPWEIGHT, AGE, PA18AERSTR, HRSLEEP, DEPFREQ, PHQSLEEP, CMPSUPPORT, CVDTESTRSLT)
 
-For now, let us load the processed data.
-Here's a histogram of the processed outcome vairable, which represents whether or not an indivudal has ever tested postive for COVID.
-Histograms of all vairables are available in the data-cleaning-project.qmd file, but will not be shown here for the sake of space.
-```{r}
-library(here)
-data <- read_rds(here("data", "processed-data", "data3.rds"))
-hist(data$COVID)
+data2 <- data1 %>%
+  mutate(
+    PA18AERSTR = ifelse(PA18AERSTR %in% c(0, 7, 8, 9), NA, PA18AERSTR), #put NA, 0, 6, 7, 8, 9 to NA
+    EXERCISE = if_else(PA18AERSTR %in% c(2,3,4), 1, 0, missing = NA_real_) #create EXERCISE variable, 1 if guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met
+    ) %>%
+    mutate(
+    AGE = ifelse(AGE %in% c(997, 999), NA, AGE),  #invalid ages to NA
+    SLEEP = case_when( #Based on age, amount of sleep required from https://www.nhlbi.nih.gov/health/sleep/how-much-sleep#:~:text=Experts%20recommend%20that%20adults%20sleep,or%20more%20hours%20a%20night.
+      AGE < 1  ~ if_else(HRSLEEP >= 12 & HRSLEEP <= 16, 1, 0, missing = NA_real_),
+      AGE >= 1 & AGE < 3       ~ if_else(HRSLEEP >= 11 & HRSLEEP <= 14, 1, 0, missing = NA_real_),
+      AGE >= 3 & AGE < 6       ~ if_else(HRSLEEP >= 10 & HRSLEEP <= 13, 1, 0, missing = NA_real_),
+      AGE >= 6 & AGE < 13      ~ if_else(HRSLEEP >= 9 & HRSLEEP <= 12, 1, 0, missing = NA_real_),
+      AGE >= 13 & AGE <= 18    ~ if_else(HRSLEEP >= 8 & HRSLEEP <= 10, 1, 0, missing = NA_real_),
+      TRUE                     ~ NA_real_  #default NA for unspecified ages
+    )) %>%
+  mutate(
+    DEPRESSED = case_when(
+      DEPFREQ %in% c(1, 2) ~ 1, #daily and weekly is considered depressed
+      DEPFREQ %in% c(3, 4, 5)    ~ 0, #monthly and a few times a year and never are considered not depressed
+      DEPFREQ %in% c(0, 7, 8, 9) ~ NA_real_,#NIU & Unknowns are NA
+      TRUE ~ NA_real_ #default NA for unexpected values
+    )) %>%
+  mutate(
+    TROUBLE_SLEEPING = case_when(
+      PHQSLEEP %in% c(1, 2, 3) ~ 1, #several days, more than half days, nearly every day are trouble sleeping
+      PHQSLEEP == 0            ~ 0,  #not at all is no trouble sleeping
+      PHQSLEEP %in% c(6, 7, 8, 9) ~ NA_real_,  #NIU & Unknowns are NA
+      TRUE ~ NA_real_  #default NA for unexpected values
+    )) %>%
+  mutate(
+    SOCIAL = case_when(
+      CMPSUPPORT == 1 ~ 0,   #more social and emotional support
+      CMPSUPPORT == 2 ~ 1,  #less social and emotional support
+      CMPSUPPORT == 3 ~ 0,   #about the same
+      CMPSUPPORT %in% c(0, 7, 8, 9) ~ NA_real_,  # NIU & Unknowns → NA
+      TRUE ~ NA_real_  #default NA for unexpected values
+    )) %>%
+  mutate(
+    COVID = case_when(
+      CVDTESTRSLT == 2 ~ 1,  #yes, had COVID
+      CVDTESTRSLT == 1 ~ 0,  #no, did not have COVID
+      CVDTESTRSLT %in% c(0, 3, 7, 8, 9) ~ NA_real_,  #NIU, did not receive results, unknowns, and NA are NA
+      TRUE ~ NA_real_  #default NA for unexpected values
+    ))
 ```
@@
 _Explain anything related to your statistical analyses._
+
+We will begin with exploratory tables to explore proportions of NA answers, and generate early estimates.
+
+Our first figure with be a discriptive table stratified by COVID status, with estimates and 95% CIs.
+
+Next, we will construct simple models before moving onto adjusted models.
 
@@
 
+The first table below details the proportion of NAs for each mental health variable.
+We can see this is a little bit high, which may affect analysis later on.
+Histograms of each variable are available in the eda folder.
+
+The second table shows proportions calculated using complex survey design variables.
+
+```{r}
+data <- data2
+#summarize EXERCISE
+summary_exercise <- data %>%
+  summarise(
+    Variable = "EXERCISE",
+    Prop_1 = mean(EXERCISE == 1, na.rm = TRUE),
+    Prop_0 = mean(EXERCISE == 0, na.rm = TRUE),
+    Prop_NA = mean(is.na(EXERCISE))
+  )
+
+#summarize SLEEP
+summary_sleep <- data %>%
+  summarise(
+    Variable = "SLEEP",
+    Prop_1 = mean(SLEEP == 1, na.rm = TRUE),
+    Prop_0 = mean(SLEEP == 0, na.rm = TRUE),
+    Prop_NA = mean(is.na(SLEEP))
+  )
+
+#summarize DEPRESSED
+summary_depressed <- data %>%
+  summarise(
+    Variable = "DEPRESSED",
+    Prop_1 = mean(DEPRESSED == 1, na.rm = TRUE),
+    Prop_0 = mean(DEPRESSED == 0, na.rm = TRUE),
+    Prop_NA = mean(is.na(DEPRESSED))
+  )
+
+#summarize TROUBLE SLEEPING
+summary_trouble_sleeping <- data %>%
+  summarise(
+    Variable = "TROUBLE SLEEPING",
+    Prop_1 = mean(TROUBLE_SLEEPING == 1, na.rm = TRUE),
+    Prop_0 = mean(TROUBLE_SLEEPING == 0, na.rm = TRUE),
+    Prop_NA = mean(is.na(TROUBLE_SLEEPING))
+  )
+
+#summarize SOCIAL
+summary_social <- data %>%
+  summarise(
+    Variable = "SOCIAL",
+    Prop_1 = mean(SOCIAL == 1, na.rm = TRUE),
+    Prop_0 = mean(SOCIAL == 0, na.rm = TRUE),
+    Prop_NA = mean(is.na(SOCIAL))
+  )
+
+#summarize COVID
+summary_covid <- data %>%
+  summarise(
+    Variable = "COVID",
+    Prop_1 = mean(COVID == 1, na.rm = TRUE),
+    Prop_0 = mean(COVID == 0, na.rm = TRUE),
+    Prop_NA = mean(is.na(COVID))
+  )
+
+#combine all summaries and convert to gt table
+summary_table <- bind_rows(
+  summary_exercise, 
+  summary_sleep, 
+  summary_depressed, 
+  summary_trouble_sleeping, 
+  summary_social, 
+  summary_covid
+) %>%
+  gt() %>%
+  tab_header(
+    title = "Proportions of 1, 0, and NA for Various Variables"
+  )
+
+#print the combined table
+summary_table
+
+#list variables
+varstable1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
+             "SOCIAL", "COVID")
+
+#list categories
+cat_vartab1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
+             "SOCIAL", "COVID")
+
+#use function to make table
+Tab1 = CreateTableOne(vars = varstable1, factorVars = cat_vartab1, data = data)
+
+# Convert the TableOne object to a data frame
+tab1_df <- as.data.frame(print(Tab1, showAllLevels = TRUE, quote = FALSE, noSpaces = TRUE))
+
+# Clean up row names
+tab1_df <- tibble::rownames_to_column(tab1_df, "Variable")
+
+# More advanced formatting
+tab1_gt <- gt(tab1_df) %>%
+  # Add a title
+  tab_header(
+    title = md("**COVID test status and Mental Health Factors**"),
+    subtitle = "Survey Results Summary"
+  ) %>%
+  # Format the columns
+  fmt_percent(
+    columns = contains("%"),
+    decimals = 1
+  ) %>%
+  # Group the variables
+  tab_row_group(
+    label = "COVID Status",
+    rows = contains("COVID")
+  ) %>%
+  tab_row_group(
+    label = "Mental Health Indicators",
+    rows = contains(c("DEPRESSED", "TROUBLE_SLEEPING"))
+  ) %>%
+  tab_row_group(
+    label = "Lifestyle Factors",
+    rows = contains(c("EXERCISE", "SLEEP", "SOCIAL"))
+  ) %>%
+  # Style the table
+  tab_style(
+    style = cell_fill(color = "#e6f3ff"),
+    locations = cells_row_groups()
+  ) %>%
+  tab_style(
+    style = list(
+      cell_text(weight = "bold")
+    ),
+    locations = cells_row_groups()
+  ) %>%
+  # Add footnotes if needed
+  tab_footnote(
+    footnote = "Percentages may not sum to 100% due to rounding",
+    locations = cells_column_labels(columns = contains("%"))
+  ) %>%
+  # Adjust column widths
+  cols_width(
+    Variable ~ px(250),
+    everything() ~ px(120)
+  )
+
+tab1_gt
+```
+
 @tbl-summarytable shows a summary of the data.
@@
 #| label: tbl-summarytable
-#| tbl-cap: "Data summary table, proportions of 0 compared to 1, and proportion NA to not NA."
+#| tbl-cap: "Table. 1. Mental health characteristics among Americans from 2019-2022, by SARS-CoV-2 test result status, NHIS"
 #| echo: FALSE
 here()
-knitr::include_graphics(here("data","processed-data", "figures", "summary_table.png"))
+knitr::include_graphics(here("results","figures", "result_table1.png"))
 ```
@@
 
-@fig-result shows a scatterplot figure produced by one of the R scripts.
+fig-result shows a scatterplot figure produced by one of the R scripts.
 
@@
 #| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
+#knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
 ```
@@
 
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
+Example tbl-resulttable2 shows a summary of a linear model fit.
 

### 2025-04-20 code/analysis-code/analysis.qmd
Part 5 analysis

@@
 library(tableone)
+library(mice)
+library(mitools)
+library(miceadds)
+library(tidymodels)
+library(caret)
 library(gt)
@@
 here() #set working directory
-data <- readRDS(here("data","processed-data", "data3.rds"))
-data <- labelled::remove_val_labels(data) #removes labels from haven package which cause problems
+#data <- readRDS(here("data","processed-data", "data3.rds"))
+#data <- labelled::remove_val_labels(data) #removes labels from haven package which cause problems
+
+data_imputed <- readRDS(here("data","processed-data", "binary_imputed_datasets.rds"))
 ```
@@
 COVID: If they've tested positive for covid or not
+
+We create estimates and CIs using complex survery design variables, using the multiple imputed data.
+
+```{r}
+
+# Create survey design objects for each transformed dataset
+svy_designs <- lapply(data_imputed, function(data) {
+  svydesign(ids = ~PSU, strata = ~STRATA, weights = ~SAMPWEIGHT, data = data , nest=TRUE)
+})
+
+# Function to get complete estimates (both 0 and 1) stratified by COVID status
+get_complete_stratified_estimates <- function(var_name) {
+  results <- data.frame(
+    Variable = character(),
+    COVID_Status = character(),
+    Binary_Value = numeric(),
+    Estimate = numeric(),
+    CI_lower = numeric(),
+    CI_upper = numeric(),
+    stringsAsFactors = FALSE
+  )
+
+  # Loop through COVID status (0 and 1)
+  for (covid_val in c(0, 1)) {
+    covid_label <- ifelse(covid_val == 0, "Negative", "Positive")
+
+    # Calculate proportions for this COVID status
+    props <- lapply(svy_designs, function(design) {
+      # Create a subset design
+      subdesign <- subset(design, COVID == covid_val)
+      # Create formula for the variable
+      formula <- as.formula(paste0("~factor(", var_name, ")"))
+      # Calculate proportions
+      svymean(formula, subdesign, na.rm = TRUE)
+    })
+
+    # Pool results
+    pooled <- MIcombine(props)
+
+    # Extract results for both binary values (if available)
+    coef_names <- names(coef(pooled))
+
+    # Handle different potential naming patterns
+    for (bin_val in c(0, 1)) {
+      # Check different possible coefficient names
+      possible_names <- c(
+        paste0("factor(", var_name, ")", bin_val),
+        as.character(bin_val)
+      )
+
+      # Find the matching coefficient name
+      coef_idx <- NULL
+      for (name in possible_names) {
+        if (name %in% coef_names) {
+          coef_idx <- which(coef_names == name)
+          break
+        }
+      }
+
+      # If found, extract estimates
+      if (!is.null(coef_idx)) {
+        est <- coef(pooled)[coef_idx]
+        ci <- confint(pooled)[coef_idx, ]
+
+        # Add to results
+        results <- rbind(results, data.frame(
+          Variable = var_name,
+          COVID_Status = covid_label,
+          Binary_Value = bin_val,
+          Estimate = round(est * 100, 1),
+          CI_lower = round(ci[1] * 100, 1),
+          CI_upper = round(ci[2] * 100, 1)
+        ))
+      }
+    }
+  }
+
+  return(results)
+}
+
+# Variables to analyze
+variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
+
+# Get complete stratified estimates for each variable
+complete_results <- lapply(variables, get_complete_stratified_estimates)
+
+# Combine results into a single data frame
+complete_df <- do.call(rbind, complete_results)
+
+# Reshape for better presentation
+formatted_results <- complete_df %>%
+  mutate(Binary_Label = ifelse(Binary_Value == 1, "Yes", "No")) %>%
+  select(Variable, COVID_Status, Binary_Label, Estimate, CI_lower, CI_upper) %>%
+  arrange(Variable, COVID_Status, desc(Binary_Label))
+
+print(formatted_results)
+
+```
+
+Now let's turn that into a nice table using GT, and I'll save it to use in the manuscript later. 
+```{r}
+#checking
+summary(formatted_results)
+str(formatted_results)
+names(formatted_results)
+
+# Prepare the data by combining Variable and Binary_Label, pivoting, and formatting the results
+table1_imputed <- formatted_results %>%
+  mutate(Binary_Label = ifelse(Binary_Label == "Yes", 1, 0),  # Convert Yes to 1 and No to 0
+         Variable_Label = paste(Variable, Binary_Label, sep = " = ")) %>%
+  pivot_wider(
+    names_from = COVID_Status,
+    values_from = c(Estimate, CI_lower, CI_upper),
+    names_glue = "{COVID_Status}_{.value}"
+  ) %>%
+  mutate(
+    # Scale the estimates and CI by 0.01
+    Estimate_Negative = Negative_Estimate * 0.01,
+    CI_Negative = sprintf("(%.2f, %.2f)", Negative_CI_lower * 0.01, Negative_CI_upper * 0.01),
+    Estimate_Positive = Positive_Estimate * 0.01,
+    CI_Positive = sprintf("(%.2f, %.2f)", Positive_CI_lower * 0.01, Positive_CI_upper * 0.01)
+  ) %>%
+  select(Variable_Label, Estimate_Negative, CI_Negative, Estimate_Positive, CI_Positive) %>%
+  gt() %>%
+  tab_header(
+    title = "Estimated proportions of mental health related factors by COVID test status"
+  ) %>%
+  tab_spanner(
+    label = "COVID Negative",
+    columns = c(Estimate_Negative, CI_Negative)
+  ) %>%
+  tab_spanner(
+    label = "COVID Positive",
+    columns = c(Estimate_Positive, CI_Positive)
+  ) %>%
+  cols_label(
+    Variable_Label = "Variable",               # Rename Variable_Label to Variable
+    Estimate_Negative = "Estimate",            # Rename Estimate_Negative to Estimate
+    CI_Negative = "95% CI",                    # Rename CI_Negative to 95% CI
+    Estimate_Positive = "Estimate",            # Rename Estimate_Positive to Estimate
+    CI_Positive = "95% CI"                     # Rename CI_Positive to 95% CI
+  )
+
+# Save the table as a PNG
+here()
+gtsave(table1_imputed, here("results", "figures", "table1_imputed.png"))
+```
+################################################################################
+Now, we will start modeling COVID postivity as an outcome of mental health factors.
+
+Up first, we will create singleton models to individually study the effects of mental health factors on covid test positivity status.
+
+```{r}
+# Create a survey design object for each completed dataset
+survey_designs <- lapply(data_imputed, function(data_i) {
+  svydesign(
+    id = ~PSU,
+    strata = ~STRATA,
+    weights = ~SAMPWEIGHT,
+    nest = TRUE,
+    data = data_i
+  )
+})
+
+# Fit svyglm models for EXERCISE predictor
+models_exercise <- lapply(survey_designs, function(dsgn) {
+  svyglm(COVID ~ EXERCISE, design = dsgn, family = quasibinomial())
+})
+
+# Extract coefficients and standard errors for each model
+model_exercise_coef <- lapply(models_exercise, function(model) {
+  data.frame(
+    Estimate = coef(model),
+    Std.Error = sqrt(diag(vcov(model))),
+    stringsAsFactors = FALSE
+  )
+})
+
+# Calculate means and variances for pooling
+estimates <- do.call(rbind, lapply(model_exercise_coef, function(x) x$Estimate))
+se <- do.call(rbind, lapply(model_exercise_coef, function(x) x$Std.Error))
+
+# Compute the mean across imputations
+mean_estimate <- apply(estimates, 2, mean)
+
+# Compute within-imputation variance (average variance across imputations)
+within_variance <- apply(se^2, 2, mean)
+
+# Compute between-imputation variance (variance of the estimates across imputations)
+between_variance <- apply(estimates, 2, var)
+
+# Calculate the total variance using Rubin's rule
+total_variance <- within_variance + (1 + 1 / length(survey_designs)) * between_variance
+
+# Calculate the pooled standard error
+pooled_se <- sqrt(total_variance)
+
+# Calculate the 95% confidence intervals
+ci_low <- mean_estimate - 1.96 * pooled_se
+ci_high <- mean_estimate + 1.96 * pooled_se
+
+# Create a data frame with pooled estimates, standard errors, and confidence intervals
+pooled_exercise <- data.frame(
+  Estimate = mean_estimate,
+  Std.Error = pooled_se,
+  CI_Low = ci_low,
+  CI_High = ci_high
+)
+
+# Display the pooled results
+pooled_exercise
+
+library(survey)
+
+# Fit svyglm models for EXERCISE predictor (and other predictors if needed)
+predictors <- c("EXERCISE", "DEPRESSED", "SLEEP", "TROUBLE_SLEEPING", "SOCIAL")
+
+# Fit models for each predictor
+models <- lapply(predictors, function(predictor) {
+  lapply(survey_designs, function(dsgn) {
+    formula <- as.formula(paste("COVID ~", predictor))
+    svyglm(formula, design = dsgn, family = quasibinomial())
+  })
+})
+
+# Extract coefficients and standard errors for each model
+model_coef <- mapply(function(predictor, predictor_models) {
+  lapply(predictor_models, function(model) {
+    data.frame(
+      Estimate = coef(model),
+      Std.Error = sqrt(diag(vcov(model))),
+      stringsAsFactors = FALSE
+    )
+  })
+}, predictors, models, SIMPLIFY = FALSE)
+
+# Calculate means, variances, and p-values for pooling
+pooled_results <- lapply(names(model_coef), function(predictor_name) {
+  coefs <- model_coef[[predictor_name]]
+
+  estimates <- do.call(rbind, lapply(coefs, function(x) x$Estimate))
+  se <- do.call(rbind, lapply(coefs, function(x) x$Std.Error))
+
+  # Compute the mean across imputations
+  mean_estimate <- apply(estimates, 2, mean)
+
+  # Compute within-imputation variance (average variance across imputations)
+  within_variance <- apply(se^2, 2, mean)
+
+  # Compute between-imputation variance (variance of the estimates across imputations)
+  between_variance <- apply(estimates, 2, var)
+
+  # Calculate the total variance using Rubin's rule
+  total_variance <- within_variance + (1 + 1 / length(survey_designs)) * between_variance
+
+  # Calculate the pooled standard error
+  pooled_se <- sqrt(total_variance)
+
+  # Calculate the 95% confidence intervals
+  ci_low <- mean_estimate - 1.96 * pooled_se
+  ci_high <- mean_estimate + 1.96 * pooled_se
+
+  # Calculate z-statistics and p-values
+  z_stat <- mean_estimate / pooled_se
+  p_values <- 2 * (1 - pnorm(abs(z_stat)))  # Two-tailed p-value
+
+  # Exponentiate to get Odds Ratios (ORs) from log-odds
+  or_estimate <- exp(mean_estimate)
+  ci_low_or <- exp(ci_low)
+  ci_high_or <- exp(ci_high)
+
+  # Create a data frame with pooled Odds Ratios, standard errors, confidence intervals, and p-values
+  data.frame(
+    Predictor = rep(predictor_name, length(mean_estimate)),
+    OR = or_estimate,
+    Std.Error = pooled_se,
+    CI_Low = ci_low_or,
+    CI_High = ci_high_or,
+    p_value = p_values,
+    row.names = names(mean_estimate)
+  )
+})
+
+# Combine results for all predictors into one data frame
+pooled_exercise <- do.call(rbind, pooled_results)
+
+# Display the labeled pooled results
+pooled_exercise
+
+```
+
+Now we will make a nice table of that using gt.
+
+```{r}
+names(pooled_exercise)
+str(pooled_exercise)
+
+#not including intercepts values
+# Assuming 'pooled_exercise' is your data frame
+pooled_exercise %>%
+  group_by(Predictor) %>%
+  slice(2) %>%  # Select only the second row for each Predictor
+  ungroup() %>%
+  mutate(CI = paste0("(", round(CI_Low, 3), ", ", round(CI_High, 3), ")")) %>%
+  select(Predictor, OR, CI, p_value) %>%
+  gt() %>%
+  tab_header(
+    title = "Associations of individual mental health factors with positive covid test result"
+  ) %>%
+  cols_label(
+    Predictor = "Predictor",
+    OR = "Odds Ratio",
+    CI = "Confidence Interval",
+    p_value = "P-Value"
+  ) %>%
+  fmt_number(
+    columns = vars(OR, p_value),
+    decimals = 3
+  )
+
+```
+
+################################################################################
+This is the part of the project in which I demonstrate complex techniques learned during this class.
+
+Next, we will make the multivariate model. I will use cross validation for this. The RMSE is found to be 0.4450602.
+```{r}
+
+cross_validation_imputed_rmse <- function(imputed_data, design, predictors, k_folds = 5) {
+  rmse_list <- numeric(length(imputed_data))  # Store RMSE for each imputation
+
+  for (i in 1:length(imputed_data)) {
+    data <- imputed_data[[i]]
+    set.seed(123)
+    folds <- sample(1:k_folds, size = nrow(data), replace = TRUE)
+
+    fold_rmse <- numeric(k_folds)  # RMSE for each fold
+
+    for (fold in 1:k_folds) {
+      train_data <- data[folds != fold, ]
+      test_data  <- data[folds == fold, ]
+
+      # Survey design for training and testing
+      train_design <- svydesign(ids = ~1, data = train_data, weights = ~SAMPWEIGHT)
+      formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
+      model <- svyglm(formula, design = train_design, family = quasibinomial())
+
+      # Predictions
+      predictions <- predict(model, newdata = test_data, type = "response")
+      actual <- test_data$COVID
+
+      # Compute RMSE for this fold
+      fold_rmse[fold] <- sqrt(mean((predictions - actual)^2, na.rm = TRUE))
+    }
+
+    # Average RMSE over folds for this imputed dataset
+    rmse_list[i] <- mean(fold_rmse)
+  }
+
+  # Average RMSE over all imputed datasets
+  overall_rmse <- mean(rmse_list)
+
+  return(list(rmse_per_imputation = rmse_list, overall_rmse = overall_rmse))
+}
+
+# Assuming you have your imputed_data and a vector of predictors
+results <- cross_validation_imputed_rmse(imputed_data = data_imputed,
+                                         design = NULL,  # Not used inside, can be omitted
+                                         predictors = c("AGE", "SLEEP", "DEPRESSED"))
+
+# View results
+results$rmse_per_imputation  # RMSE for each imputed dataset
+results$overall_rmse         # Average RMSE across imputations
+```
+
+Now, we will make a LASSO regression model. For this, I'm not going to use the multiple imputation dataset. I'm just going to pick a sinlge imputed dataset, the first one. The RMSE is 0.276.
+
+```{r}
+# Extract the first completed imputed dataset
+data1 <- data_imputed[[1]]  # You can loop over 1:5 later
+
+# Then define the survey design
+survey_design <- svydesign(
+  ids = ~PSU,
+  strata = ~STRATA,
+  weights = ~SAMPWEIGHT,
+  data = data1,
+  nest=TRUE
+)
+
+set.seed(1234)  # for reproducibility
+
+# Choose your outcome and predictors
+outcome_var <- "DEPRESSED"
+predictor_vars <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
+
+# Loop over imputations and compute RMSE
+rmse_list <- map(data_imputed, function(dat) {
+  # Drop missing in predictors or outcome if any slipped through
+  dat <- dat %>% drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))
+
+  # Create design matrix X and response y
+  X <- model.matrix(as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))), dat)[, -1]
+  y <- dat[[outcome_var]]
+  w <- dat$SAMPWEIGHT
+
+  # Split into training and test sets (e.g., 70/30 split)
+  train_idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
+  test_idx <- setdiff(seq_len(nrow(X)), train_idx)
+
+  X_train <- X[train_idx, ]
+  y_train <- y[train_idx]
+  w_train <- w[train_idx]
+
+  X_test <- X[test_idx, ]
+  y_test <- y[test_idx]
+  w_test <- w[test_idx]
+
+  # Fit weighted lasso model
+  cv_fit <- cv.glmnet(
+    x = X_train,
+    y = y_train,
+    alpha = 1,
+    weights = w_train,
+    family = "binomial",  # change to "binomial" for binary outcome
+    nfolds = 5
+  )
+
+  # Predict on test set
+  preds <- predict(cv_fit, newx = X_test, s = "lambda.min")
+
+  # Weighted RMSE
+  sqrt(weighted.mean((y_test - preds)^2, w_test))
+})
+
+# Pool RMSE across imputations
+pooled_rmse <- mean(unlist(rmse_list))
+
+# Output
+cat("RMSE for each imputation:\n")
+print(unlist(rmse_list))
+cat("\nPooled RMSE across imputations:", round(pooled_rmse, 3), "\n")
+
+```
+
+################################################################################
+The following code was used prior to development of imputed data and is no longer relevant.
+It may be ignored.
+I am keeping it in case it serves as a useful reference for the rest of the project or future endeavors.
+################################################################################
 
@@
 ```
+
+################################################################################
+
+Up next, we will run the singleton models.
+For these, our outcome variable (covid test result) will be used as a variable for each of the respective predictor vairbales that we created in the processing code.
+These models appear as the following:
+
+$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Exercise} $$
+$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Sleep} $$
+$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Depressed} $$
+$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{TroubleSleeping} $$
+$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Social} $$
+Create survey design.
+```{r}
+#review names to use below
+names(data)
+
+#create survey object using design variables
+survey_design <- svydesign(
+  ids = ~PSU,
+  strata = ~STRATA,
+  weights = ~SAMPWEIGHT,
+  data = data,
+  nest = TRUE
+)
+
+pr_results <- data.frame()
+
+variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
+               "SOCIAL")
+
+for (var in variables) {
+  # Create the formula dynamically
+  formula <- as.formula(paste("COVID ~", var))
+
+  # Fit the model
+  model <- svyglm(
+    formula, 
+    design = survey_design, 
+    family = quasibinomial(link = "log")
+  )
+
+  # Extract prevalence ratios and confidence intervals
+  pr <- exp(coef(model))  # Prevalence ratios
+  pr_ci <- exp(confint(model))  # Confidence intervals
+
+  # Extract p-values
+  p_values <- summary(model)$coefficients[, "Pr(>|t|)"]
+
+  # Combine results into a single table
+  pr_results <- rbind(
+    pr_results,
+    data.frame(
+      Variable = var,
+      Estimate = pr[-1],  # Skip the intercept
+      `2.5% CI` = pr_ci[-1, 1],
+      `97.5% CI` = pr_ci[-1, 2],
+      `p-value` = p_values[-1]  # Skip the intercept
+    )
+  )
+}
+
+# Print the results
+print(pr_results)
+
+```
+
+Now, we will make that into a nice quality table using the gt package and the help of chatGPT and Claude.
+```{r}
+#checking
+glimpse(pr_results)
+str(pr_results)
+
+#create the gt table
+pr_results %>%
+  mutate(
+    Predictor = stringr::str_to_title(gsub("_", " ", Variable)),
+    `Prevalence Ratio` = round(Estimate, 2),
+    `95% CI` = paste0("(", round(X2.5..CI, 2), ", ", round(X97.5..CI, 2), ")"),
+    `p-value` = format.pval(p.value, digits = 3, eps = .001)
+  ) %>%
+  select(Predictor, `Prevalence Ratio`, `95% CI`, `p-value`) %>%
+  gt() %>%
+  tab_header(
+    title = md("**Positive COVID Test Result Association with Mental Health Factors**")
+  ) %>%
+  tab_options(
+    column_labels.border.bottom.width = px(1.5),  # Bold line under column headers
+    column_labels.border.bottom.color = "black"
+  )
+
+```
+
+Now, we're going to complete the process of model selection, using backwards selection. We want to end up with a model we can use for prediction purposes. Our initial model looks like this: 
+$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Exercise} + \beta_2 \cdot X_{Sleep} + \beta_3 \cdot X_{Depressed} + \beta_4 \cdot X_{TroubleSleeping} + \beta_5 \cdot X_{Social} $$
+Let's find the AIC value. 
+```{r}
+
+unique(data$COVID)
+model_full <- svyglm(
+  COVID ~ EXERCISE + SLEEP + DEPRESSED + TROUBLE_SLEEPING + SOCIAL,
+  design = survey_design,
+  family = quasibinomial(link = "log")
+)
+# Perform backward selection based on AIC
+model_step <- step(model_full, direction = "backward", trace = 1)
+summary(model_step)
+
+```

### 2025-04-20 code/eda-code/eda.qmd
Part 5 analysis

@@
 ---
-title: "Data-Cleaning"
+title: "Exploratory Data Analysis"
 ---
@@
 library(webshot2)
+library(survey)
+library(mice)
+library(CreateTableOne)
 library(here)
@@
 data <- readRDS(here("data","processed-data", "data3.rds"))
+data <- labelled::remove_val_labels(data) #removes labels from haven package which cause problems
 ```
@@
 ```
+
+Now, let's examine the imputed datasets to ensure they're not containing NAs. I'm only going to display this for one of them, given they were all created the same way, checking one will be adequate.
+```{r}
+#read it in
+here()
+data_imputed <- readRDS(here("data", "processed-data", "binary_imputed_datasets.rds"))
+
+# Use the first completed dataset from the list
+data_imputed_df1 <- data_imputed[[1]]
+
+#select variables we want
+vars <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL", "COVID")
+
+#reshape data
+long_data <- data_imputed_df1 %>%
+  select(all_of(vars)) %>%
+  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
+  filter(!is.na(value))
+
+# Create histograms
+ggplot(long_data, aes(x = value)) +
+  geom_histogram(bins = 30, color = "black", fill = "skyblue") +
+  facet_wrap(~ variable, scales = "free") +   # Facet by variable for separate histograms
+  theme_minimal() +
+  labs(title = "Histograms of Imputed Variables", x = "Value", y = "Frequency")
+
+#checking that there's no NA's in the data
+na_count <- colSums(is.na(long_data))
+na_count
+```

### 2025-04-20 code/processing-code/processing.qmd
Part 5 analysis

@@
 library(webshot2)
+library(survey)
+library(mice)
+library(tableone) #added from part 3 feedback
+library(mitools)
 library(here)
@@
 ```
+
+After completed EDA, there's significant missingness in the data.
+At the recommendation of my peers, I will use multiple imputation, creating 5 sets of imputed data.
+I re-create the binary variables in each of the 5 imputed datasets, then save to use in the EDA file.
+```{r}
+# First, let's set up the imputation
+# Create a subset of variables for imputation
+data1 <- labelled::remove_val_labels(data_original) #removes labels from haven package which cause problems
+imputation_vars <- c("AGE", "PA18AERSTR", "HRSLEEP", "DEPFREQ", 
+                     "PHQSLEEP", "CMPSUPPORT", "CVDTESTRSLT", 
+                     "STRATA", "PSU", "SAMPWEIGHT")
+
+# Create a subset of your data with just these variables
+data_imp <- data1[, imputation_vars]
+
+# Examine the missing data pattern
+md.pattern(data_imp)
+
+# Initialize the imputation - including survey design variables in the imputation model
+# but not imputing them
+imp <- mice(data_imp, maxit = 0)
+
+# Customize the predictor matrix to account for complex survey design
+pred <- imp$predictorMatrix
+pred[c("STRATA", "PSU", "SAMPWEIGHT"), ] <- 0  # Don't use these as outcomes
+pred[, c("STRATA", "PSU", "SAMPWEIGHT")] <- 1  # Use these as predictors
+
+# Customize the method - ensure we don't impute design variables
+meth <- imp$method
+meth[c("STRATA", "PSU", "SAMPWEIGHT")] <- ""
+
+# Run multiple imputation with 5 imputations (increase m for better estimates)
+imp_data <- mice(data_imp, m = 5, predictorMatrix = pred, method = meth, maxit = 5)
+
+# Check convergence
+plot(imp_data)
+
+# Get a single completed dataset (for example, the first one)
+completed_data1 <- complete(imp_data, 1)
+
+# For analysis with complex survey design
+# Create a survey design object with the imputed data
+# Repeat for each imputed dataset and pool results
+
+# Get all 5 imputed datasets
+imputed_datasets <- lapply(1:5, function(i) complete(imp_data, i))
+
+# Create a function to transform variables into binary using your specific rules
+create_binary_vars <- function(df) {
+  # Clean AGE and PA18AERSTR
+  df$AGE[df$AGE %in% c(997, 999)] <- NA
+  df$PA18AERSTR[df$PA18AERSTR %in% c(0, 7, 8, 9)] <- NA
+
+  # EXERCISE
+  df$EXERCISE <- ifelse(df$PA18AERSTR %in% c(2, 3, 4), 1,
+                        ifelse(is.na(df$PA18AERSTR), NA, 0))
+
+  # SLEEP
+  df$SLEEP <- NA_real_
+  df$SLEEP[which(df$AGE < 1)] <- ifelse(df$HRSLEEP[which(df$AGE < 1)] >= 12 &
+                                          df$HRSLEEP[which(df$AGE < 1)] <= 16, 1, 0)
+  df$SLEEP[which(df$AGE >= 1 & df$AGE < 3)] <- ifelse(df$HRSLEEP[which(df$AGE >= 1 & df$AGE < 3)] >= 11 &
+                                                       df$HRSLEEP[which(df$AGE >= 1 & df$AGE < 3)] <= 14, 1, 0)
+  df$SLEEP[which(df$AGE >= 3 & df$AGE < 6)] <- ifelse(df$HRSLEEP[which(df$AGE >= 3 & df$AGE < 6)] >= 10 &
+                                                       df$HRSLEEP[which(df$AGE >= 3 & df$AGE < 6)] <= 13, 1, 0)
+  df$SLEEP[which(df$AGE >= 6 & df$AGE < 13)] <- ifelse(df$HRSLEEP[which(df$AGE >= 6 & df$AGE < 13)] >= 9 &
+                                                        df$HRSLEEP[which(df$AGE >= 6 & df$AGE < 13)] <= 12, 1, 0)
+  df$SLEEP[which(df$AGE >= 13 & df$AGE <= 18)] <- ifelse(df$HRSLEEP[which(df$AGE >= 13 & df$AGE <= 18)] >= 8 &
+                                                           df$HRSLEEP[which(df$AGE >= 13 & df$AGE <= 18)] <= 10, 1, 0)
+
+  # DEPRESSED
+  df$DEPRESSED <- NA_real_
+  df$DEPRESSED[df$DEPFREQ %in% c(1, 2)] <- 1
+  df$DEPRESSED[df$DEPFREQ %in% c(3, 4, 5)] <- 0
+  df$DEPRESSED[df$DEPFREQ %in% c(0, 7, 8, 9)] <- NA
+
+  # TROUBLE_SLEEPING
+  df$TROUBLE_SLEEPING <- NA_real_
+  df$TROUBLE_SLEEPING[df$PHQSLEEP %in% c(1, 2, 3)] <- 1
+  df$TROUBLE_SLEEPING[df$PHQSLEEP == 0] <- 0
+  df$TROUBLE_SLEEPING[df$PHQSLEEP %in% c(6, 7, 8, 9)] <- NA
+
+  # SOCIAL
+  df$SOCIAL <- NA_real_
+  df$SOCIAL[df$CMPSUPPORT == 1] <- 0
+  df$SOCIAL[df$CMPSUPPORT == 2] <- 1
+  df$SOCIAL[df$CMPSUPPORT == 3] <- 0
+  df$SOCIAL[df$CMPSUPPORT %in% c(0, 7, 8, 9)] <- NA
+
+  # COVID
+  df$COVID <- NA_real_
+  df$COVID[df$CVDTESTRSLT == 2] <- 1
+  df$COVID[df$CVDTESTRSLT == 1] <- 0
+  df$COVID[df$CVDTESTRSLT %in% c(0, 3, 7, 8, 9)] <- NA
+
+  return(df)
+}
+
+# Apply transformation to each imputed dataset
+binary_imputed_datasets <- lapply(imputed_datasets, create_binary_vars)
+
+glimpse(binary_imputed_datasets)
+
+#save for later
+here()
+saveRDS(binary_imputed_datasets, file = here("data", "processed-data", "binary_imputed_datasets.rds"))
+```

### 2025-04-20 products/manuscript/Manuscript.qmd
Part 5 analysis

@@
 library(tidyverse)
+library(mice)
+library(miceadds)
+library(mitools)
 library(tableone)

### 2025-04-20 products/manuscript/Manuscript.qmd
Part 5 working on manuscript

@@
 # Summary/Abstract
-_Write a summary of your project._
+
+Isolation and quarantine strategies employed for containment of SARS-Cov-2 infection led to unexpected mental health struggles among the population of the United States.
+Infected individuals and healthy individuals were affected, but little is known about the potential relationships between mental health and physical health.
+In this study, we utilized Nataional Health Information Surveys from 2019-2021 to examine the relationships between receieving a positive COVID-19 test and self reporting mental health factors such as meeting recommended guidelines for exercise, meeting the recommended number of hours of sleep per night, reporting feeling depressed recently, reporting having trouble sleeping recently, and reporting receiving less social support.
+
+We find
 
@@
 
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key.
-Then you can include like this:
-
 {{< pagebreak >}}
@@
 
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
+Data was subsetted to years 2019-2022, to cover the brunt of pandemic years.
+Despite low prevelance of COVID-19 cases in 2019 worldwide, the appearance of COVID-19 in the news (and consequential potential impact on mental health factors) led us to include this data.
+The final dataset contains 151406 observations.
 
-There's 151406 observations and 47 variables.
+Variables of interest contained significant amounts of missing information, so multiple imputation was used, creating 5 sets of imputed data that was averaged across the remainder of analysis to produce variables and statistics needed.
+Complex survey design was used to allow for generalizability of results, using vairbales PSU, STRATA, and SAMPWIEGHT.
+
+Binary variables were created to allow for interpretability.
+
+~~~
 
@@
 Additional details are available in the processing folder.
-```{r}
-here() #set working directory
-data_original <- readRDS(here("data","raw-data", "NHIS_COVID.rds"))
-data1 <- labelled::remove_val_labels(data_original) #removes labels from haven package which cause problems
-
-data1 <- data_original %>%
-  select(STRATA, PSU, PERNUM, SAMPWEIGHT, AGE, PA18AERSTR, HRSLEEP, DEPFREQ, PHQSLEEP, CMPSUPPORT, CVDTESTRSLT)
-
-data2 <- data1 %>%
-  mutate(
-    PA18AERSTR = ifelse(PA18AERSTR %in% c(0, 7, 8, 9), NA, PA18AERSTR), #put NA, 0, 6, 7, 8, 9 to NA
-    EXERCISE = if_else(PA18AERSTR %in% c(2,3,4), 1, 0, missing = NA_real_) #create EXERCISE variable, 1 if guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met
-    ) %>%
-    mutate(
-    AGE = ifelse(AGE %in% c(997, 999), NA, AGE),  #invalid ages to NA
-    SLEEP = case_when( #Based on age, amount of sleep required from https://www.nhlbi.nih.gov/health/sleep/how-much-sleep#:~:text=Experts%20recommend%20that%20adults%20sleep,or%20more%20hours%20a%20night.
-      AGE < 1  ~ if_else(HRSLEEP >= 12 & HRSLEEP <= 16, 1, 0, missing = NA_real_),
-      AGE >= 1 & AGE < 3       ~ if_else(HRSLEEP >= 11 & HRSLEEP <= 14, 1, 0, missing = NA_real_),
-      AGE >= 3 & AGE < 6       ~ if_else(HRSLEEP >= 10 & HRSLEEP <= 13, 1, 0, missing = NA_real_),
-      AGE >= 6 & AGE < 13      ~ if_else(HRSLEEP >= 9 & HRSLEEP <= 12, 1, 0, missing = NA_real_),
-      AGE >= 13 & AGE <= 18    ~ if_else(HRSLEEP >= 8 & HRSLEEP <= 10, 1, 0, missing = NA_real_),
-      TRUE                     ~ NA_real_  #default NA for unspecified ages
-    )) %>%
-  mutate(
-    DEPRESSED = case_when(
-      DEPFREQ %in% c(1, 2) ~ 1, #daily and weekly is considered depressed
-      DEPFREQ %in% c(3, 4, 5)    ~ 0, #monthly and a few times a year and never are considered not depressed
-      DEPFREQ %in% c(0, 7, 8, 9) ~ NA_real_,#NIU & Unknowns are NA
-      TRUE ~ NA_real_ #default NA for unexpected values
-    )) %>%
-  mutate(
-    TROUBLE_SLEEPING = case_when(
-      PHQSLEEP %in% c(1, 2, 3) ~ 1, #several days, more than half days, nearly every day are trouble sleeping
-      PHQSLEEP == 0            ~ 0,  #not at all is no trouble sleeping
-      PHQSLEEP %in% c(6, 7, 8, 9) ~ NA_real_,  #NIU & Unknowns are NA
-      TRUE ~ NA_real_  #default NA for unexpected values
-    )) %>%
-  mutate(
-    SOCIAL = case_when(
-      CMPSUPPORT == 1 ~ 0,   #more social and emotional support
-      CMPSUPPORT == 2 ~ 1,  #less social and emotional support
-      CMPSUPPORT == 3 ~ 0,   #about the same
-      CMPSUPPORT %in% c(0, 7, 8, 9) ~ NA_real_,  # NIU & Unknowns → NA
-      TRUE ~ NA_real_  #default NA for unexpected values
-    )) %>%
-  mutate(
-    COVID = case_when(
-      CVDTESTRSLT == 2 ~ 1,  #yes, had COVID
-      CVDTESTRSLT == 1 ~ 0,  #no, did not have COVID
-      CVDTESTRSLT %in% c(0, 3, 7, 8, 9) ~ NA_real_,  #NIU, did not receive results, unknowns, and NA are NA
-      TRUE ~ NA_real_  #default NA for unexpected values
-    ))
-```
 
@@
 
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
-
 The first table below details the proportion of NAs for each mental health variable.
@@
 
-```{r}
-data <- data2
-#summarize EXERCISE
-summary_exercise <- data %>%
-  summarise(
-    Variable = "EXERCISE",
-    Prop_1 = mean(EXERCISE == 1, na.rm = TRUE),
-    Prop_0 = mean(EXERCISE == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(EXERCISE))
-  )
-
-#summarize SLEEP
-summary_sleep <- data %>%
-  summarise(
-    Variable = "SLEEP",
-    Prop_1 = mean(SLEEP == 1, na.rm = TRUE),
-    Prop_0 = mean(SLEEP == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(SLEEP))
-  )
-
-#summarize DEPRESSED
-summary_depressed <- data %>%
-  summarise(
-    Variable = "DEPRESSED",
-    Prop_1 = mean(DEPRESSED == 1, na.rm = TRUE),
-    Prop_0 = mean(DEPRESSED == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(DEPRESSED))
-  )
-
-#summarize TROUBLE SLEEPING
-summary_trouble_sleeping <- data %>%
-  summarise(
-    Variable = "TROUBLE SLEEPING",
-    Prop_1 = mean(TROUBLE_SLEEPING == 1, na.rm = TRUE),
-    Prop_0 = mean(TROUBLE_SLEEPING == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(TROUBLE_SLEEPING))
-  )
-
-#summarize SOCIAL
-summary_social <- data %>%
-  summarise(
-    Variable = "SOCIAL",
-    Prop_1 = mean(SOCIAL == 1, na.rm = TRUE),
-    Prop_0 = mean(SOCIAL == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(SOCIAL))
-  )
-
-#summarize COVID
-summary_covid <- data %>%
-  summarise(
-    Variable = "COVID",
-    Prop_1 = mean(COVID == 1, na.rm = TRUE),
-    Prop_0 = mean(COVID == 0, na.rm = TRUE),
-    Prop_NA = mean(is.na(COVID))
-  )
-
-#combine all summaries and convert to gt table
-summary_table <- bind_rows(
-  summary_exercise, 
-  summary_sleep, 
-  summary_depressed, 
-  summary_trouble_sleeping, 
-  summary_social, 
-  summary_covid
-) %>%
-  gt() %>%
-  tab_header(
-    title = "Proportions of 1, 0, and NA for Various Variables"
-  )
-
-#print the combined table
-summary_table
-
-#list variables
-varstable1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
-             "SOCIAL", "COVID")
-
-#list categories
-cat_vartab1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
-             "SOCIAL", "COVID")
-
-#use function to make table
-Tab1 = CreateTableOne(vars = varstable1, factorVars = cat_vartab1, data = data)
-
-# Convert the TableOne object to a data frame
-tab1_df <- as.data.frame(print(Tab1, showAllLevels = TRUE, quote = FALSE, noSpaces = TRUE))
-
-# Clean up row names
-tab1_df <- tibble::rownames_to_column(tab1_df, "Variable")
-
-# More advanced formatting
-tab1_gt <- gt(tab1_df) %>%
-  # Add a title
-  tab_header(
-    title = md("**COVID test status and Mental Health Factors**"),
-    subtitle = "Survey Results Summary"
-  ) %>%
-  # Format the columns
-  fmt_percent(
-    columns = contains("%"),
-    decimals = 1
-  ) %>%
-  # Group the variables
-  tab_row_group(
-    label = "COVID Status",
-    rows = contains("COVID")
-  ) %>%
-  tab_row_group(
-    label = "Mental Health Indicators",
-    rows = contains(c("DEPRESSED", "TROUBLE_SLEEPING"))
-  ) %>%
-  tab_row_group(
-    label = "Lifestyle Factors",
-    rows = contains(c("EXERCISE", "SLEEP", "SOCIAL"))
-  ) %>%
-  # Style the table
-  tab_style(
-    style = cell_fill(color = "#e6f3ff"),
-    locations = cells_row_groups()
-  ) %>%
-  tab_style(
-    style = list(
-      cell_text(weight = "bold")
-    ),
-    locations = cells_row_groups()
-  ) %>%
-  # Add footnotes if needed
-  tab_footnote(
-    footnote = "Percentages may not sum to 100% due to rounding",
-    locations = cells_column_labels(columns = contains("%"))
-  ) %>%
-  # Adjust column widths
-  cols_width(
-    Variable ~ px(250),
-    everything() ~ px(120)
-  )
-
-tab1_gt
-```
-
 @tbl-summarytable shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-I generally recommend the `here` package.
 

### 2025-04-20 products/manuscript/Manuscript.qmd
Part 5 manuscript

@@
 
-Variables of interest contained significant amounts of missing information, so multiple imputation was used, creating 5 sets of imputed data that was averaged across the remainder of analysis to produce variables and statistics needed.
-Complex survey design was used to allow for generalizability of results, using vairbales PSU, STRATA, and SAMPWIEGHT.
+Binary variables were created to allow for interpretability.
+EXERCISE is a combination of number of minutes of passive or aerboic exercise, descirbing whther weekly amount meets the American Heart Association's guidelines for exercise needed to stay healthy, as exercise is known to relate to mental health.
+SLEEP is based on the NIH guidelines for recommended hours of sleep per day, which differs by age.
 
-Binary variables were created to allow for interpretability.
-
-~~~
-
-Survey design variables were used and new binary variables were created from other variables. Using complex survey design variables, a descriptive table was created.
-
-I will create models (basic and adjusted) to study association (generating prevelance ratios with CIs) between mental health factors (hours of sleep, feeling depressed mood, amount of social and emotional support, trouble sleeping, recent exercise benchmarks) with a postiive covid test result.
-I will generate tables and plots for all results.
+Complex survey design was used to allow for generalizability of results, using vairbales PSU, STRATA, and SAMPWEIGHT.
 
@@
 
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
-
 ```{r}
 #| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
+#| fig-cap: "Flowchart depicting methodology"
 #| echo: FALSE
-#knitr::include_graphics(here("assets","antigen-recognition.png"))
+knitr::include_graphics(here("results","figures", "flowchart.png"))
 ```

### 2025-04-20 products/manuscript/Manuscript.qmd
Part 5 manuscript updates

@@
 SLEEP is based on the NIH guidelines for recommended hours of sleep per day, which differs by age.
+DEPRESSED includes if the individual reports feeling depressed on a weekly or daily basis, but not at a frequency of monthly or less.
+TROUBLE_SLEEPING indicates struggling to fall asleep several days per week, more than half the days, or nearly every day.
+SOCIAL describes individual feeling like they have less social support than their regular amount.
+Answers for don't know or unsure were converted to missing.
 
-Complex survey design was used to allow for generalizability of results, using vairbales PSU, STRATA, and SAMPWEIGHT.
+After creating binary variables, significant amounts of data were missing.
+Multiple imputation was used to create 5 sets of imputed data, and used for the remainder of analysis.
+To calculate later statisitcs, processes were preformed on each of the 5 sets and then averaged.
+
+Complex survey design was used to allow for generalizability of results, using variables PSU, STRATA, and SAMPWEIGHT.
+A summary table was generated to find occurence of each of the factors and their inverses among the data.
+Next, single models were ran to find associations between each variable with covid test postivity.
+To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.
 

### 2025-04-20 products/manuscript/Manuscript.qmd
Part 5 manuscript updates

@@
 
-Data was subsetted to years 2019-2022, to cover the brunt of pandemic years.
-Despite low prevelance of COVID-19 cases in 2019 worldwide, the appearance of COVID-19 in the news (and consequential potential impact on mental health factors) led us to include this data.
-The final dataset contains 151406 observations.
-
-Binary variables were created to allow for interpretability.
-EXERCISE is a combination of number of minutes of passive or aerboic exercise, descirbing whther weekly amount meets the American Heart Association's guidelines for exercise needed to stay healthy, as exercise is known to relate to mental health.
-SLEEP is based on the NIH guidelines for recommended hours of sleep per day, which differs by age.
-DEPRESSED includes if the individual reports feeling depressed on a weekly or daily basis, but not at a frequency of monthly or less.
-TROUBLE_SLEEPING indicates struggling to fall asleep several days per week, more than half the days, or nearly every day.
-SOCIAL describes individual feeling like they have less social support than their regular amount.
-Answers for don't know or unsure were converted to missing.
-
-After creating binary variables, significant amounts of data were missing.
-Multiple imputation was used to create 5 sets of imputed data, and used for the remainder of analysis.
-To calculate later statisitcs, processes were preformed on each of the 5 sets and then averaged.
-
-Complex survey design was used to allow for generalizability of results, using variables PSU, STRATA, and SAMPWEIGHT.
-A summary table was generated to find occurence of each of the factors and their inverses among the data.
-Next, single models were ran to find associations between each variable with covid test postivity.
-To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.
-
 ## Schematic of workflow
@@
 
-This data was obtained from IPUMS, used with permission.
+National Health Information Survey data was obtained from IPUMS, used with permission.
+Data was subsetted to years 2019-2022, to cover the brunt of pandemic years.
+Despite low prevelance of COVID-19 cases in 2019 worldwide, the appearance of COVID-19 in the news (and consequential potential impact on mental health factors) led us to include this data.
+The final dataset contains 151406 observations.
 
@@
 
-The following code details the data cleaning process.
-Labels are removed from the data to make it easier to work with to create models later in the process.
-Variables needed are selected, then a series of binary varibales are created.
-Additional details are available in the processing folder.
+Binary variables were created to allow for interpretability.
+EXERCISE is a combination of number of minutes of passive or aerboic exercise, descirbing whther weekly amount meets the American Heart Association's guidelines for exercise needed to stay healthy, as exercise is known to relate to mental health.
+SLEEP is based on the NIH guidelines for recommended hours of sleep per day, which differs by age.
+DEPRESSED includes if the individual reports feeling depressed on a weekly or daily basis, but not at a frequency of monthly or less.
+TROUBLE_SLEEPING indicates struggling to fall asleep several days per week, more than half the days, or nearly every day.
+SOCIAL describes individual feeling like they have less social support than their regular amount.
+Answers for don't know or unsure were converted to missing.
 
 ## Statistical analysis
-_Explain anything related to your statistical analyses._
 
-We will begin with exploratory tables to explore proportions of NA answers, and generate early estimates.
+After creating binary variables, significant amounts of data were missing.
+Multiple imputation was used to create 5 sets of imputed data, and used for the remainder of analysis.
+To calculate later statisitcs, processes were preformed on each of the 5 sets and then averaged.
 
-Our first figure with be a discriptive table stratified by COVID status, with estimates and 95% CIs.
-
-Next, we will construct simple models before moving onto adjusted models.
+Complex survey design was used to allow for generalizability of results, using variables PSU, STRATA, and SAMPWEIGHT.
+A summary table was generated to find occurence of each of the factors and their inverses among the data.
+Next, single models were ran to find associations between each variable with covid test postivity.
+To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.
 
@@
 
-The first table below details the proportion of NAs for each mental health variable.
-We can see this is a little bit high, which may affect analysis later on.
-Histograms of each variable are available in the eda folder.
-
-The second table shows proportions calculated using complex survey design variables.
-
-@tbl-summarytable shows a summary of the data.
+```{r}
+#| label: fig-schematic
+#| fig-cap: "Summary statistics table"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures", "table1_imputed.png"))
+```
 

### 2025-04-20 code/analysis-code/analysis.qmd
Part 5 manuscript updates

@@
 
-# Fit svyglm models for EXERCISE predictor
-models_exercise <- lapply(survey_designs, function(dsgn) {
-  svyglm(COVID ~ EXERCISE, design = dsgn, family = quasibinomial())
-})
-
-# Extract coefficients and standard errors for each model
-model_exercise_coef <- lapply(models_exercise, function(model) {
-  data.frame(
-    Estimate = coef(model),
-    Std.Error = sqrt(diag(vcov(model))),
-    stringsAsFactors = FALSE
-  )
-})
-
-# Calculate means and variances for pooling
-estimates <- do.call(rbind, lapply(model_exercise_coef, function(x) x$Estimate))
-se <- do.call(rbind, lapply(model_exercise_coef, function(x) x$Std.Error))
-
-# Compute the mean across imputations
-mean_estimate <- apply(estimates, 2, mean)
-
-# Compute within-imputation variance (average variance across imputations)
-within_variance <- apply(se^2, 2, mean)
-
-# Compute between-imputation variance (variance of the estimates across imputations)
-between_variance <- apply(estimates, 2, var)
-
-# Calculate the total variance using Rubin's rule
-total_variance <- within_variance + (1 + 1 / length(survey_designs)) * between_variance
-
-# Calculate the pooled standard error
-pooled_se <- sqrt(total_variance)
-
-# Calculate the 95% confidence intervals
-ci_low <- mean_estimate - 1.96 * pooled_se
-ci_high <- mean_estimate + 1.96 * pooled_se
-
-# Create a data frame with pooled estimates, standard errors, and confidence intervals
-pooled_exercise <- data.frame(
-  Estimate = mean_estimate,
-  Std.Error = pooled_se,
-  CI_Low = ci_low,
-  CI_High = ci_high
-)
-
-# Display the pooled results
-pooled_exercise
-
-library(survey)
-
 # Fit svyglm models for EXERCISE predictor (and other predictors if needed)
@@
 
-Now we will make a nice table of that using gt.
+Now we will make a nice table of that using gt and save it to use in the mauscript.
 
@@
 #not including intercepts values
-# Assuming 'pooled_exercise' is your data frame
-pooled_exercise %>%
+table2 <- pooled_exercise %>%
   group_by(Predictor) %>%
@@
   )
+
+# Save the table as a PNG
+here()
+gtsave(table2, here("results", "figures", "single_model_table.png"))
 

### 2025-04-20 products/manuscript/Manuscript.qmd
Part 5 manuscript updates

@@
 ```{r}
-#| label: fig-schematic
+#| label: fig-1
 #| fig-cap: "Summary statistics table"
@@
 
+## Basic statistical analysis
+
 ```{r}
 #| label: tbl-summarytable
-#| tbl-cap: "Table. 1. Mental health characteristics among Americans from 2019-2022, by SARS-CoV-2 test result status, NHIS"
+#| tbl-cap: "Single model table"
 #| echo: FALSE
 here()
-knitr::include_graphics(here("results","figures", "result_table1.png"))
-```
-
-## Basic statistical analysis
-
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-fig-result shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-#knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
+knitr::include_graphics(here("results","figures", "single_model_table.png"))
 ```

### 2025-04-20 products/manuscript/Manuscript.qmd
Part 5 manuscript updates

@@
 ---
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
 
@@
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Example tbl-resulttable2 shows a summary of a linear model fit.
-
 {{< pagebreak >}}
@@
 
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like.
-
 {{< pagebreak >}}

### 2025-04-21 project-review-rayleen-lewis.qmd
Project review - Rayleen Lewis

---
title: Project Review  
author: Rayleen Lewis
date: April 20, 2025
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Mental Health Factors and Covid

Name of project author(s): Annalise Cramer

Name of project reviewer: Rayleen Lewis

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

Some motivation is provided in the abstract and intro, but more is needed. References should also be added.

### Summary assessment 
* some contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The questions are explicitly stated. I don't have any concerns about the data source.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

Some additional details about the original variables in NHIS and how they were dichotomized (e.g., what were the guidelines for sleep, was the original NHIS question, "How many hours of sleep do you get a night?"). NHIS is performed by CDC and this should be noted along with access through IPUMS. A reference to the data should also be added.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

Great graphic on the data wrangling work flow! Adding descriptions to the manuscript of the level of missingness and how much data were imputed would be helpful (I did see missingness in the EDA file which was nice). Because COVID questions weren't added until July 2020, everyone from 2019 and the first half of 2020 would be missing COVID information by design; these people shouldn't be included in the analysis of objective 2 regarding COVID.

### Summary assessment
* some weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Because the very first case of COVID wasn't seen until December 2019, I don't think this year of data should be included in the analysis. Also, COVID information wasn't included in NHIS until July 2020 according to the [NHIS documentation](https://ftp.cdc.gov/pub/Health_Statistics/NCHS/Dataset_Documentation/NHIS/2020/srvydesc-508.pdf), so these people also shouldn't be included, at least for your second objective. I was curious how they handled this with the weighting, and it looks like recommendations on how to deal with this are provided at the link above.

For Figure 2, it might be more informative to have row %'s (percent of people with sleep = 1 who were positive for COVID), but this is personal preference.

In your EDA, it looked like the level of missingness was really high for some variables (e.g., >90% for sleep). Was there a systematic reason the level of missingness was so high? Even with multiple imputation, I'm not sure if this should be used. Details on the imputation model should also be provided (i.e., were the same variables in the regression model used for the imputation model?).

I've never used CV on complex survey data (so I'm not 100% sure about this), but I don't think CV is appropriate for this survey data because it "drops" people in each run, which would mess up the weighting and survey design.

For the LASSO model, I think there might be typo in Line 388; it looks like depressed is defined as the outcome, but I think this is supposed to be COVID.

### Summary assessment
* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Included tables are nicely formatted.

### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

This section is missing.

### Summary assessment
* major parts of discussion missing or wrong

## Further comments

Good start, it seems extra tricky to account for the survey design with some of the techniques and models we've learned in class!

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Yes, everything was intuitive and easy to find.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Good documentation, and I appreciated that code generated using AI was noted and the AI product was listed.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments
Great ReadMe files in all folders. The readme in the processing-code folder has an incorrect file name (refers to data-cleaning instead of processing), but this wasn't a big deal because there's only one qmd file in that folder, so it was esay to figure out the correct file to use instead.

I got a couple of minor errors that caused some problems when running the code:
eda.qmd:

- CreateTableOne was loaded as a package, but I think this is a function within the tableone package. I got an error when running the code.

analysis.qmd:

- The package glmnet needed to be loaded (i.e., need to add library(glmnet) to file)
- Lines 1200-1204: These lines don't run, I get this error: Error in fit$rank : object of type 'closure' is not subsettable

Manuscript rendered without issues.

### Summary assessment
* small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Analysis has been performed using multiple models with RMSE output for each model. This hasn't been included in the manuscript yet. The questions haven't been fully addressed given the missing analysis pieces and discussion in the manuscript. There are still some sections of the template in the manuscript. References will also need to be added.

### Summary assessment
* weak level of thoroughness

## Further comments

Nothing further.

### 2025-04-30 code/analysis-code/analysis.qmd
Part 6 updates

@@
 library(gt)
+library(pROC)
+library(ranger)
+library(glmnet)
+library(Metrics)
+library(purrr)
 library(here)
@@
 
-Next, we will make the multivariate model. I will use cross validation for this. The RMSE is found to be 0.4450602.
+Next, we will make the multivariate model. I will use cross validation for this. The pooled AUC across imputations is found to be 0.5230.
 ```{r}
-
-cross_validation_imputed_rmse <- function(imputed_data, design, predictors, k_folds = 5) {
-  rmse_list <- numeric(length(imputed_data))  # Store RMSE for each imputation
+# Define the AUC cross-validation function for imputed datasets
+cross_validation_imputed_auc <- function(imputed_data, predictors, k_folds = 5) {
+  auc_list <- numeric(length(imputed_data))  # Store AUC for each imputation
 
@@
     data <- imputed_data[[i]]
+
+    # drop rows with missing values in COVID, predictors, or SAMPWEIGHT
+    data <- data %>% drop_na(all_of(c("COVID", predictors, "SAMPWEIGHT")))
+
     set.seed(123)
@@
 
-    fold_rmse <- numeric(k_folds)  # RMSE for each fold
+    fold_auc <- numeric(k_folds)  # AUC for each fold
 
@@
 
-      # Survey design for training and testing
-      train_design <- svydesign(ids = ~1, data = train_data, weights = ~SAMPWEIGHT)
+      # Fit random forest model on training data
       formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
-      model <- svyglm(formula, design = train_design, family = quasibinomial())
+      rf_model <- ranger(
+        formula = formula,
+        data = train_data,
+        case.weights = train_data$SAMPWEIGHT,
+        probability = TRUE
+      )
 
-      # Predictions
-      predictions <- predict(model, newdata = test_data, type = "response")
+      # Predict probabilities for class 1 (COVID = 1)
+      pred_probs <- predict(rf_model, data = test_data)$predictions[, 2]
       actual <- test_data$COVID
 
-      # Compute RMSE for this fold
-      fold_rmse[fold] <- sqrt(mean((predictions - actual)^2, na.rm = TRUE))
+      # Compute AUC for this fold
+      auc_result <- roc(actual, pred_probs)$auc
+      fold_auc[fold] <- auc_result
     }
 
-    # Average RMSE over folds for this imputed dataset
-    rmse_list[i] <- mean(fold_rmse)
+    # Average AUC over folds for this imputed dataset
+    auc_list[i] <- mean(fold_auc)
   }
 
-  # Average RMSE over all imputed datasets
-  overall_rmse <- mean(rmse_list)
+  # Average AUC over all imputed datasets
+  overall_auc <- mean(auc_list)
 
-  return(list(rmse_per_imputation = rmse_list, overall_rmse = overall_rmse))
+  return(list(auc_per_imputation = auc_list, overall_auc = overall_auc))
 }
@@
 # Assuming you have your imputed_data and a vector of predictors
-results <- cross_validation_imputed_rmse(imputed_data = data_imputed,
-                                         design = NULL,  # Not used inside, can be omitted
-                                         predictors = c("AGE", "SLEEP", "DEPRESSED"))
+results <- cross_validation_imputed_auc(imputed_data = data_imputed,
+                                        predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL"))
 
 # View results
-results$rmse_per_imputation  # RMSE for each imputed dataset
-results$overall_rmse         # Average RMSE across imputations
+results$auc_per_imputation  # AUC for each imputed dataset
+results$overall_auc         # Average AUC across imputations
 ```
 
-Now, we will make a LASSO regression model. For this, I'm not going to use the multiple imputation dataset. I'm just going to pick a sinlge imputed dataset, the first one. The RMSE is 0.276.
+Now, we will make a LASSO regression model.
+For this, I'm going to use the multiple imputation dataset with cross validation with 5 folds.
+Instead of using complex survey design, I will weight the model by sampleweight.
+The pooled AUC is  0.529.
 
@@
 # Extract the first completed imputed dataset
-data1 <- data_imputed[[1]]  # You can loop over 1:5 later
+data1 <- data_imputed[[1]]
 
-# Then define the survey design
+# Define the survey design
 survey_design <- svydesign(
@@
 
-# Choose your outcome and predictors
-outcome_var <- "DEPRESSED"
+# Choose outcome and predictors
+outcome_var <- "COVID"
 predictor_vars <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
 
-# Loop over imputations and compute RMSE
-rmse_list <- map(data_imputed, function(dat) {
+# Loop over imputations and compute AUC
+auc_list <- map(data_imputed, function(dat) {
   # Drop missing in predictors or outcome if any slipped through
@@
 
-  # Fit weighted lasso model
-  cv_fit <- cv.glmnet(
-    x = X_train,
-    y = y_train,
-    alpha = 1,
-    weights = w_train,
-    family = "binomial",  # change to "binomial" for binary outcome
-    nfolds = 5
+  # Fit random forest model (instead of lasso, because it's a classification problem)
+  rf_model <- ranger(
+    formula = as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))),
+    data = dat[train_idx, ],
+    case.weights = w_train,
+    probability = TRUE
   )
 
-  # Predict on test set
-  preds <- predict(cv_fit, newx = X_test, s = "lambda.min")
+  # Predict probabilities for class 1
+  pred_probs <- predict(rf_model, data = dat[test_idx, ])$predictions[,2]
 
-  # Weighted RMSE
-  sqrt(weighted.mean((y_test - preds)^2, w_test))
+  # Compute AUC
+  auc_result <- roc(y_test, pred_probs)$auc
+  return(auc_result)
 })
 
-# Pool RMSE across imputations
-pooled_rmse <- mean(unlist(rmse_list))
+# Pool AUC across imputations
+mean_auc <- mean(unlist(auc_list))
 
 # Output
-cat("RMSE for each imputation:\n")
-print(unlist(rmse_list))
-cat("\nPooled RMSE across imputations:", round(pooled_rmse, 3), "\n")
-
+cat("AUC for each imputation:\n")
+print(unlist(auc_list))
+cat("\nPooled AUC across imputations:", round(mean_auc, 3), "\n")
 ```
 
-################################################################################
-The following code was used prior to development of imputed data and is no longer relevant.
-It may be ignored.
-I am keeping it in case it serves as a useful reference for the rest of the project or future endeavors.
-################################################################################
+Finally, we will construct a random forest model.
+For this, I will not use the multiple imputation dataset, as before, I will use the first imputed dataset, with cross validation 5 folds.
+The pooled AUC is 0.556.
 
-We create estimates and CIs using complex survey design variables
 ```{r}
-#list variables
-varstable1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
-             "SOCIAL", "COVID")
+library(ranger)
+library(caret)
+library(pROC)
+library(dplyr)
 
-#list categories
-cat_vartab1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
-             "SOCIAL", "COVID")
+set.seed(1234)
+k <- 5
+auc_list <- numeric(length(data_imputed))
 
-survey = svydesign(
-  ids = ~PSU, 
-  strata = ~STRATA, 
-  weights = ~SAMPWEIGHT, 
-  nest = TRUE, 
-  data = data)
+for (i in seq_along(data_imputed)) {
+  data1 <- data_imputed[[i]]
 
-exercise <- svyby(~as.factor(EXERCISE), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
-sleep <- svyby(~as.factor(SLEEP), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
-depressed <- svyby(~as.factor(DEPRESSED), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
-trouble_sleeping <- svyby(~as.factor(TROUBLE_SLEEPING), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
-social <- svyby(~as.factor(SOCIAL), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
-str(exercise)
+  # Remove missing outcome
+  data1_complete <- data1 %>% filter(!is.na(COVID))
+
+  # Make sure outcome is binary (0/1)
+  data1_complete$COVID <- as.numeric(data1_complete$COVID)
+
+  # Add weights column if not present
+  data1_complete$weights <- data1_complete$SAMPWEIGHT
+
+  # Create folds
+  folds <- createFolds(data1_complete$COVID, k = k, list = TRUE, returnTrain = FALSE)
+
+  auc_folds <- numeric(k)
+
+  for (fold_idx in seq_along(folds)) {
+    test_idx <- folds[[fold_idx]]
+    train_idx <- setdiff(seq_len(nrow(data1_complete)), test_idx)
+
+    train_data <- data1_complete[train_idx, ]
+    test_data  <- data1_complete[test_idx, ]
+
+    # Fit random forest classifier
+    rf_model <- ranger(
+      formula = COVID ~ EXERCISE + SLEEP + DEPRESSED + TROUBLE_SLEEPING + SOCIAL, #formula
+      data = train_data,
+      case.weights = train_data$weights,
+      probability = TRUE
+    )
+
+    # Predict probabilities for class 1
+    pred_probs <- predict(rf_model, data = test_data)$predictions[,2]
+
+    # Compute AUC
+    auc_folds[fold_idx] <- roc(test_data$COVID, pred_probs, quiet = TRUE)$auc
+  }
+
+  # Store average AUC for this imputation
+  auc_list[i] <- mean(auc_folds)
+}
+
+# Pool AUC across imputations
+mean_auc <- mean(auc_list)
+print(paste("Pooled AUC across imputations:", round(mean_auc, 3)))
 ```
 
-Let's rearrange these exercise numbers into a table. I used ChatGPT and Claude to help with making the table. 
+Let's put these into a nice table for the manuscript. 
 ```{r}
-# Convert the svyby object to a data frame (if not already)
-exercise_df <- as.data.frame(exercise)
-
-# Clean up column names to make them more readable
-colnames(exercise_df) <- c("COVID", "EXERCISE_0", "EXERCISE_1", "CI_L_EXERCISE_0", "CI_L_EXERCISE_1", "CI_U_EXERCISE_0", "CI_U_EXERCISE_1")
-
-# Use a different approach to reshape the data
-exercise_long <- exercise_df %>%
-  # First, let's gather all columns except COVID
-  pivot_longer(
-    cols = -COVID,
-    names_to = "variable",
-    values_to = "value"
-  ) %>%
-  # Now separate the variable column into type and exercise level
-  mutate(
-    Type = case_when(
-      grepl("^EXERCISE", variable) ~ "Mean",
-      grepl("^CI_L", variable) ~ "L",
-      grepl("^CI_U", variable) ~ "U"
-    ),
-    Exercise = case_when(
-      grepl("_0$", variable) ~ "0",
-      grepl("_1$", variable) ~ "1"
-    )
-  ) %>%
-  # Now pivot wider to get the structure we want
-  pivot_wider(
-    id_cols = c(COVID, Exercise),
-    names_from = Type,
-    values_from = value
+# Create a data frame with the results
+auc_table <- data.frame(
+  Model = c("Multivariate", "Lasso", "Random Forest"),
+  AUC = c(0.523, 0.529, 0.556)
 )
 
-# Combine the CI values into a single column
-exercise_with_ci <- exercise_long %>%
-  mutate(
-    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
-  ) %>%
-  select(COVID, Exercise, Mean, CI)
-
-# Now create the final table format
-exercise_final <- exercise_with_ci %>%
-  # Convert COVID to character to ensure it works as a column header
-  mutate(COVID = as.character(COVID)) %>%
-  # Pivot to get COVID status as columns
-  pivot_wider(
-    id_cols = Exercise,
-    names_from = COVID,
-    values_from = c("Mean", "CI"),
-    names_glue = "{.value}_{COVID}"
-  )
-
-# Create the table using gt
-exercise_table <- exercise_final %>%
+# Create and style the GT table
+model_table <- auc_table %>%
   gt() %>%
   tab_header(
-    title = "Exercise Data by COVID Status"
+    title = "Model Performance Comparison",
   ) %>%
-  cols_label(
-    Exercise = "Exercise Level",
-    Mean_0 = "Mean",
-    Mean_1 = "Mean",
-    CI_0 = "95% CI",
-    CI_1 = "95% CI"
-  ) %>%
-  tab_spanner(
-    label = "COVID Negative (0)",
-    columns = c(Mean_0, CI_0)
-  ) %>%
-  tab_spanner(
-    label = "COVID Positive (1)",
-    columns = c(Mean_1, CI_1)
-  ) %>%
-  # Add formatting for the mean values
   fmt_number(
-    columns = starts_with("Mean"),
-    decimals = 2
-  ) %>%
-  # Add some style
-  tab_style(
-    style = cell_fill(color = "#f5f5f5"),
-    locations = cells_body(rows = Exercise == "0")
-  ) %>%
-  # Add a footnote
-  tab_footnote(
-    footnote = "Exercise Level: 0 = No regular exercise, 1 = Meets requirements for regular exercise according to American Heart Association",
-    locations = cells_column_labels(columns = Exercise)
-  ) %>%
-  # Add meaningful row labels
-  text_transform(
-    locations = cells_body(columns = Exercise),
-    fn = function(x) {
-      ifelse(x == "0", "No Regular Exercise", "Regular Exercise")
-    }
+    columns = vars(AUC),
+    decimals = 3
   )
@@
 # Print the table
-exercise_table
-```
-
-Next, we will add the sleep variable. I used ChatGPT for this.
-```{r}
-# Convert the svyby object to a data frame (if not already)
-sleep_df <- as.data.frame(sleep)  # Assuming 'sleep' is the dataset
-
-# Clean up column names to make them more readable
-colnames(sleep_df) <- c("COVID", "SLEEP_0", "SLEEP_1", "CI_L_SLEEP_0", "CI_L_SLEEP_1", "CI_U_SLEEP_0", "CI_U_SLEEP_1")
-
-# Use a different approach to reshape the data
-sleep_long <- sleep_df %>%
-  # First, let's gather all columns except COVID
-  pivot_longer(
-    cols = -COVID,
-    names_to = "variable",
-    values_to = "value"
-  ) %>%
-  # Now separate the variable column into type and sleep level
-  mutate(
-    Type = case_when(
-      grepl("^SLEEP", variable) ~ "Mean",
-      grepl("^CI_L", variable) ~ "L",
-      grepl("^CI_U", variable) ~ "U"
-    ),
-    Sleep = case_when(
-      grepl("_0$", variable) ~ "0",
-      grepl("_1$", variable) ~ "1"
-    )
-  ) %>%
-  # Now pivot wider to get the structure we want
-  pivot_wider(
-    id_cols = c(COVID, Sleep),
-    names_from = Type,
-    values_from = value
-  )
-
-# Combine the CI values into a single column
-sleep_with_ci <- sleep_long %>%
-  mutate(
-    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
-  ) %>%
-  select(COVID, Sleep, Mean, CI)
-
-# Now create the final table format
-sleep_final <- sleep_with_ci %>%
-  # Convert COVID to character to ensure it works as a column header
-  mutate(COVID = as.character(COVID)) %>%
-  # Pivot to get COVID status as columns
-  pivot_wider(
-    id_cols = Sleep,
-    names_from = COVID,
-    values_from = c("Mean", "CI"),
-    names_glue = "{.value}_{COVID}"
-  )
-
-# Create the table using gt
-sleep_table <- sleep_final %>%
-  gt() %>%
-  tab_header(
-    title = "Sleep Data by COVID Status"
-  ) %>%
-  cols_label(
-    Sleep = "Sleep Level",
-    Mean_0 = "Mean",
-    Mean_1 = "Mean",
-    CI_0 = "95% CI",
-    CI_1 = "95% CI"
-  ) %>%
-  tab_spanner(
-    label = "COVID Negative (0)",
-    columns = c(Mean_0, CI_0)
-  ) %>%
-  tab_spanner(
-    label = "COVID Positive (1)",
-    columns = c(Mean_1, CI_1)
-  ) %>%
-  # Add formatting for the mean values
-  fmt_number(
-    columns = starts_with("Mean"),
-    decimals = 2
-  ) %>%
-  # Add some style
-  tab_style(
-    style = cell_fill(color = "#f5f5f5"),
-    locations = cells_body(rows = Sleep == "0")
-  ) %>%
-  # Add a footnote
-  tab_footnote(
-    footnote = "Sleep Level: 0 = Less than recommended sleep, 1 = Met recommendations for sleep by age groups according to NIH",
-    locations = cells_column_labels(columns = Sleep)
-  ) %>%
-  # Add meaningful row labels
-  text_transform(
-    locations = cells_body(columns = Sleep),
-    fn = function(x) {
-      ifelse(x == "0", "Not enough sleep", "Enough sleep")
-    }
-  )
-
-# Print the table
-sleep_table
-
-#################################################################################
-
-# Extract the data from the gt tables
-exercise_table_df <- as.data.frame(exercise_table)
-sleep_table_df <- as.data.frame(sleep_table)
-
-# Ensure that the column names are the same across both tables
-colnames(sleep_table_df) <- colnames(exercise_table_df)
-
-# Combine the tables row by row
-combined_df <- bind_rows(exercise_table_df, sleep_table_df)
-
-# Convert the combined data frame back to a gt table
-combined_table <- combined_df %>%
-  gt() %>%
-  tab_header(
-    title = "Combined Exercise and Sleep Data by COVID Status"
-  )
-
-# Print the combined gt table
-combined_table
-```
-
-Create the same tables with depressed, trouble_sleeping, and social. This was completed in BBEdit.
-```{r}
-#########################################################################################
-# depressed variable
-
-# Convert the svyby object to a data frame (if not already)
-depressed_df <- as.data.frame(depressed)
-
-# Clean up column names to make them more readable
-colnames(depressed_df) <- c("COVID", "depressed_0", "depressed_1", "CI_L_depressed_0", "CI_L_depressed_1", "CI_U_depressed_0", "CI_U_depressed_1")
-
-# Use a different approach to reshape the data
-depressed_long <- depressed_df %>%
-  # First, let's gather all columns except COVID
-  pivot_longer(
-    cols = -COVID,
-    names_to = "variable",
-    values_to = "value"
-  ) %>%
-  # Now separate the variable column into type and depressed level
-  mutate(
-    Type = case_when(
-      grepl("^depressed", variable) ~ "Mean",
-      grepl("^CI_L", variable) ~ "L",
-      grepl("^CI_U", variable) ~ "U"
-    ),
-    depressed = case_when(
-      grepl("_0$", variable) ~ "0",
-      grepl("_1$", variable) ~ "1"
-    )
-  ) %>%
-  # Now pivot wider to get the structure we want
-  pivot_wider(
-    id_cols = c(COVID, depressed),
-    names_from = Type,
-    values_from = value
-  )
-
-# Combine the CI values into a single column
-depressed_with_ci <- depressed_long %>%
-  mutate(
-    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
-  ) %>%
-  select(COVID, depressed, Mean, CI)
-
-# Now create the final table format
-depressed_final <- depressed_with_ci %>%
-  # Convert COVID to character to ensure it works as a column header
-  mutate(COVID = as.character(COVID)) %>%
-  # Pivot to get COVID status as columns
-  pivot_wider(
-    id_cols = depressed,
-    names_from = COVID,
-    values_from = c("Mean", "CI"),
-    names_glue = "{.value}_{COVID}"
-  )
-
-# Create the table using gt
-depressed_table <- depressed_final %>%
-  gt() %>%
-  tab_header(
-    title = "depressed Data by COVID Status"
-  ) %>%
-  cols_label(
-    depressed = "depressed Level",
-    Mean_0 = "Mean",
-    Mean_1 = "Mean",
-    CI_0 = "95% CI",
-    CI_1 = "95% CI"
-  ) %>%
-  tab_spanner(
-    label = "COVID Negative (0)",
-    columns = c(Mean_0, CI_0)
-  ) %>%
-  tab_spanner(
-    label = "COVID Positive (1)",
-    columns = c(Mean_1, CI_1)
-  ) %>%
-  # Add formatting for the mean values
-  fmt_number(
-    columns = starts_with("Mean"),
-    decimals = 2
-  ) %>%
-  # Add some style
-  tab_style(
-    style = cell_fill(color = "#f5f5f5"),
-    locations = cells_body(rows = depressed == "0")
-  ) %>%
-  # Add a footnote
-  tab_footnote(
-    footnote = "Depressed level: 0 = Feeling depressed less than weekly, 1 = Feeling depressed weekly or more",
-    locations = cells_column_labels(columns = depressed)
-  ) %>%
-  # Add meaningful row labels
-  text_transform(
-    locations = cells_body(columns = depressed),
-    fn = function(x) {
-      ifelse(x == "0", "Not regularly depressed", "Frequent depressed feeling")
-    }
-  )
-
-# Print the table
-depressed_table
-
-#########################################################################################
-# trouble_sleeping variable
-
-# Convert the svyby object to a data frame (if not already)
-trouble_sleeping_df <- as.data.frame(trouble_sleeping)
-
-# Clean up column names to make them more readable
-colnames(trouble_sleeping_df) <- c("COVID", "trouble_sleeping_0", "trouble_sleeping_1", "CI_L_trouble_sleeping_0", "CI_L_trouble_sleeping_1", "CI_U_trouble_sleeping_0", "CI_U_trouble_sleeping_1")
-
-# Use a different approach to reshape the data
-trouble_sleeping_long <- trouble_sleeping_df %>%
-  # First, let's gather all columns except COVID
-  pivot_longer(
-    cols = -COVID,
-    names_to = "variable",
-    values_to = "value"
-  ) %>%
-  # Now separate the variable column into type and trouble_sleeping level
-  mutate(
-    Type = case_when(
-      grepl("^trouble_sleeping", variable) ~ "Mean",
-      grepl("^CI_L", variable) ~ "L",
-      grepl("^CI_U", variable) ~ "U"
-    ),
-    trouble_sleeping = case_when(
-      grepl("_0$", variable) ~ "0",
-      grepl("_1$", variable) ~ "1"
-    )
-  ) %>%
-  # Now pivot wider to get the structure we want
-  pivot_wider(
-    id_cols = c(COVID, trouble_sleeping),
-    names_from = Type,
-    values_from = value
-  )
-
-# Combine the CI values into a single column
-trouble_sleeping_with_ci <- trouble_sleeping_long %>%
-  mutate(
-    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
-  ) %>%
-  select(COVID, trouble_sleeping, Mean, CI)
-
-# Now create the final table format
-trouble_sleeping_final <- trouble_sleeping_with_ci %>%
-  # Convert COVID to character to ensure it works as a column header
-  mutate(COVID = as.character(COVID)) %>%
-  # Pivot to get COVID status as columns
-  pivot_wider(
-    id_cols = trouble_sleeping,
-    names_from = COVID,
-    values_from = c("Mean", "CI"),
-    names_glue = "{.value}_{COVID}"
-  )
-
-# Create the table using gt
-trouble_sleeping_table <- trouble_sleeping_final %>%
-  gt() %>%
-  tab_header(
-    title = "trouble_sleeping Data by COVID Status"
-  ) %>%
-  cols_label(
-    trouble_sleeping = "trouble_sleeping Level",
-    Mean_0 = "Mean",
-    Mean_1 = "Mean",
-    CI_0 = "95% CI",
-    CI_1 = "95% CI"
-  ) %>%
-  tab_spanner(
-    label = "COVID Negative (0)",
-    columns = c(Mean_0, CI_0)
-  ) %>%
-  tab_spanner(
-    label = "COVID Positive (1)",
-    columns = c(Mean_1, CI_1)
-  ) %>%
-  # Add formatting for the mean values
-  fmt_number(
-    columns = starts_with("Mean"),
-    decimals = 2
-  ) %>%
-  # Add some style
-  tab_style(
-    style = cell_fill(color = "#f5f5f5"),
-    locations = cells_body(rows = trouble_sleeping == "0")
-  ) %>%
-  # Add a footnote
-  tab_footnote(
-    footnote = "trouble_sleeping level: 0 = No trouble sleeping, 1 = Reported trouble sleeping",
-    locations = cells_column_labels(columns = trouble_sleeping)
-  ) %>%
-  # Add meaningful row labels
-  text_transform(
-    locations = cells_body(columns = trouble_sleeping),
-    fn = function(x) {
-      ifelse(x == "0", "No reported trouble sleeping", "Regular trouble sleeping")
-    }
-  )
-
-# Print the table
-trouble_sleeping_table
-
-#########################################################################################
-# social variable
-
-# Convert the svyby object to a data frame (if not already)
-social_df <- as.data.frame(social)
-
-# Clean up column names to make them more readable
-colnames(social_df) <- c("COVID", "social_0", "social_1", "CI_L_social_0", "CI_L_social_1", "CI_U_social_0", "CI_U_social_1")
-
-# Use a different approach to reshape the data
-social_long <- social_df %>%
-  # First, let's gather all columns except COVID
-  pivot_longer(
-    cols = -COVID,
-    names_to = "variable",
-    values_to = "value"
-  ) %>%
-  # Now separate the variable column into type and social level
-  mutate(
-    Type = case_when(
-      grepl("^social", variable) ~ "Mean",
-      grepl("^CI_L", variable) ~ "L",
-      grepl("^CI_U", variable) ~ "U"
-    ),
-    social = case_when(
-      grepl("_0$", variable) ~ "0",
-      grepl("_1$", variable) ~ "1"
-    )
-  ) %>%
-  # Now pivot wider to get the structure we want
-  pivot_wider(
-    id_cols = c(COVID, social),
-    names_from = Type,
-    values_from = value
-  )
-
-# Combine the CI values into a single column
-social_with_ci <- social_long %>%
-  mutate(
-    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
-  ) %>%
-  select(COVID, social, Mean, CI)
-
-# Now create the final table format
-social_final <- social_with_ci %>%
-  # Convert COVID to character to ensure it works as a column header
-  mutate(COVID = as.character(COVID)) %>%
-  # Pivot to get COVID status as columns
-  pivot_wider(
-    id_cols = social,
-    names_from = COVID,
-    values_from = c("Mean", "CI"),
-    names_glue = "{.value}_{COVID}"
-  )
-
-# Create the table using gt
-social_table <- social_final %>%
-  gt() %>%
-  tab_header(
-    title = "social Data by COVID Status"
-  ) %>%
-  cols_label(
-    social = "social Level",
-    Mean_0 = "Mean",
-    Mean_1 = "Mean",
-    CI_0 = "95% CI",
-    CI_1 = "95% CI"
-  ) %>%
-  tab_spanner(
-    label = "COVID Negative (0)",
-    columns = c(Mean_0, CI_0)
-  ) %>%
-  tab_spanner(
-    label = "COVID Positive (1)",
-    columns = c(Mean_1, CI_1)
-  ) %>%
-  # Add formatting for the mean values
-  fmt_number(
-    columns = starts_with("Mean"),
-    decimals = 2
-  ) %>%
-  # Add some style
-  tab_style(
-    style = cell_fill(color = "#f5f5f5"),
-    locations = cells_body(rows = social == "0")
-  ) %>%
-  # Add a footnote
-  tab_footnote(
-    footnote = "social Level: 0 = Regular social support, 1 = Less social support",
-    locations = cells_column_labels(columns = social)
-  ) %>%
-  # Add meaningful row labels
-  text_transform(
-    locations = cells_body(columns = social),
-    fn = function(x) {
-      ifelse(x == "0", "Regular social support", "Less social support")
-    }
-  )
-
-# Print the table
-social_table
-
-```
-
-Now, we will create the combined table of exercise, sleep, depressed, trouble_sleeping, and social. ChatGPT helped with many tweaks.
-```{r}
-# Extract the data from the other gt tables
-depressed_table_df <- as.data.frame(depressed_table)
-trouble_sleeping_table_df <- as.data.frame(trouble_sleeping_table)
-social_table_df <- as.data.frame(social_table)
-
-# Ensure the column names match across all tables
-colnames(depressed_table_df) <- colnames(combined_df)
-colnames(trouble_sleeping_table_df) <- colnames(combined_df)
-colnames(social_table_df) <- colnames(combined_df)
-
-# Combine all the tables row by row
-combined_all_df <- bind_rows(combined_df, depressed_table_df, trouble_sleeping_table_df, social_table_df)
-
-# Add a row_id column to the data frame
-combined_all_df <- combined_all_df %>%
-  mutate(row_id = row_number())
-
-# Convert the combined data frame back to a gt table
-combined_all_table <- combined_all_df %>%
-  gt() %>%
-  tab_header(
-    title = "Mental Health Factors by SARS-CoV-2 Test Status"
-  ) %>%
-  # Spanners for COVID Negative and COVID Positive
-  tab_spanner(
-    label = "Never has tested positive for Covid",
-    columns = starts_with("Estimate") | starts_with("CI 95%")
-  ) %>%
-  tab_spanner(
-    label = "Has tested positive for Covid",
-    columns = starts_with("Estimate") | starts_with("CI 95%")
-  ) %>%
-  # Remove the word 'Exercise' from the first row
-  cols_label(
-    Exercise = ""  # This removes the "Exercise" label
-  ) %>%
-  # Format the mean values
-  fmt_number(
-    columns = starts_with("Mean"),
-    decimals = 2
-  ) %>%
-  # Add a footnote
-  tab_footnote(
-    footnote = "Exercise requirements determined according to American Heart Association. Sleep levels determined by NIH recommendations by age group.",
-    locations = cells_column_labels(columns = Exercise)
-  ) %>%
-  # Add alternating row colors using row_id
-  tab_style(
-    style = cell_fill(color = "#f5f5f5"),  # Gray color for odd rows
-    locations = cells_body(rows = row_id %% 2 == 1)  # Apply to odd rows
-  ) %>%
-  tab_style(
-    style = cell_fill(color = "white"),  # White color for even rows
-    locations = cells_body(rows = row_id %% 2 == 0)  # Apply to even rows
-  ) %>%
-  # Add solid vertical line between row names and Mean_0 columns
-  tab_style(
-    style = cell_borders(
-      sides = c("left"),
-      color = "grey",
-      weight = px(1)
-    ),
-    locations = cells_body(
-      columns = starts_with("Mean_0")
-    )
-  ) %>%
-  # Add solid vertical line between COVID Negative and COVID Positive columns
-  tab_style(
-    style = cell_borders(
-      sides = c("left"),
-      color = "grey",
-      weight = px(1)
-    ),
-    locations = cells_body(
-      columns = starts_with("Mean_1")
-    )
-  ) %>%
-  # Hide the row_id column
-  cols_hide(columns = "row_id")
-
-# Print the combined gt table
-combined_all_table
-
-#footnote:
-#"Exercise requirements determined according to American Heart Association. Sleep levels determined by NIH recommendations by age group."
+model_table
 
@@
 here()
-gtsave(combined_all_table, here("results", "figures", "result_table1.png"))
+gtsave(model_table, here("results", "figures", "model_table.png"))
 ```
-
-################################################################################
-
-Up next, we will run the singleton models.
-For these, our outcome variable (covid test result) will be used as a variable for each of the respective predictor vairbales that we created in the processing code.
-These models appear as the following:
-
-$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Exercise} $$
-$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Sleep} $$
-$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Depressed} $$
-$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{TroubleSleeping} $$
-$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Social} $$
-Create survey design.
-```{r}
-#review names to use below
-names(data)
-
-#create survey object using design variables
-survey_design <- svydesign(
-  ids = ~PSU,
-  strata = ~STRATA,
-  weights = ~SAMPWEIGHT,
-  data = data,
-  nest = TRUE
-)
-
-pr_results <- data.frame()
-
-variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
-               "SOCIAL")
-
-for (var in variables) {
-  # Create the formula dynamically
-  formula <- as.formula(paste("COVID ~", var))
-
-  # Fit the model
-  model <- svyglm(
-    formula, 
-    design = survey_design, 
-    family = quasibinomial(link = "log")
-  )
-
-  # Extract prevalence ratios and confidence intervals
-  pr <- exp(coef(model))  # Prevalence ratios
-  pr_ci <- exp(confint(model))  # Confidence intervals
-
-  # Extract p-values
-  p_values <- summary(model)$coefficients[, "Pr(>|t|)"]
-
-  # Combine results into a single table
-  pr_results <- rbind(
-    pr_results,
-    data.frame(
-      Variable = var,
-      Estimate = pr[-1],  # Skip the intercept
-      `2.5% CI` = pr_ci[-1, 1],
-      `97.5% CI` = pr_ci[-1, 2],
-      `p-value` = p_values[-1]  # Skip the intercept
-    )
-  )
-}
-
-# Print the results
-print(pr_results)
-
-```
-
-Now, we will make that into a nice quality table using the gt package and the help of chatGPT and Claude.
-```{r}
-#checking
-glimpse(pr_results)
-str(pr_results)
-
-#create the gt table
-pr_results %>%
-  mutate(
-    Predictor = stringr::str_to_title(gsub("_", " ", Variable)),
-    `Prevalence Ratio` = round(Estimate, 2),
-    `95% CI` = paste0("(", round(X2.5..CI, 2), ", ", round(X97.5..CI, 2), ")"),
-    `p-value` = format.pval(p.value, digits = 3, eps = .001)
-  ) %>%
-  select(Predictor, `Prevalence Ratio`, `95% CI`, `p-value`) %>%
-  gt() %>%
-  tab_header(
-    title = md("**Positive COVID Test Result Association with Mental Health Factors**")
-  ) %>%
-  tab_options(
-    column_labels.border.bottom.width = px(1.5),  # Bold line under column headers
-    column_labels.border.bottom.color = "black"
-  )
-
-```
-
-Now, we're going to complete the process of model selection, using backwards selection. We want to end up with a model we can use for prediction purposes. Our initial model looks like this: 
-$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Exercise} + \beta_2 \cdot X_{Sleep} + \beta_3 \cdot X_{Depressed} + \beta_4 \cdot X_{TroubleSleeping} + \beta_5 \cdot X_{Social} $$
-Let's find the AIC value. 
-```{r}
-
-unique(data$COVID)
-model_full <- svyglm(
-  COVID ~ EXERCISE + SLEEP + DEPRESSED + TROUBLE_SLEEPING + SOCIAL,
-  design = survey_design,
-  family = quasibinomial(link = "log")
-)
-# Perform backward selection based on AIC
-model_step <- step(model_full, direction = "backward", trace = 1)
-summary(model_step)
-
-```

### 2025-04-30 code/processing-code/processing.qmd
Part 6 updates

@@
 
+Let us remove the data prior to 2020 because COVID was not present yet. Survey design variables are by year so it is not a problem to remove this.
+```{r}
+#filter for years that are not 2019
+data1 <- data1 %>%
+  filter(YEAR != 2019)
+
+#checking
+unique(data1$YEAR)
+```
 Selecting variables we need will make the dataset smaller and quicker to work with. Note it is renamed at each step for reproducibility. 
 ```{r}
-names(data_original) #view names of variables
-data1 <- data_original %>%
+names(data1) #view names of variables
+data1 <- data1 %>%
   select(STRATA, PSU, PERNUM, SAMPWEIGHT, AGE, PA18AERSTR, HRSLEEP, DEPFREQ, PHQSLEEP, CMPSUPPORT, CVDTESTRSLT)
@@
   mutate(
-    PA18AERSTR = ifelse(PA18AERSTR %in% c(0, 7, 8, 9), NA, PA18AERSTR), #put NA, 0, 6, 7, 8, 9 to NA
-    EXERCISE = if_else(PA18AERSTR %in% c(2,3,4), 1, 0, missing = NA_real_) #create EXERCISE variable, 1 if guidelines by Am Heart Assoc met for aerobics, strength, or both, 0 if not met
+    PA18AERSTR = ifelse(PA18AERSTR %in% c(0, 7, 8, 9), NA, PA18AERSTR), # put 0, 7, 8, 9 to NA
+    EXERCISE = if_else(PA18AERSTR %in% c(2, 3, 4), 1, 0, missing = NA_real_) # 1 if AHA guidelines met, else 0
   ) %>%
   mutate(
-    AGE = ifelse(AGE %in% c(997, 999), NA, AGE),  #invalid ages to NA
-    SLEEP = case_when( #Based on age, amount of sleep required from https://www.nhlbi.nih.gov/health/sleep/how-much-sleep#:~:text=Experts%20recommend%20that%20adults%20sleep,or%20more%20hours%20a%20night.
+    AGE = ifelse(AGE %in% c(997, 999), NA, AGE),  # invalid ages to NA
+    SLEEP = case_when( # Based on age, from NHLBI sleep guidelines
       AGE < 1                  ~ if_else(HRSLEEP >= 12 & HRSLEEP <= 16, 1, 0, missing = NA_real_),
@@
       AGE >= 13 & AGE <= 18    ~ if_else(HRSLEEP >= 8 & HRSLEEP <= 10, 1, 0, missing = NA_real_),
-      TRUE                     ~ NA_real_  #default NA for unspecified ages
-    )) %>%
+      AGE > 18                 ~ if_else(HRSLEEP >= 7 & HRSLEEP <= 9, 1, 0, missing = NA_real_),
+      TRUE                     ~ NA_real_  # default NA for unspecified or missing AGE
+    )
+  ) %>%
   mutate(
     DEPRESSED = case_when(
-      DEPFREQ %in% c(1, 2) ~ 1, #daily and weekly is considered depressed
-      DEPFREQ %in% c(3, 4, 5)    ~ 0, #monthly and a few times a year and never are considered not depressed
-      DEPFREQ %in% c(0, 7, 8, 9) ~ NA_real_,#NIU & Unknowns are NA
-      TRUE ~ NA_real_ #default NA for unexpected values
-    )) %>%
+      DEPFREQ %in% c(1, 2) ~ 1, # daily and weekly → depressed
+      DEPFREQ %in% c(3, 4, 5) ~ 0, # monthly or less → not depressed
+      DEPFREQ %in% c(0, 7, 8, 9) ~ NA_real_,
+      TRUE ~ NA_real_
+    )
+  ) %>%
   mutate(
     TROUBLE_SLEEPING = case_when(
-      PHQSLEEP %in% c(1, 2, 3) ~ 1, #several days, more than half days, nearly every day are trouble sleeping
-      PHQSLEEP == 0            ~ 0,  #not at all is no trouble sleeping
-      PHQSLEEP %in% c(6, 7, 8, 9) ~ NA_real_,  #NIU & Unknowns are NA
-      TRUE ~ NA_real_  #default NA for unexpected values
-    )) %>%
+      PHQSLEEP %in% c(1, 2, 3) ~ 1,  # several days or more → trouble sleeping
+      PHQSLEEP == 0            ~ 0,  # not at all
+      PHQSLEEP %in% c(6, 7, 8, 9) ~ NA_real_,
+      TRUE ~ NA_real_
+    )
+  ) %>%
   mutate(
     SOCIAL = case_when(
-      CMPSUPPORT == 1 ~ 0,   #more social and emotional support
-      CMPSUPPORT == 2 ~ 1,  #less social and emotional support
-      CMPSUPPORT == 3 ~ 0,   #about the same
-      CMPSUPPORT %in% c(0, 7, 8, 9) ~ NA_real_,  # NIU & Unknowns → NA
-      TRUE ~ NA_real_  #default NA for unexpected values
-    )) %>%
+      CMPSUPPORT == 1 ~ 0,  # more support
+      CMPSUPPORT == 2 ~ 1,  # less support
+      CMPSUPPORT == 3 ~ 0,  # about the same
+      CMPSUPPORT %in% c(0, 7, 8, 9) ~ NA_real_,
+      TRUE ~ NA_real_
+    )
+  ) %>%
   mutate(
     COVID = case_when(
-      CVDTESTRSLT == 2 ~ 1,  #yes, had COVID
-      CVDTESTRSLT == 1 ~ 0,  #no, did not have COVID
-      CVDTESTRSLT %in% c(0, 3, 7, 8, 9) ~ NA_real_,  #NIU, did not receive results, unknowns, and NA are NA
-      TRUE ~ NA_real_  #default NA for unexpected values
-    ))
+      CVDTESTRSLT == 2 ~ 1,  # had COVID
+      CVDTESTRSLT == 1 ~ 0,  # did not have COVID
+      CVDTESTRSLT %in% c(0, 3, 7, 8, 9) ~ NA_real_,
+      TRUE ~ NA_real_
+    )
+  )
 
@@
 # Create a subset of variables for imputation
-data1 <- labelled::remove_val_labels(data_original) #removes labels from haven package which cause problems
+
 imputation_vars <- c("AGE", "PA18AERSTR", "HRSLEEP", "DEPFREQ", 
@@
                                                            df$HRSLEEP[which(df$AGE >= 13 & df$AGE <= 18)] <= 10, 1, 0)
+  df$SLEEP[which(df$AGE > 18)] <- ifelse(df$HRSLEEP[which(df$AGE > 18)] >= 7 &
+                                           df$HRSLEEP[which(df$AGE > 18)] <= 9, 1, 0)
 

### 2025-04-30 products/manuscript/Manuscript.qmd
Part 6 updates

@@
 
-During the SARS-CoV pandemic, methods implemented to control the transmission of the disease had broader societal impacts.
-A common complaint surrounding social distancing and quarantine policies is the effects on mental health 1.
-This study aims to better understand the associations between mental health factors and contracting covid.
-Using data from the National Health Interview Surveys collected during the SARS-CoV-2 pandemic, simple and multiple regression models are created.
+When the SARS-CoV-2 pandemic began in late 2019, there were very few control measures enacted within the United States.
+By Spring of 2020, this had drastically changed, and control measures were broadly put into effect across the country.
+Social distancing, quantining, isolation, and lockdowns were prioritized control of transmission, but societal impacts were not initially considered.
+Disruptions to daily routine and social interactions quickly became a widespread concerns for their potential impacts on mental health, with widespread increases in anxiety, depression, loneliness, and sleep disturbances (CITATION NEEDED).
+
+Pandemic control strategies have become commonly criticized for the impacts on individuals' psychological well-being.
+Research documents the mental health consequences of pandemic control strageies (CITATION SAME AS BEFORE), but the potential relationships between mental health factors and vulnerability to SARS-CoV-2 infection is less understood.
+
+This study seeks to add to the limited understanding of mental health and COVID-19 vulnerability by expliorung associations between mental health factors and COVID-19 test positivity status.
+Self reported bouts of depressed mood, perceived social support, amount of exercise, and sleep quality (via hours of sleep and reporting difficulty falling asleep) are examined.
+(CITATION HERE) Simple and multiple regression models including Lasso and Random Forest are used to explore the predicitve power of mental health factors to a positive COVID-19 test.
 
@@
 
-Factors studied will include hours of sleep, reported depressed mood, reported trouble sleeping, recent exercise, and reported social/emotional support.
+Factors studied include:
+1.) Average self-reported hours of sleep per night, 
+2.) Self-reported depressed mood, 
+3.) Self-reported trouble sleeping, 
+4.) Self-reported average weekly exercise, and
+ 5.) Self-reported changes to social and emotional support.
 
@@
 
-Binary variables were created to allow for interpretability.
-EXERCISE is a combination of number of minutes of passive or aerboic exercise, descirbing whther weekly amount meets the American Heart Association's guidelines for exercise needed to stay healthy, as exercise is known to relate to mental health.
+Data was used in 2020 and after, given the NHIS didn't have questions about COVID-19 testing in 2019, these questions began in 2020. Binary variables were created to allow for interpretability.
+
+EXERCISE is a combination of number of minutes of passive or aerboic exercise, descirbing whether weekly amount meets the American Heart Association's guidelines for exercise needed to stay healthy.
+For adults, this includes 150 minutes per week of moderate intensity aerobic activity, 75 minutes per week of vigorous aerobic activity, or a combination of both.Children ages 6-17 should get 60 minutes of moderate to vigorous intenisty physical activity.
+A value of 1 for this binary variable means that the individual met exercise recommendations, and a 0 means that the individual did not meet exercise recommendations.
+
 SLEEP is based on the NIH guidelines for recommended hours of sleep per day, which differs by age.
-DEPRESSED includes if the individual reports feeling depressed on a weekly or daily basis, but not at a frequency of monthly or less.
-TROUBLE_SLEEPING indicates struggling to fall asleep several days per week, more than half the days, or nearly every day.
-SOCIAL describes individual feeling like they have less social support than their regular amount.
-Answers for don't know or unsure were converted to missing.
+Individuals under age 1 are suggested to get 12-16 hours of sleep per day, children ages 1 to 2 are suggested to get 11-14 hours of sleep per day, children ages 3 to 5 years are suggested to get 10-13 hours of sleep per day, children ages 6 to 12 years are suggested to get 9-12 hours of sleep per day, teens ages 13-18 are suggested to get 8-10 hours of sleep per day, and adults (over the age of 18) are suggested to get 7-9 hours of sleep per day.
+Guidelines do not require that sleep takes place at nighttime, and details that naps count towards sleep totals.
+A value of 1 for this binary variable means that the individual met recommendations for hours of sleep per day, and a 0 means that the individual did not meet recommendations for hours of sleep recommendations.
+
+DEPRESSED includes a 1 if the individual reports feeling depressed on a weekly or daily basis, but a 0 for feeling depressed at a frequency of monthly or less.
+
+TROUBLE_SLEEPING includes a 1 if the individual reports struggling to fall asleep several days per week, more than half the days, or nearly every day, and 0 if they report otherwise.
+
+SOCIAL includes a value of 1 for an individual describing feeling like they have less social support than their regular amount, and a 0 for feeling like they recieve the same or more support.
+
+For all variables, answers for don't know or unsure were converted to missing.
 
@@
 
-After creating binary variables, significant amounts of data were missing.
-Multiple imputation was used to create 5 sets of imputed data, and used for the remainder of analysis.
-To calculate later statisitcs, processes were preformed on each of the 5 sets and then averaged.
+After creating binary variables, significant amounts of data were missing, seen in table below.
+The proportions of values for answers 1 and 0 for binary variables are shown out out total answers, and then a column including NAs are shown, which details the proportions of NA answers among total individuals included.
 
-Complex survey design was used to allow for generalizability of results, using variables PSU, STRATA, and SAMPWEIGHT.
-A summary table was generated to find occurence of each of the factors and their inverses among the data.
-Next, single models were ran to find associations between each variable with covid test postivity.
-To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.
+Multiple imputation was used to create 5 sets of imputed data, and used for the remainder of analysis. To calculate later statistics, processes were preformed on each of the 5 sets and then averaged.
+
+```{r}
+#| label: tab-1
+#| fig-cap: "Missingness table"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures", "summary_table_with_nas.png"))
+```
 
@@
 ## Exploratory/Descriptive analysis
+
+Complex survey design was used to allow for generalizability of results, using variables PSU, STRATA, and SAMPWEIGHT.
+A summary table was generated to find occurence of each of the factors and their inverses among the data.
+Next, single models were ran to find associations between each variable with covid test postivity.
+To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.
 

### 2025-04-30 products/manuscript/Manuscript.qmd
Part 6 updates

@@
 
-Complex survey design was used to allow for generalizability of results, using variables PSU, STRATA, and SAMPWEIGHT.
-A summary table was generated to find occurence of each of the factors and their inverses among the data.
-Next, single models were ran to find associations between each variable with covid test postivity.
-To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.
+To examine prevelance of mental health factors among the United States population, compmlex survey design was used to allow for generalizability of results, specifically, variables PSU, STRATA, and SAMPWEIGHT.
+A summary table was generated to find occurence of each of the factors and their inverses among the data, seen below.
 
@@
 
+Next, single models were ran to find associations between each variable with covid test postivity. Complex survey design was used to allow generalizability of results to the United States population.
+
 ```{r}
@@
 ## Full analysis
+
+To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.
 

### 2025-05-01 Joseph_project-review.qmd
Joseph Review

---
title: Mental Health Factors and Covid
author: Asmith Joseph 
date: date-modified
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Mental Health Factors and Covid

Name of project author(s): Annalise Cramer

Name of project reviewer: Asmith Joseph

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/renders into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The introduction addresses an important and timely issue but would be stronger with a more focused narrative. 
Some Suggestions: 
•	Start by summarizing three to five pivotal studies that examine the direct effects of COVID-19 infection, rather than lockdown alone, on sleep quality, physical activity, mood disturbances, and social support.
•	After reviewing these findings, highlight the remaining gap, no research has yet leveraged nationally representative NHIS data to compare mental health outcomes between adults who were infected and those who were not across multiple domains. 
•	Support each point with concise, informative citations. Conclude with a clear gap-to-aim statement: “Although prior studies have documented the individual impacts of COVID-19 infection on sleep, exercise, mood, and social connectedness, no study to date has simultaneously evaluated these outcomes in a nationally representative cohort; this analysis will fill that void.”

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* some contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The questions do provide the readers a sense of the general aims; however, they lack the precision and framing of formal hypotheses, for example, a null and alternative hypothesis or clearly articulated expected direction of association. It isn’t immediately obvious how each question maps onto the NHIS data structure.

I recommend rewriting this section as two or three specific aims (e.g., “Aim 1: Estimate the weighted prevalence of self-reported poor mental health among U.S. adults during 2019–2021”; “Aim 2: Test whether adults with laboratory-confirmed COVID-19 have higher odds of reporting poor mental health after adjusting for age, sex, and survey year”), and explicitly noting which NHIS variables and sample weights will be used to answer each aim. That will make the relationship between question, hypothesis, and data crystal clear.

### Summary assessment
* question/hypotheses somewhat explained

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The manuscript tells us you’re using a large national survey of adults during the pandemic, but the description of the data stops at “we used NHIS.” You never specify which survey years or waves, how Covid‐infection status was ascertained, which mental‐health scales were used or where readers can find their scoring rules, nor do you point to a public codebook or variable glossary.

Without at least a brief “Data Source and Measures” subsection, complete with the exact dataset version, links to the NHIS codebook (or appendix file), and a table of your key variables (name, question wording, response options, handling of missing values), it’s hard for readers to know exactly what you measured or to replicate your work.

### Summary assessment
* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

In term of Data wrangling and exploratory analysis, only broad cleaning (dropping missing cases, recoding infection) and doesn’t detail how you screened for outliers, chose inclusion thresholds, or compared imputation methods. Exploratory analysis is likewise confined to one table, with no supplemental histograms, correlations, or group comparisons to verify variable behavior.

Some suggestions: 
Add a “Data Cleaning & Exploration” section that (1) lists each preprocessing step and its rationale, (2) briefly discusses alternative approaches you considered (e.g., multiple imputation vs. complete case), and (3) includes key EDA visuals (histograms, correlation matrices, stratified plots) in an appendix so readers can see your data are sound.

### Summary assessment
* major weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The combination of logistic regression with interactions and a machine-learning pipeline is appropriate for linking COVID-19 infection to mental health outcomes. However, the manuscript should:

•	Clarify how predictors were chosen (theory, univariate screening, or regularization).
•	Detail data pre-processing (missing data, outliers, categorical encoding).
•	Go beyond AUC—report sensitivity, specificity, calibration, or decision-curve analyses.
•	Address class imbalance and threshold selection.

Overall, the methods are sound but need greater transparency in variable selection, preprocessing, and a richer set of evaluation metrics.

### Summary assessment
* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

So far only a few tables and figures, but they are well structured. Most tables use sensible row–column layouts and include confidence intervals or p-values, and the figures effectively visualize interaction effects and subgroup trends. 
Some suggestions
•	Streamline dense tables: Move secondary details (e.g., extensive covariate lists or robustness checks) into supplementary material to keep primary tables focused.
•	Standardize formatting: Use consistent fonts, decimal places, and table borders throughout.
•	Enhance figure legends and output quality: Write concise, panel‐specific captions; produce high‐resolution graphics; and choose color palettes that remain distinguishable in black‐and‐white print and are accessible to color‐blind readers.

### Summary assessment

* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Since a Discussion section hasn’t been drafted yet, here are some tips to guide its structure:

- Begin with a concise summary of your key findings, emphasizing, for example, that COVID-19 infection was linked to poorer sleep, mood disturbances, and reduced social support in a nationally representative sample spanning multiple mental-health domains.
- Contextualize these results by comparing them to prior studies—highlight where you confirm or extend existing evidence—and explicitly acknowledge your study’s strengths (e.g., large, population-based data; comprehensive mental-health measures) alongside its weaknesses (e.g., cross-sectional design, self-report bias, possible residual confounding).
- Interpret your findings through plausible mechanisms, such as inflammation-mediated neurochemical changes that may drive mood and sleep disruptions, and discuss how unmeasured factors might influence the observed associations.
- Propose concrete next steps, like longitudinal follow-up to assess causality and incorporation of biological markers (e.g., cytokine levels) to validate mechanistic hypotheses.
- Cite relevant literature throughout—both to support your comparisons with past work and to bolster mechanistic explanations—so readers can readily see how your results fit into and advance the field.

### Summary assessment
* major parts of discussion missing or wrong

## Further comments

_Add any other comments regarding the different aspects of the project here. Write anything you think can help your classmate improve their project._

So far you have a strong foundation, a good start. However, you still have some work to do. Keep at it. You can do it!
Some Suggestions: 
•	Tighten transitions between sections, explicitly introduce each table and figure in the text, and make sure your methods naturally flow into your results.
•	Add a few descriptive plots or a correlation matrix (in supplementary material if necessary) to justify your choice of predictors and to highlight any surprising patterns before modeling.
•	Anchor your findings in biological or psychosocial theory (e.g., inflammation-driven mood changes), acknowledge potential residual confounding, and suggest concrete next steps like longitudinal follow-up or biomarker validation.
•	Standardize fonts, decimal places, and table borders; use concise figure legends that map directly to panels; and ensure high-resolution, accessible color palettes.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Yes, no issues navigating through dufferent folders

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Yes, well structured so far.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

No, not all the way reproducible. But I am sure once everything is completed it should be.

### Summary assessment
* small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The manuscript is not complete. There are still a lot of sections missing

### Summary assessment
* weak level of thoroughness

## Further comments

_Add any other comments regarding the overall project here. Write anything you think can help your classmate improve their project._

Keep working, you got this! best of luck.

### 2025-05-01 code/analysis-code/analysis.qmd
Part 6 updates

@@
 library(gt)
+library(scales)
 library(pROC)
@@
 
-Next, we will make the multivariate model. I will use cross validation for this. The pooled AUC across imputations is found to be 0.5230.
+Next, we will make the multivariate model. I will use cross validation for this.Pooled AUC is calculated and a observed-expected plot is generated.
 ```{r}
-# Define the AUC cross-validation function for imputed datasets
-cross_validation_imputed_auc <- function(imputed_data, predictors, k_folds = 5) {
+library(pROC)
+library(dplyr)
+
+cross_validation_imputed_auc_logit <- function(imputed_data, predictors, k_folds = 5) {
   auc_list <- numeric(length(imputed_data))  # Store AUC for each imputation
 
-  for (i in 1:length(imputed_data)) {
+  for (i in seq_along(imputed_data)) {
     data <- imputed_data[[i]]
 
-    # drop rows with missing values in COVID, predictors, or SAMPWEIGHT
+    # Drop rows with missing values in outcome, predictors, or weights
     data <- data %>% drop_na(all_of(c("COVID", predictors, "SAMPWEIGHT")))
@@
 
-    fold_auc <- numeric(k_folds)  # AUC for each fold
+    fold_auc <- numeric(k_folds)
 
@@
 
-      # Fit random forest model on training data
+      # Fit logistic regression model
       formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
-      rf_model <- ranger(
+      logit_model <- glm(
         formula = formula,
         data = train_data,
-        case.weights = train_data$SAMPWEIGHT,
-        probability = TRUE
+        family = binomial(),
+        weights = SAMPWEIGHT
       )
 
-      # Predict probabilities for class 1 (COVID = 1)
-      pred_probs <- predict(rf_model, data = test_data)$predictions[, 2]
+      # Predict probabilities on test data
+      pred_probs <- predict(logit_model, newdata = test_data, type = "response")
       actual <- test_data$COVID
@@
 
-    # Average AUC over folds for this imputed dataset
     auc_list[i] <- mean(fold_auc)
@@
 
-  # Average AUC over all imputed datasets
   overall_auc <- mean(auc_list)
@@
 
-# Assuming you have your imputed_data and a vector of predictors
-results <- cross_validation_imputed_auc(imputed_data = data_imputed,
-                                        predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL"))
+# Usage
+results_logit <- cross_validation_imputed_auc_logit(
+  imputed_data = data_imputed,
+  predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
+)
 
 # View results
-results$auc_per_imputation  # AUC for each imputed dataset
-results$overall_auc         # Average AUC across imputations
+results_logit$auc_per_imputation
+results_logit$overall_auc
+
+# Function to compute calibration data per imputation
+get_calibration_data <- function(data, predictors, n_bins = 10) {
+  # Fit logistic regression model
+  formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
+  model <- glm(formula, data = data, family = binomial(), weights = SAMPWEIGHT)
+
+  # Predict probabilities
+  data$predicted <- predict(model, type = "response")
+
+  # Bin predictions into deciles
+  data <- data %>%
+    mutate(bin = ntile(predicted, n_bins)) %>%
+    group_by(bin) %>%
+    summarise(
+      mean_pred = mean(predicted),
+      obs_rate = mean(COVID),
+      .groups = "drop"
+    )
+
+  return(data)
+}
+
+# Combine calibration data across imputations
+all_calib_data <- purrr::map_dfr(
+  .x = seq_along(data_imputed),
+  .f = function(i) {
+    data <- data_imputed[[i]] %>% 
+      drop_na(COVID, EXERCISE, SLEEP, DEPRESSED, TROUBLE_SLEEPING, SOCIAL, SAMPWEIGHT)
+
+    calib <- get_calibration_data(data, predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL"))
+    calib$imputation <- paste0("Imputation ", i)
+    return(calib)
+  }
+)
+
+# Plot calibration curves
+LogReg_plot <- ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
+  geom_line() +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
+  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  labs(title = "Calibration Plot: Logistic Regression (All 5 Imputations)", color = "Imputation") +
+  theme_minimal()
+
+# Save the plot using the `here` package
+ggsave(filename = here::here("results", "figures", "log_calibration_plot.png"),
+       plot = LogReg_plot,
+       width = 8, height = 6, dpi = 300)
+
 ```
@@
 Instead of using complex survey design, I will weight the model by sampleweight.
-The pooled AUC is  0.529.
+Pooled AUC is calculated and a observed-expected plot is generated.
 
 ```{r}
-# Extract the first completed imputed dataset
-data1 <- data_imputed[[1]]
+library(glmnet)
+library(caret)
+library(Metrics)  # for rmse()
+library(dplyr)
 
-# Define the survey design
-survey_design <- svydesign(
-  ids = ~PSU,
-  strata = ~STRATA,
-  weights = ~SAMPWEIGHT,
-  data = data1,
-  nest=TRUE
+set.seed(1234)
+k <- 5
+auc_list <- c()
+
+# Loop through 5 imputed datasets
+for (i in 1:5) {
+  data1 <- data_imputed[[i]]
+
+  # Remove rows with missing outcome
+  data1_complete <- data1 %>% filter(!is.na(COVID))
+
+  # Create folds
+  folds <- createFolds(data1_complete$COVID, k = k, list = TRUE, returnTrain = FALSE)
+
+  preds <- rep(NA, nrow(data1_complete))
+
+  for (fold_idx in seq_along(folds)) {
+    test_idx <- folds[[fold_idx]]
+    train_idx <- setdiff(seq_len(nrow(data1_complete)), test_idx)
+
+    train_data <- data1_complete[train_idx, ] %>% drop_na()
+    test_data  <- data1_complete[test_idx, ] %>% drop_na()
+
+    # Create model matrix (remove outcome column)
+    x_train <- model.matrix(COVID ~ ., data = train_data)[, -1]
+    y_train <- train_data$COVID
+
+    x_test <- model.matrix(COVID ~ ., data = test_data)[, -1]
+
+    # Fit Lasso with CV
+    lasso_fit <- cv.glmnet(x = x_train, y = y_train, alpha = 1, family = "binomial")
+
+    # Predict probabilities (type = "response" for prob)
+    preds[test_idx] <- predict(lasso_fit, newx = x_test, s = "lambda.min", type = "response")
+  }
+
+  actuals <- data1_complete$COVID
+
+  # AUC using pROC
+  roc_obj <- roc(actuals, preds)
+  fold_auc <- auc(roc_obj)
+  auc_list[i] <- fold_auc
+}
+
+pooled_auc <- mean(auc_list)
+print(pooled_auc)
+
+################################################################################
+## plot expected vs observed
+
+set.seed(1234)
+k <- 5
+all_preds_df <- data.frame()
+
+for (i in 1:5) {
+  data1 <- data_imputed[[i]]
+  data1_complete <- data1 %>% filter(!is.na(COVID))
+
+  folds <- createFolds(data1_complete$COVID, k = k, list = TRUE, returnTrain = FALSE)
+
+  preds <- rep(NA, nrow(data1_complete))
+
+  for (fold_idx in seq_along(folds)) {
+    test_idx <- folds[[fold_idx]]
+    train_idx <- setdiff(seq_len(nrow(data1_complete)), test_idx)
+
+    train_data <- data1_complete[train_idx, ] %>% drop_na()
+    test_data  <- data1_complete[test_idx, ] %>% drop_na()
+
+    x_train <- model.matrix(COVID ~ ., data = train_data)[, -1]
+    y_train <- train_data$COVID
+    x_test  <- model.matrix(COVID ~ ., data = test_data)[, -1]
+
+    lasso_fit <- cv.glmnet(x = x_train, y = y_train, alpha = 1, family = "binomial")
+
+    preds[test_idx] <- predict(lasso_fit, newx = x_test, s = "lambda.min", type = "response")
+  }
+
+  all_preds_df <- bind_rows(all_preds_df, data.frame(
+    imputation = i,
+    predicted = preds,
+    observed = data1_complete$COVID
+  ))
+}
+
+# Bin predictions into 10 bins and compute calibration stats
+calib_data <- all_preds_df %>%
+  mutate(bin = ntile(predicted, 10)) %>%
+  group_by(imputation, bin) %>%
+  summarise(
+    mean_pred = mean(predicted, na.rm = TRUE),
+    obs_rate = mean(observed, na.rm = TRUE),
+    .groups = "drop"
   )
 
-set.seed(1234)  # for reproducibility
+Lasso_plot <- ggplot(calib_data, aes(x = mean_pred, y = obs_rate, color = factor(imputation))) +
+  geom_line() +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
+  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  labs(
+    title = "Calibration Plot: Lasso Logistic Regression (All 5 Imputations)",
+    color = "Imputation"
+  ) +
+  theme_minimal()
 
-# Choose outcome and predictors
-outcome_var <- "COVID"
-predictor_vars <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
+print(Lasso_plot)
 
-# Loop over imputations and compute AUC
-auc_list <- map(data_imputed, function(dat) {
-  # Drop missing in predictors or outcome if any slipped through
-  dat <- dat %>% drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))
+# Save the plot using the `here` package
+ggsave(filename = here::here("results", "figures", "lasso_calibration_plot.png"),
+       plot = Lasso_plot,
+       width = 8, height = 6, dpi = 300)
 
-  # Create design matrix X and response y
+################################################################################
+# Keep variable names for predictors
+formula <- as.formula("COVID ~ .")
+
+# Explicitly remove rows with any NA in predictors (even if COVID is not NA)
+train_data <- train_data %>% select(COVID, everything()) %>% drop_na()
+
+# Now get model matrix
+x_train <- model.matrix(formula, data = train_data)[, -1]
+y_train <- train_data$COVID
+
+nrow(x_train)
+length(y_train)
+```
+
+```{r}
+# Load required libraries
+library(glmnet)      # For Lasso regression
+library(pROC)        # For AUC calculation
+library(ggplot2)     # For plotting
+library(dplyr)       # For data manipulation
+library(mice)        # For working with imputed data
+library(tidyr)       # For data reshaping
+
+# --- LASSO MODEL WITH CROSS-VALIDATION ON IMPUTED DATA ---
+
+# Function to run Lasso with sample weights and calculate performance metrics
+run_lasso_cv <- function(imp_data, outcome = "COVID", 
+                        predictors = c("SLEEP", "DEPRESSED", "SOCIAL", "TROUBLE_SLEEPING", "EXERCISE"),
+                        weight_var = "sampleweight") {
+
+  # Extract data from current imputation
+  X <- as.matrix(imp_data[, predictors])
+  y <- imp_data[[outcome]]
+  weights <- imp_data[[weight_var]]
+
+  # Standardize predictors (important for Lasso)
+  X_scaled <- scale(X)
+
+  # Run cross-validated Lasso with sample weights
+  set.seed(123) # For reproducibility
+  cv_model <- cv.glmnet(
+    x = X_scaled,
+    y = y,
+    family = "binomial",
+    alpha = 1,       # 1 = Lasso
+    weights = weights,
+    nfolds = 5,
+    type.measure = "auc"
+  )
+
+  # Get the best lambda (penalty parameter)
+  best_lambda <- cv_model$lambda.min
+
+  # Refit model with best lambda
+  final_model <- glmnet(
+    x = X_scaled,
+    y = y,
+    family = "binomial",
+    alpha = 1,
+    weights = weights,
+    lambda = best_lambda
+  )
+
+  # Get predictions
+  pred_probs <- predict(final_model, newx = X_scaled, type = "response")
+
+  # Calculate AUC
+  roc_obj <- roc(y, pred_probs)
+  auc_value <- auc(roc_obj)
+
+  # Create calibration data for expected vs observed plot
+  n_groups <- 10  # Number of bins for calibration plot
+
+  # Create calibration data
+  calib_data <- data.frame(
+    predicted = as.vector(pred_probs),
+    observed = y
+  )
+
+  # Sort by predicted probability and divide into bins
+  calib_data <- calib_data[order(calib_data$predicted), ]
+  calib_data$bin <- cut(seq(1, nrow(calib_data)), breaks = n_groups, labels = FALSE)
+
+  # Calculate mean predicted and observed values for each bin
+  calibration <- calib_data %>%
+    group_by(bin) %>%
+    summarize(
+      n = n(),
+      mean_predicted = mean(predicted),
+      mean_observed = mean(observed)
+    )
+
+  # Return results
+  return(list(
+    model = final_model,
+    lambda = best_lambda,
+    auc = auc_value,
+    calibration = calibration,
+    coefs = coef(final_model)
+  ))
+}
+
+# ----------------------------
+# MAIN ANALYSIS
+# ----------------------------
+
+# Check the structure of the imputed data
+print(class(data_imputed))
+
+# Initialize results containers
+auc_values <- numeric(length = 5)
+calibration_results <- list()
+coef_results <- list()
+
+# Function to run Lasso with sample weights and calculate performance metrics
+run_lasso_cv <- function(imp_data, outcome = "COVID", 
+                        predictors = c("SLEEP", "DEPRESSED", "SOCIAL", "TROUBLE_SLEEPING", "EXERCISE"),
+                        weight_var = "sampleweight") {
+
+  # Extract data from current imputation and handle missing values
+  # First, check for and handle missing data
+  complete_cases <- complete.cases(imp_data[, c(outcome, predictors, weight_var)])
+  if (!all(complete_cases)) {
+    message("Removing ", sum(!complete_cases), " rows with missing values")
+    imp_data <- imp_data[complete_cases, ]
+  }
+
+  X <- as.matrix(imp_data[, predictors])
+  y <- imp_data[[outcome]]
+  weights <- imp_data[[weight_var]]
+
+  # Standardize predictors (important for Lasso)
+  X_scaled <- scale(X)
+
+  # Run cross-validated Lasso with sample weights
+  set.seed(123) # For reproducibility
+  cv_model <- cv.glmnet(
+    x = X_scaled,
+    y = y,
+    family = "binomial",
+    alpha = 1,       # 1 = Lasso
+    weights = weights,
+    nfolds = 5,
+    type.measure = "auc"
+  )
+
+  # Get the best lambda (penalty parameter)
+  best_lambda <- cv_model$lambda.min
+
+  # Refit model with best lambda
+  final_model <- glmnet(
+    x = X_scaled,
+    y = y,
+    family = "binomial",
+    alpha = 1,
+    weights = weights,
+    lambda = best_lambda
+  )
+
+# Combine calibration results from all imputations
+all_calibration <- do.call(rbind, calibration_results)
+
+# ----------------------------
+# OUTPUT RESULTS
+# ----------------------------
+
+# Print AUC values for each imputation
+cat("\nAUC values for each imputation:\n")
+print(auc_values)
+cat("\nMean AUC across imputations:", mean(auc_values), "\n")
+
+# Plot expected vs observed probabilities for all imputations
+calibration_plot <- ggplot(all_calibration, aes(x = mean_predicted, y = mean_observed, color = factor(imputation))) +
+  geom_point(size = 3) +
+  geom_line() +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
+  labs(
+    title = "Expected vs. Observed Probabilities",
+    subtitle = paste("Mean AUC =", round(mean(auc_values), 3)),
+    x = "Expected (Predicted Probability)",
+    y = "Observed (Actual Proportion)",
+    color = "Imputation"
+  ) +
+  theme_minimal() +
+  coord_equal() +
+  xlim(0, 1) + 
+  ylim(0, 1)
+
+print(calibration_plot)
+
+# Print average coefficients across imputations
+cat("\nAverage Lasso coefficients across imputations:\n")
+# Extract coefficient names from the first imputation
+coef_names <- rownames(coef_results[[1]])
+
+# Create matrix to store coefficients from all imputations
+all_coefs <- matrix(0, nrow = length(coef_names), ncol = 5)
+rownames(all_coefs) <- coef_names
+
+# Fill in the coefficient matrix
+for (i in 1:5) {
+  all_coefs[, i] <- as.vector(coef_results[[i]])
+}
+
+# Calculate mean coefficients
+mean_coefs <- rowMeans(all_coefs)
+coef_summary <- data.frame(
+  Variable = coef_names,
+  Mean_Coefficient = mean_coefs
+)
+
+# Print non-zero coefficients
+non_zero_coefs <- coef_summary[abs(coef_summary$Mean_Coefficient) > 0.001, ]
+print(non_zero_coefs[order(abs(non_zero_coefs$Mean_Coefficient), decreasing = TRUE), ])
+```
+
+```{r}
+library(dplyr)
+library(ggplot2)
+library(scales)
+
+calib_data_list <- vector("list", length(data_imputed))
+
+for (i in seq_along(data_imputed)) {
+  dat <- data_imputed[[i]] %>%
+    drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))
+
   X <- model.matrix(as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))), dat)[, -1]
@@
 
-  # Split into training and test sets (e.g., 70/30 split)
+  set.seed(1234 + i)
   train_idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
@@
   y_test <- y[test_idx]
-  w_test <- w[test_idx]
 
-  # Fit random forest model (instead of lasso, because it's a classification problem)
-  rf_model <- ranger(
-    formula = as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))),
-    data = dat[train_idx, ],
-    case.weights = w_train,
-    probability = TRUE
+  cv_fit <- cv.glmnet(
+    x = X_train,
+    y = y_train,
+    weights = w_train,
+    alpha = 1,
+    family = "binomial",
+    standardize = TRUE
   )
 
-  # Predict probabilities for class 1
-  pred_probs <- predict(rf_model, data = dat[test_idx, ])$predictions[,2]
+  pred_probs <- predict(cv_fit, newx = X_test, s = "lambda.min", type = "response") %>% as.vector()
+
+  # Count unique predicted probabilities
+  n_unique_preds <- length(unique(pred_probs))
+
+  # Only proceed with binning if there are at least 2 unique predicted probabilities
+  if (n_unique_preds >= 2) {
+    # Safely bin the predicted probabilities
+    n_bins <- min(5, n_unique_preds)
+
+    calib_data <- tibble(
+      actual = y_test,
+      predicted = pred_probs
+    ) %>%
+      mutate(bin = cut_number(predicted, n = n_bins, labels = FALSE)) %>%
+      group_by(bin) %>%
+      summarise(
+        mean_pred = mean(predicted),
+        obs_rate = mean(actual),
+        .groups = "drop"
+      ) %>%
+      mutate(imputation = paste0("Imputation ", i))
+
+    calib_data_list[[i]] <- calib_data
+  } else {
+    message("Skipping calibration plot for imputation ", i, ": too few unique predicted values (", n_unique_preds, " unique values)")
+  }
+}
+
+# Combine all calibration data
+all_calib_data <- bind_rows(calib_data_list)
+
+# Create calibration plot
+ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
+  geom_line() +
+  geom_point(size = 2) +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
+  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
+  theme_minimal()
+
+# Combine calibration data
+all_calib_data <- bind_rows(calib_data_list)
+
+# Plot calibration curve
+ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
+  geom_line() +
+  geom_point(size = 2) +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
+  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
+  theme_minimal()
+
+# Combine calibration data
+all_calib_data <- bind_rows(calib_data_list)
+
+# Plot calibration curve
+ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
+  geom_line() +
+  geom_point(size = 2) +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
+  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
+  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
+  theme_minimal()
+# Save the plot
+ggsave(filename = here::here("results", "figures", "lasso_calibration_plot.png"),
+       plot = calibration_plot_lasso,
+       width = 8, height = 6, dpi = 300)
+
+# Initialize list to store AUCs
+auc_list <- vector("numeric", length(data_imputed))
+
+for (i in seq_along(data_imputed)) {
+  dat <- data_imputed[[i]] %>%
+    drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))
+
+  X <- model.matrix(as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))), dat)[, -1]
+  y <- dat[[outcome_var]]
+  w <- dat$SAMPWEIGHT
+
+  set.seed(1234 + i)
+  train_idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
+  test_idx <- setdiff(seq_len(X), train_idx)
+
+  X_train <- X[train_idx, ]
+  y_train <- y[train_idx]
+  w_train <- w[train_idx]
+
+  X_test <- X[test_idx, ]
+  y_test <- y[test_idx]
+
+  # Fit lasso model
+  cv_fit <- cv.glmnet(
+    x = X_train,
+    y = y_train,
+    weights = w_train,
+    alpha = 1,
+    family = "binomial",
+    standardize = TRUE
+  )
+
+  # Predict probabilities
+  pred_probs <- predict(cv_fit, newx = X_test, s = "lambda.min", type = "response") %>% as.vector()
 
   # Compute AUC
-  auc_result <- roc(y_test, pred_probs)$auc
-  return(auc_result)
-})
+  auc_val <- roc(y_test, pred_probs)$auc
+  auc_list[i] <- auc_val
+}
 
-# Pool AUC across imputations
-mean_auc <- mean(unlist(auc_list))
+# Print AUCs
+cat("AUC for each imputation:\n")
+print(round(auc_list, 3))
 
-# Output
-cat("AUC for each imputation:\n")
-print(unlist(auc_list))
-cat("\nPooled AUC across imputations:", round(mean_auc, 3), "\n")
+# Mean (pooled) AUC
+cat("\nPooled AUC across imputations:", round(mean(auc_list), 3), "\n")
 ```
@@
 For this, I will not use the multiple imputation dataset, as before, I will use the first imputed dataset, with cross validation 5 folds.
-The pooled AUC is 0.556.
+Pooled AUC is calculated and a observed-expected plot is generated.
 
 ```{r}
-library(ranger)
-library(caret)
-library(pROC)
-library(dplyr)
+cross_validation_imputed_predictions <- function(imputed_data, predictors, k_folds = 5) {
+  auc_list <- numeric(length(imputed_data))  # Store AUCs
+  predictions_list <- vector("list", length(imputed_data))  # Store predictions and actuals
 
-set.seed(1234)
-k <- 5
-auc_list <- numeric(length(data_imputed))
+  for (i in 1:length(imputed_data)) {
+    data <- imputed_data[[i]] %>%
+      drop_na(all_of(c("COVID", predictors, "SAMPWEIGHT")))
 
-for (i in seq_along(data_imputed)) {
-  data1 <- data_imputed[[i]]
+    set.seed(123)
+    folds <- sample(1:k_folds, size = nrow(data), replace = TRUE)
 
-  # Remove missing outcome
-  data1_complete <- data1 %>% filter(!is.na(COVID))
+    fold_auc <- numeric(k_folds)
+    imputation_preds <- tibble()  # Empty tibble to collect preds for this imputation
 
-  # Make sure outcome is binary (0/1)
-  data1_complete$COVID <- as.numeric(data1_complete$COVID)
+    for (fold in 1:k_folds) {
+      train_data <- data[folds != fold, ]
+      test_data  <- data[folds == fold, ]
 
-  # Add weights column if not present
-  data1_complete$weights <- data1_complete$SAMPWEIGHT
-
-  # Create folds
-  folds <- createFolds(data1_complete$COVID, k = k, list = TRUE, returnTrain = FALSE)
-
-  auc_folds <- numeric(k)
-
-  for (fold_idx in seq_along(folds)) {
-    test_idx <- folds[[fold_idx]]
-    train_idx <- setdiff(seq_len(nrow(data1_complete)), test_idx)
-
-    train_data <- data1_complete[train_idx, ]
-    test_data  <- data1_complete[test_idx, ]
-
-    # Fit random forest classifier
+      formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
       rf_model <- ranger(
-      formula = COVID ~ EXERCISE + SLEEP + DEPRESSED + TROUBLE_SLEEPING + SOCIAL, #formula
+        formula = formula,
         data = train_data,
-      case.weights = train_data$weights,
+        case.weights = train_data$SAMPWEIGHT,
         probability = TRUE
@@
 
-    # Predict probabilities for class 1
-    pred_probs <- predict(rf_model, data = test_data)$predictions[,2]
+      pred_probs <- predict(rf_model, data = test_data)$predictions[, 2]
+      actual <- test_data$COVID
 
-    # Compute AUC
-    auc_folds[fold_idx] <- roc(test_data$COVID, pred_probs, quiet = TRUE)$auc
+      # Store predictions
+      imputation_preds <- bind_rows(imputation_preds, tibble(actual = actual, predicted = pred_probs))
+
+      # AUC
+      auc_result <- roc(actual, pred_probs)$auc
+      fold_auc[fold] <- auc_result
     }
 
-  # Store average AUC for this imputation
-  auc_list[i] <- mean(auc_folds)
+    auc_list[i] <- mean(fold_auc)
+    predictions_list[[i]] <- imputation_preds
   }
 
-# Pool AUC across imputations
-mean_auc <- mean(auc_list)
-print(paste("Pooled AUC across imputations:", round(mean_auc, 3)))
+  list(
+    auc_per_imputation = auc_list,
+    overall_auc = mean(auc_list),
+    predictions = predictions_list  # List of tibbles with actual + predicted
+  )
+}
+# Assuming you have your imputed_data and a vector of predictors
+results_with_preds <- cross_validation_imputed_predictions(
+  imputed_data = data_imputed,
+  predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
+)
+
+# View results
+results$auc_per_imputation  # AUC for each imputed dataset
+results$overall_auc         # Average AUC across imputations
+
+# Create calibration plot from predictions across imputations
+plot_calibration_multiple <- function(predictions_list, n_bins = 10, title = "Calibration Plot") {
+  calib_data_all <- map2_dfr(predictions_list, seq_along(predictions_list), function(df, i) {
+    df %>%
+      mutate(bin = ntile(predicted, n_bins)) %>%
+      group_by(bin) %>%
+      summarise(
+        mean_pred = mean(predicted),
+        obs_rate = mean(actual),
+        .groups = "drop"
+      ) %>%
+      mutate(imputation = paste0("Imputation ", i))
+  })
+
+  ggplot(calib_data_all, aes(x = mean_pred, y = obs_rate, color = imputation)) +
+    geom_line() +
+    geom_point(size = 2) +
+    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
+    scale_x_continuous("Expected probability", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
+    scale_y_continuous("Observed proportion", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
+    labs(title = title, color = "Imputation") +
+    theme_minimal()
+}
+
+# Plot it
+plot_cal <- plot_calibration_multiple(results_with_preds$predictions, title = "Calibration Plot: RF across 5 Imputations")
+print(plot_cal)
+
+# Save it
+ggsave(here("results", "figures", "rf_calibration_all_imputations.png"), plot_cal, width = 8, height = 6, dpi = 300)
+
 ```
@@
 gtsave(model_table, here("results", "figures", "model_table.png"))
+
+# Store your AUC lists here
+aauc_multivar <- results$auc_per_imputation  # From multivariate logistic regression with CV
+auc_lasso     <- unlist(auc_list)                      # From your LASSO section
+auc_rf        <- auc_list                              # From your final RF model section
+
+# Create a data frame with AUC values and model type
+auc_df <- data.frame(
+  Imputation = rep(1:5, 3),
+  AUC = c(auc_multivar, auc_lasso, auc_rf),
+  Model = rep(c("Multivariate Logistic Regression", "LASSO", "Random Forest"), each = 5)
+)
+
+# Plot: AUCs across imputations by model
+aucs <- ggplot(auc_df, aes(x = Model, y = AUC, color = Model)) +
+  geom_jitter(width = 0.2, size = 2) +         # dots for individual AUCs
+  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, color = "black") +
+  stat_summary(fun = mean, geom = "text", aes(label = round(..y.., 3)), 
+               vjust = -1, color = "black", fontface = "bold") +
+  labs(title = "AUC Across Imputations by Model",
+       x = "Model",
+       y = "AUC") +
+  theme_minimal()
+
+# Save the table as a PNG
+here()
+ggsave(
+  filename = here("results", "figures", "auc_plot.png"),
+  plot = aucs,  # replace with your plot object name
+  width = 8,
+  height = 6,
+  dpi = 300
+)
+
 ```
+
+Now, we will make expected versus observed plots.
+```{r}
+plot_calibration_multiple <- function(actual, predicted_list, n_bins = 10, title = "Calibration Plot") {
+  # Combine calibration data across imputations
+  calib_data_all <- map2_dfr(predicted_list, seq_along(predicted_list), function(predicted, i) {
+    tibble(actual = actual, predicted = predicted) %>%
+      mutate(bin = ntile(predicted, n_bins)) %>%
+      group_by(bin) %>%
+      summarise(
+        mean_pred = mean(predicted),
+        obs_rate = mean(actual),
+        .groups = "drop"
+      ) %>%
+      mutate(imputation = paste0("Imputation ", i))
+  })
+
+  # Plot
+  ggplot(calib_data_all, aes(x = mean_pred, y = obs_rate, color = imputation)) +
+    geom_line() +
+    geom_point(size = 2) +
+    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
+    scale_x_continuous("Expected probability", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
+    scale_y_continuous("Observed proportion", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
+    labs(title = title, color = "Imputation") +
+    theme_minimal()
+}
+
+# Example usage:
+plot_cal <- plot_calibration_multiple(
+  actual = test_data$COVID,
+  predicted_list = pred_probs_list,  # this should be a list of 5 vectors
+  title = "Random Forest Calibration Across 5 Imputations"
+)
+
+# Display plot
+print(plot_cal)
+
+# Save it
+ggsave(
+  filename = here("results", "figures", "rf_calibration_all_imputations.png"),
+  plot = plot_cal,
+  width = 8,
+  height = 6,
+  dpi = 300
+)
+
+```

### 2025-05-01 products/manuscript/Manuscript.qmd
Part 6 updates

@@
 
-To explore the data further, a multivariate model, LASSO regression, and a third advanced model (replace this) were ran with cross validation.
+To better understand the relationship between COVID-19 test status and mental health factors, three new models were constructed: a multivariate logistic regression model, LASSO regression, and a random forest model.
+Each model used the imputed data and five fold cross validation to affect predictive performance.
+All models accounted for sample wighting, but not clustering.
+Notably, survey design could not be fully accounted for due to the complexity of the LASSO and random forest models, making results not interpretable to the country-wide level, rather, predictive results are relevant within the NHIS population, which accounts for 110,216 individuals.
+
+All models included COVID-19 test status as the outcome and all mental helath variables as predictors (hours of sleep, touble sleeping, feeling depressed, meeting exercise guidelines, and amount of social support).
+The multivariate logistic regression was chosen for it's ease of interpretability and to serve as a baseline against the more advanced models.
+The LASSO model was chosen for it's ability to penalize predictors to indicate which mental health factors were most important.
+The random forest model was used to capture potential complex relationships between variables.
+
+To evalaute the performance of these models, the area under the curve (AUC) was calculated.
+This metric was chosen due to it's discrimination between binary outcome of the COVID-19 variable.
+Higher AUC indicates better model preformance.
+
+```{r}
+#| label: fig-aucplot
+#| fig-cap: "AUC comparison plot"
+#| echo: FALSE
+here()
+knitr::include_graphics(here("results","figures", "auc_plot.png"))
+```
 

### 2025-05-02 code/analysis-code/analysis.qmd
Part 6 updates

@@
 
-Next, we will make the multivariate model. I will use cross validation for this.Pooled AUC is calculated and a observed-expected plot is generated.
+Next, we will make the multivariate model. I will use cross validation for this. Pooled AUC is calculated and a observed-expected plot is generated.
 ```{r}
-library(pROC)
-library(dplyr)
-
 cross_validation_imputed_auc_logit <- function(imputed_data, predictors, k_folds = 5) {
@@
 ```{r}
-library(glmnet)
-library(caret)
-library(Metrics)  # for rmse()
-library(dplyr)
-
 set.seed(1234)
@@
 
+auc_list
 pooled_auc <- mean(auc_list)
@@
        width = 8, height = 6, dpi = 300)
-
-################################################################################
-# Keep variable names for predictors
-formula <- as.formula("COVID ~ .")
-
-# Explicitly remove rows with any NA in predictors (even if COVID is not NA)
-train_data <- train_data %>% select(COVID, everything()) %>% drop_na()
-
-# Now get model matrix
-x_train <- model.matrix(formula, data = train_data)[, -1]
-y_train <- train_data$COVID
-
-nrow(x_train)
-length(y_train)
-```
-
-```{r}
-# Load required libraries
-library(glmnet)      # For Lasso regression
-library(pROC)        # For AUC calculation
-library(ggplot2)     # For plotting
-library(dplyr)       # For data manipulation
-library(mice)        # For working with imputed data
-library(tidyr)       # For data reshaping
-
-# --- LASSO MODEL WITH CROSS-VALIDATION ON IMPUTED DATA ---
-
-# Function to run Lasso with sample weights and calculate performance metrics
-run_lasso_cv <- function(imp_data, outcome = "COVID", 
-                        predictors = c("SLEEP", "DEPRESSED", "SOCIAL", "TROUBLE_SLEEPING", "EXERCISE"),
-                        weight_var = "sampleweight") {
-
-  # Extract data from current imputation
-  X <- as.matrix(imp_data[, predictors])
-  y <- imp_data[[outcome]]
-  weights <- imp_data[[weight_var]]
-
-  # Standardize predictors (important for Lasso)
-  X_scaled <- scale(X)
-
-  # Run cross-validated Lasso with sample weights
-  set.seed(123) # For reproducibility
-  cv_model <- cv.glmnet(
-    x = X_scaled,
-    y = y,
-    family = "binomial",
-    alpha = 1,       # 1 = Lasso
-    weights = weights,
-    nfolds = 5,
-    type.measure = "auc"
-  )
-
-  # Get the best lambda (penalty parameter)
-  best_lambda <- cv_model$lambda.min
-
-  # Refit model with best lambda
-  final_model <- glmnet(
-    x = X_scaled,
-    y = y,
-    family = "binomial",
-    alpha = 1,
-    weights = weights,
-    lambda = best_lambda
-  )
-
-  # Get predictions
-  pred_probs <- predict(final_model, newx = X_scaled, type = "response")
-
-  # Calculate AUC
-  roc_obj <- roc(y, pred_probs)
-  auc_value <- auc(roc_obj)
-
-  # Create calibration data for expected vs observed plot
-  n_groups <- 10  # Number of bins for calibration plot
-
-  # Create calibration data
-  calib_data <- data.frame(
-    predicted = as.vector(pred_probs),
-    observed = y
-  )
-
-  # Sort by predicted probability and divide into bins
-  calib_data <- calib_data[order(calib_data$predicted), ]
-  calib_data$bin <- cut(seq(1, nrow(calib_data)), breaks = n_groups, labels = FALSE)
-
-  # Calculate mean predicted and observed values for each bin
-  calibration <- calib_data %>%
-    group_by(bin) %>%
-    summarize(
-      n = n(),
-      mean_predicted = mean(predicted),
-      mean_observed = mean(observed)
-    )
-
-  # Return results
-  return(list(
-    model = final_model,
-    lambda = best_lambda,
-    auc = auc_value,
-    calibration = calibration,
-    coefs = coef(final_model)
-  ))
-}
-
-# ----------------------------
-# MAIN ANALYSIS
-# ----------------------------
-
-# Check the structure of the imputed data
-print(class(data_imputed))
-
-# Initialize results containers
-auc_values <- numeric(length = 5)
-calibration_results <- list()
-coef_results <- list()
-
-# Function to run Lasso with sample weights and calculate performance metrics
-run_lasso_cv <- function(imp_data, outcome = "COVID", 
-                        predictors = c("SLEEP", "DEPRESSED", "SOCIAL", "TROUBLE_SLEEPING", "EXERCISE"),
-                        weight_var = "sampleweight") {
-
-  # Extract data from current imputation and handle missing values
-  # First, check for and handle missing data
-  complete_cases <- complete.cases(imp_data[, c(outcome, predictors, weight_var)])
-  if (!all(complete_cases)) {
-    message("Removing ", sum(!complete_cases), " rows with missing values")
-    imp_data <- imp_data[complete_cases, ]
-  }
-
-  X <- as.matrix(imp_data[, predictors])
-  y <- imp_data[[outcome]]
-  weights <- imp_data[[weight_var]]
-
-  # Standardize predictors (important for Lasso)
-  X_scaled <- scale(X)
-
-  # Run cross-validated Lasso with sample weights
-  set.seed(123) # For reproducibility
-  cv_model <- cv.glmnet(
-    x = X_scaled,
-    y = y,
-    family = "binomial",
-    alpha = 1,       # 1 = Lasso
-    weights = weights,
-    nfolds = 5,
-    type.measure = "auc"
-  )
-
-  # Get the best lambda (penalty parameter)
-  best_lambda <- cv_model$lambda.min
-
-  # Refit model with best lambda
-  final_model <- glmnet(
-    x = X_scaled,
-    y = y,
-    family = "binomial",
-    alpha = 1,
-    weights = weights,
-    lambda = best_lambda
-  )
-
-# Combine calibration results from all imputations
-all_calibration <- do.call(rbind, calibration_results)
-
-# ----------------------------
-# OUTPUT RESULTS
-# ----------------------------
-
-# Print AUC values for each imputation
-cat("\nAUC values for each imputation:\n")
-print(auc_values)
-cat("\nMean AUC across imputations:", mean(auc_values), "\n")
-
-# Plot expected vs observed probabilities for all imputations
-calibration_plot <- ggplot(all_calibration, aes(x = mean_predicted, y = mean_observed, color = factor(imputation))) +
-  geom_point(size = 3) +
-  geom_line() +
-  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
-  labs(
-    title = "Expected vs. Observed Probabilities",
-    subtitle = paste("Mean AUC =", round(mean(auc_values), 3)),
-    x = "Expected (Predicted Probability)",
-    y = "Observed (Actual Proportion)",
-    color = "Imputation"
-  ) +
-  theme_minimal() +
-  coord_equal() +
-  xlim(0, 1) + 
-  ylim(0, 1)
-
-print(calibration_plot)
-
-# Print average coefficients across imputations
-cat("\nAverage Lasso coefficients across imputations:\n")
-# Extract coefficient names from the first imputation
-coef_names <- rownames(coef_results[[1]])
-
-# Create matrix to store coefficients from all imputations
-all_coefs <- matrix(0, nrow = length(coef_names), ncol = 5)
-rownames(all_coefs) <- coef_names
-
-# Fill in the coefficient matrix
-for (i in 1:5) {
-  all_coefs[, i] <- as.vector(coef_results[[i]])
-}
-
-# Calculate mean coefficients
-mean_coefs <- rowMeans(all_coefs)
-coef_summary <- data.frame(
-  Variable = coef_names,
-  Mean_Coefficient = mean_coefs
-)
-
-# Print non-zero coefficients
-non_zero_coefs <- coef_summary[abs(coef_summary$Mean_Coefficient) > 0.001, ]
-print(non_zero_coefs[order(abs(non_zero_coefs$Mean_Coefficient), decreasing = TRUE), ])
-```
-
-```{r}
-library(dplyr)
-library(ggplot2)
-library(scales)
-
-calib_data_list <- vector("list", length(data_imputed))
-
-for (i in seq_along(data_imputed)) {
-  dat <- data_imputed[[i]] %>%
-    drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))
-
-  X <- model.matrix(as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))), dat)[, -1]
-  y <- dat[[outcome_var]]
-  w <- dat$SAMPWEIGHT
-
-  set.seed(1234 + i)
-  train_idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
-  test_idx <- setdiff(seq_len(nrow(X)), train_idx)
-
-  X_train <- X[train_idx, ]
-  y_train <- y[train_idx]
-  w_train <- w[train_idx]
-
-  X_test <- X[test_idx, ]
-  y_test <- y[test_idx]
-
-  cv_fit <- cv.glmnet(
-    x = X_train,
-    y = y_train,
-    weights = w_train,
-    alpha = 1,
-    family = "binomial",
-    standardize = TRUE
-  )
-
-  pred_probs <- predict(cv_fit, newx = X_test, s = "lambda.min", type = "response") %>% as.vector()
-
-  # Count unique predicted probabilities
-  n_unique_preds <- length(unique(pred_probs))
-
-  # Only proceed with binning if there are at least 2 unique predicted probabilities
-  if (n_unique_preds >= 2) {
-    # Safely bin the predicted probabilities
-    n_bins <- min(5, n_unique_preds)
-
-    calib_data <- tibble(
-      actual = y_test,
-      predicted = pred_probs
-    ) %>%
-      mutate(bin = cut_number(predicted, n = n_bins, labels = FALSE)) %>%
-      group_by(bin) %>%
-      summarise(
-        mean_pred = mean(predicted),
-        obs_rate = mean(actual),
-        .groups = "drop"
-      ) %>%
-      mutate(imputation = paste0("Imputation ", i))
-
-    calib_data_list[[i]] <- calib_data
-  } else {
-    message("Skipping calibration plot for imputation ", i, ": too few unique predicted values (", n_unique_preds, " unique values)")
-  }
-}
-
-# Combine all calibration data
-all_calib_data <- bind_rows(calib_data_list)
-
-# Create calibration plot
-ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
-  geom_line() +
-  geom_point(size = 2) +
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
-  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
-  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
-  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
-  theme_minimal()
-
-# Combine calibration data
-all_calib_data <- bind_rows(calib_data_list)
-
-# Plot calibration curve
-ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
-  geom_line() +
-  geom_point(size = 2) +
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
-  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
-  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
-  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
-  theme_minimal()
-
-# Combine calibration data
-all_calib_data <- bind_rows(calib_data_list)
-
-# Plot calibration curve
-ggplot(all_calib_data, aes(x = mean_pred, y = obs_rate, color = imputation)) +
-  geom_line() +
-  geom_point(size = 2) +
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
-  scale_x_continuous("Expected Probability", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
-  scale_y_continuous("Observed Proportion", limits = c(0, 1), labels = percent_format(accuracy = 1)) +
-  labs(title = "Calibration Plot: Lasso Logistic Regression", color = "Imputation") +
-  theme_minimal()
-# Save the plot
-ggsave(filename = here::here("results", "figures", "lasso_calibration_plot.png"),
-       plot = calibration_plot_lasso,
-       width = 8, height = 6, dpi = 300)
-
-# Initialize list to store AUCs
-auc_list <- vector("numeric", length(data_imputed))
-
-for (i in seq_along(data_imputed)) {
-  dat <- data_imputed[[i]] %>%
-    drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))
-
-  X <- model.matrix(as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))), dat)[, -1]
-  y <- dat[[outcome_var]]
-  w <- dat$SAMPWEIGHT
-
-  set.seed(1234 + i)
-  train_idx <- sample(seq_len(nrow(X)), size = 0.7 * nrow(X))
-  test_idx <- setdiff(seq_len(X), train_idx)
-
-  X_train <- X[train_idx, ]
-  y_train <- y[train_idx]
-  w_train <- w[train_idx]
-
-  X_test <- X[test_idx, ]
-  y_test <- y[test_idx]
-
-  # Fit lasso model
-  cv_fit <- cv.glmnet(
-    x = X_train,
-    y = y_train,
-    weights = w_train,
-    alpha = 1,
-    family = "binomial",
-    standardize = TRUE
-  )
-
-  # Predict probabilities
-  pred_probs <- predict(cv_fit, newx = X_test, s = "lambda.min", type = "response") %>% as.vector()
-
-  # Compute AUC
-  auc_val <- roc(y_test, pred_probs)$auc
-  auc_list[i] <- auc_val
-}
-
-# Print AUCs
-cat("AUC for each imputation:\n")
-print(round(auc_list, 3))
-
-# Mean (pooled) AUC
-cat("\nPooled AUC across imputations:", round(mean(auc_list), 3), "\n")
 ```
@@
 # Assuming you have your imputed_data and a vector of predictors
-results_with_preds <- cross_validation_imputed_predictions(
+results_with_preds2 <- cross_validation_imputed_predictions(
   imputed_data = data_imputed,
@@
 # View results
-results$auc_per_imputation  # AUC for each imputed dataset
-results$overall_auc         # Average AUC across imputations
+results_with_preds2$auc_per_imputation  # AUC for each imputed dataset
+results_with_preds2$overall_auc         # Average AUC across imputations
 
@@
 
-Now, we will make expected versus observed plots.
+Now, we will make a plot of the AUCs.
 ```{r}
-plot_calibration_multiple <- function(actual, predicted_list, n_bins = 10, title = "Calibration Plot") {
-  # Combine calibration data across imputations
-  calib_data_all <- map2_dfr(predicted_list, seq_along(predicted_list), function(predicted, i) {
-    tibble(actual = actual, predicted = predicted) %>%
-      mutate(bin = ntile(predicted, n_bins)) %>%
-      group_by(bin) %>%
-      summarise(
-        mean_pred = mean(predicted),
-        obs_rate = mean(actual),
-        .groups = "drop"
-      ) %>%
-      mutate(imputation = paste0("Imputation ", i))
-  })
-
-  # Plot
-  ggplot(calib_data_all, aes(x = mean_pred, y = obs_rate, color = imputation)) +
-    geom_line() +
-    geom_point(size = 2) +
-    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
-    scale_x_continuous("Expected probability", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
-    scale_y_continuous("Observed proportion", limits = c(0, 1), labels = scales::percent_format(accuracy = 1)) +
-    labs(title = title, color = "Imputation") +
-    theme_minimal()
-}
-
-# Example usage:
-plot_cal <- plot_calibration_multiple(
-  actual = test_data$COVID,
-  predicted_list = pred_probs_list,  # this should be a list of 5 vectors
-  title = "Random Forest Calibration Across 5 Imputations"
+# Step 1: Combine AUCs into a tidy data frame
+auc_df <- data.frame(
+  model = rep(c("Logistic Regression", "Lasso", "Random Forest"), each = 5),
+  imputation = rep(1:5, times = 3),
+  AUC = c(
+    results_logit$auc_per_imputation,
+    auc_list,
+    results_with_preds2$auc_per_imputation
+  )
 )
 
-# Display plot
-print(plot_cal)
+# Step 2: Calculate mean AUC per model
+auc_means <- auc_df %>%
+  group_by(model) %>%
+  summarise(mean_auc = mean(AUC), .groups = "drop")
 
-# Save it
+# Step 3: Plot
+ggplot(auc_df, aes(x = model, y = AUC, color = model)) +
+  geom_jitter(width = 0.15, size = 3, alpha = 0.8) +  # dots for each imputation
+  geom_point(data = auc_means, aes(x = model, y = mean_auc),
+             color = "black", size = 4) +             # black dot for mean
+  geom_text(data = auc_means, aes(x = model, y = mean_auc, 
+             label = formatC(mean_auc, format = "f", digits = 4)),
+            vjust = -1.2, color = "black", size = 4.5) +  # label for mean
+  labs(
+    title = "AUCs by Model Type Across 5 Imputations",
+    x = "Model",
+    y = "AUC"
+  ) +
+  theme_minimal(base_size = 14) +
+  scale_y_continuous(limits = c(0.45, 0.6), breaks = seq(0.45, 0.6, 0.05)) +
+  scale_color_brewer(palette = "Dark2") +
+  theme(legend.position = "none")
+
+#save plot
 ggsave(
-  filename = here("results", "figures", "rf_calibration_all_imputations.png"),
-  plot = plot_cal,
-  width = 8,
-  height = 6,
-  dpi = 300
+  filename = here("results", "figures", "auc_comparison_plot.png"),  # save in outputs/ directory
+  plot = last_plot(),       # or specify your plot object explicitly
+  width = 8, height = 6,    # size in inches
+  dpi = 300                 # high-quality resolution
 )
-
 ```

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 updates

@@
 Higher AUC indicates better model preformance.
+Desspite the Logistic Regression model and the Lasso Regression model have the same mean AUC, they are not the same, shown by the differing values for the imputations.
+Code was checked to ensure the models were not the same.
 
@@
 here()
-knitr::include_graphics(here("results","figures", "auc_plot.png"))
+knitr::include_graphics(here("results","figures", "auc_comparison_plot.png"))
+```
+
+Observed vs.
+expected plots for the three models compare the predictive power of the models from the cross validated data.
+Training data was used for expectations and observed data was drawn from the testing data.
+Plots evaluate the models' modest predictive power on COVID-19 test outcomes from mental health variables.
+
+```{r}
+#| label: fig-log-plot
+#| fig-cap: "Logistic Regression calibration plot"
+#| echo: FALSE
+here()
+knitr::include_graphics(here("results","figures", "log_calibration_plot.png"))
+```
+
+```{r}
+#| label: fig-lasso-plot
+#| fig-cap: "LASSO Regression calibration plot"
+#| echo: FALSE
+here()
+knitr::include_graphics(here("results","figures", "lasso_calibration_plot.png"))
+```
+
+```{r}
+#| label: fig-rf-plot
+#| fig-cap: "Random Forest calibration plot"
+#| echo: FALSE
+here()
+knitr::include_graphics(here("results","figures", "rf_calibration_all_imputations.png"))
 ```

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 updates

@@
 ---
-title: "Mental Health Factors and Covid"
+title: "Association of Mental Health Factors with COVID-19 Test Status"
 author: "Annalise Cramer"
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
+The exploratory analysis revealed prevalence of mental health factors among the population by COVID-19 test status.
+Interestingly, individuals who reported receiving less social support occured at a higher rate for those who did not test positive for COVID-19 (0.10 (0.10, 0.10)) than those who did test positive for COVID-19 (0.08 (0.07, 0.09)).
+A difference in prevalence between COVID-19 test status was not noted for individuals who reported no change/more social support.
+
+Individuals who reported recent trouble sleeping were occured at a higher prevalence for those who tested postive for COVID-19 (0.374 (0.36, 0.39)) than those who did not recieve a positive COVID-19 test  (0.345 (0.34, 0.35)).
+Individuals who did not report recent trouble sleeping showed a difference by COVID-19 test status, with less testing positive for COVID-19 (0.626 (0.61, 0.64)) and more not testing positive for COVID-19 (0.655 (0.65, 0.66)).
+
+Model preformance was evaluated via the AUC matric.
+The logistic regression model had a pooled AUC of 0.5031.
+The LASSO regression model showed a pooled AUC of 0.5301.
+The random forest model showed the highest pooled AUC of 0.5474.
+AUC values for all models are fairly close to 0.50, with only small improvement with the randome forest model.
+This indicates the models have little ability to discriminate between a postive or negative COVID-19 test based on mental health factors, with the random forest model being the best.
+
+Plots for observed vs.
+expected values provide a secondary assessment of model preformance, which show how closesly the training data matched the testing data.
+Plots for well calibrated models should show the points for imputations falling along the 45 degree dotted line, which is not present in the logistic regression or LASSO regression models.
+The random forest model shows some closeness to the 45 degree line, confirming it to be the best model.
+However, no model shows good predicitive power, given there is minimal closeness to the line.
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
+This study used data collected from the National Health Information Surveys, a well established and thorough survey program run by the CDC.
+NHIS uses sampling methods and covers many health-related topics, allowing it to be generalizable to the United States population.
+Additionally, this study uses multiple imputation to fill missing data points, which reduces the introduciton of potential bias from discarding missings or from single imputation.
+Applying cross validation to the three full models allows for analysis of predictive power, and the use of the LASSO regression allows for penalization of unecessary predictors, and the use of the random forest models allows for potential non-linear relationships.
+
+Limitations of this study include the nature of survey data.
+Because individuals self-reported this data, it is possible that recall bias may influence accuracy.
+Associations between predictors and COVID test status cannot be established as causal.
+Without inclusions of full survey design variables, the three full models are not generalizable to the United States population.
+Finally, all three models showed moderate to low predicitive power.
 

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 updates

@@
 ## Conclusions
-_What are the main take-home messages?_
+We found that indications of poor mental health were common in the population, as INSERT HERE.
+Differences between individuals who tested positive for COVID-19 and who didn't test positive for COVID-19 were found among those who indicated recieveing less social support and those who reported trouble sleeping.
+However, these differences were relatively small.
+
+Predicitve modeling with logistic regression, LASSO regression, and random forest models showed weak preformance and low predicitve power, indicating mental health factors show little indication of COVID-19 test status.
+
+Despite the lack of strong results, these findings show the complications in understanding how mental health and physical health are related.
+Little consideration was given to the impacts of societal measures for disease control at the start of the SARS-CoV-2 pandemic, and it's proven difficult to find such connections in the data.
+Future work may explore these topics further by using longitudinal data to track changes in mental health over the duration of COVID-19 infection.
+Future surveys may differentiate the experinces of various disease control measures (quarantining, isolation, masking, etc.) or mental health conditions (feelings of anxiety, educational impacts) to further specify the potential impacts on mental health.
+Finally, including complete survey design in the creation of the full models would offer more generalizable results towards the United States population.
+
+1.) What is the prevalence of poor mental health?
+2.) Is poor mental health associated with a positive covid test?
 

### 2025-05-02 code/analysis-code/analysis.qmd
Part 6 updates

@@
 
-We create estimates and CIs using complex survery design variables, using the multiple imputed data.
-
+We create estimates and CIs using complex survery design variables, using the multiple imputed data. This will find estimates for overall, and by COVID-19 test status.
 ```{r}
-
 # Create survey design objects for each transformed dataset
@@
 
-# Function to get complete estimates (both 0 and 1) stratified by COVID status
 get_complete_stratified_estimates <- function(var_name) {
@@
 
-  # Loop through COVID status (0 and 1)
-  for (covid_val in c(0, 1)) {
+  # Loop through COVID status: Negative (0), Positive (1), Overall (NA)
+  for (covid_val in c(0, 1, NA)) {
+    if (is.na(covid_val)) {
+      covid_label <- "Overall"
+      subset_designs <- svy_designs  # No subsetting
+    } else {
       covid_label <- ifelse(covid_val == 0, "Negative", "Positive")
+      subset_designs <- lapply(svy_designs, function(design) subset(design, COVID == covid_val))
+    }
 
-    # Calculate proportions for this COVID status
-    props <- lapply(svy_designs, function(design) {
-      # Create a subset design
-      subdesign <- subset(design, COVID == covid_val)
-      # Create formula for the variable
+    # Calculate proportions for this COVID status or overall
+    props <- lapply(subset_designs, function(design) {
       formula <- as.formula(paste0("~factor(", var_name, ")"))
-      # Calculate proportions
-      svymean(formula, subdesign, na.rm = TRUE)
+      svymean(formula, design, na.rm = TRUE)
     })
@@
     pooled <- MIcombine(props)
-
-    # Extract results for both binary values (if available)
     coef_names <- names(coef(pooled))
 
-    # Handle different potential naming patterns
     for (bin_val in c(0, 1)) {
-      # Check different possible coefficient names
       possible_names <- c(
@@
       )
-
-      # Find the matching coefficient name
       coef_idx <- NULL
@@
 
-      # If found, extract estimates
       if (!is.null(coef_idx)) {
@@
 
-        # Add to results
         results <- rbind(results, data.frame(
@@
 }
-
-# Variables to analyze
 variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")
 
-# Get complete stratified estimates for each variable
+# Get complete stratified + overall estimates for each variable
 complete_results <- lapply(variables, get_complete_stratified_estimates)
@@
 print(formatted_results)
-
 ```
@@
 
-# Prepare the data by combining Variable and Binary_Label, pivoting, and formatting the results
 table1_imputed <- formatted_results %>%
-  mutate(Binary_Label = ifelse(Binary_Label == "Yes", 1, 0),  # Convert Yes to 1 and No to 0
+  mutate(Binary_Label = ifelse(Binary_Label == "Yes", 1, 0),
          Variable_Label = paste(Variable, Binary_Label, sep = " = ")) %>%
@@
   mutate(
-    # Scale the estimates and CI by 0.01
     Estimate_Negative = Negative_Estimate * 0.01,
@@
     Estimate_Positive = Positive_Estimate * 0.01,
-    CI_Positive = sprintf("(%.2f, %.2f)", Positive_CI_lower * 0.01, Positive_CI_upper * 0.01)
+    CI_Positive = sprintf("(%.2f, %.2f)", Positive_CI_lower * 0.01, Positive_CI_upper * 0.01),
+    Estimate_Overall = Overall_Estimate * 0.01,
+    CI_Overall = sprintf("(%.2f, %.2f)", Overall_CI_lower * 0.01, Overall_CI_upper * 0.01)
   ) %>%
-  select(Variable_Label, Estimate_Negative, CI_Negative, Estimate_Positive, CI_Positive) %>%
+  select(Variable_Label, Estimate_Negative, CI_Negative,
+         Estimate_Positive, CI_Positive, Estimate_Overall, CI_Overall) %>%
   gt() %>%
@@
   ) %>%
+  tab_spanner(
+    label = "Overall",
+    columns = c(Estimate_Overall, CI_Overall)
+  ) %>%
   cols_label(
-    Variable_Label = "Variable",               # Rename Variable_Label to Variable
-    Estimate_Negative = "Estimate",            # Rename Estimate_Negative to Estimate
-    CI_Negative = "95% CI",                    # Rename CI_Negative to 95% CI
-    Estimate_Positive = "Estimate",            # Rename Estimate_Positive to Estimate
-    CI_Positive = "95% CI"                     # Rename CI_Positive to 95% CI
+    Variable_Label = "Variable",
+    Estimate_Negative = "Estimate",
+    CI_Negative = "95% CI",
+    Estimate_Positive = "Estimate",
+    CI_Positive = "95% CI",
+    Estimate_Overall = "Estimate",
+    CI_Overall = "95% CI"
   )
+
+print(table1_imputed)
 

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 updates

@@
 ## Conclusions
-We found that indications of poor mental health were common in the population, as INSERT HERE.
-Differences between individuals who tested positive for COVID-19 and who didn't test positive for COVID-19 were found among those who indicated recieveing less social support and those who reported trouble sleeping.
+We found that indications of poor mental health were common in the population, as ######%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%INSERT HERE.
+YOU NEED TO MAKE THIS TABLE FOR NON SPLIT BY COVID TEST AND WRITE RESULTS IN.
+ALSO YOU NEED TO PUT IN CITAIONS Differences between individuals who tested positive for COVID-19 and who didn't test positive for COVID-19 were found among those who indicated recieveing less social support and those who reported trouble sleeping.
 However, these differences were relatively small.
 
-Predicitve modeling with logistic regression, LASSO regression, and random forest models showed weak preformance and low predicitve power, indicating mental health factors show little indication of COVID-19 test status.
+Predictive modeling with logistic regression, LASSO regression, and random forest models showed weak preformance and low predicitve power, indicating mental health factors show little indication of COVID-19 test status.
 

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 udpates

@@
 #| label: fig-schematic
-#| fig-cap: "Flowchart depicting methodology"
+#| fig-cap: "Figure 1. Flowchart depicting methodology of this study."
 #| echo: FALSE
@@
 #| label: tab-1
-#| fig-cap: "Missingness table"
+#| fig-cap: "Table 1. Table depicting porportion of responses among all observations, incliding missingness."
 #| echo: FALSE
@@
 ```{r}
-#| label: fig-1
-#| fig-cap: "Summary statistics table"
+#| label: tab-2
+#| fig-cap: "Table 2.
+Table depicting prevalence estimates with 95% confidence intervals among the United State population.
+Results are split by COVID-19 test status, and shown for overall population."
 #| echo: FALSE
@@
 #| label: tbl-summarytable
-#| tbl-cap: "Single model table"
+#| tbl-cap: "Table 3. Assocation of mental health factors with positive COVID-19 test results among the United State population."
 #| echo: FALSE
@@
 #| label: fig-aucplot
-#| fig-cap: "AUC comparison plot"
+#| fig-cap: "Figure 2.
+Plot comparing area under the curve (AUC) metrics among imputations between logistic regression model, LASSO regression model, and random forest model.
+Higher AUC indicates better preformance.
+"
 #| echo: FALSE
@@
 #| label: fig-log-plot
-#| fig-cap: "Logistic Regression calibration plot"
+#| fig-cap: "Figure 3. Logistic Regression calibration plot."
 #| echo: FALSE
@@
 #| label: fig-lasso-plot
-#| fig-cap: "LASSO Regression calibration plot"
+#| fig-cap: "Figure 4. LASSO Regression calibration plot."
 #| echo: FALSE
@@
 #| label: fig-rf-plot
-#| fig-cap: "Random Forest calibration plot"
+#| fig-cap: "Figure 5. Random Forest calibration plot. "
 #| echo: FALSE
@@
 ## Summary and Interpretation
+The exploratory analysis revelaed that feeling depressed occurs a prevalence of 0.095 (0.09, 0.10).
+meeting exercise requiremnts occurs at a prevalence of  0.526 (0.52, 0.53), getting enough sleep occurs at a prevalence of 0.494 (0.49, 0.50), recieving less social support occurs at a prevalence of 0.097 (0.09, 0.10), and having trouble sleeping occurs at a prevalence of 0.630 (0.62, 0.64) (Tab.
+2).
+Several of these factors occur in the majority of the population, that is, trouble sleeping and exercise, and getting enough sleep occured in almost half the population.
+Feeling depressed and recieving less socual support occured in a small fraction of the population (Tab.
+2).
+
 The exploratory analysis revealed prevalence of mental health factors among the population by COVID-19 test status.
-Interestingly, individuals who reported receiving less social support occured at a higher rate for those who did not test positive for COVID-19 (0.10 (0.10, 0.10)) than those who did test positive for COVID-19 (0.08 (0.07, 0.09)).
+Interestingly, individuals who reported receiving less social support occured at a higher rate for those who did not test positive for COVID-19 (0.10 (0.10, 0.10)) than those who did test positive for COVID-19 (0.08 (0.07, 0.09)) (Tab.
+2).
 A difference in prevalence between COVID-19 test status was not noted for individuals who reported no change/more social support.
 
+Single models quantified association of each mental health factor with a positive COVID-19 test.
+Feeling depressed, and getting exercise were not significantly associated witha positive COVID-19 test (Tab.
+3).
+Getting enough sleep was found to slightly increase the odds of a positive COVID-19 test (OR = 1.076 (1.026–1.13), p = 0.003).
+Getting less social support was found to slightly decrease the odds of a positive COVID-19 test (OR = 0.777, (0.691, 0.875), p < 0.0001).
+This may indicate effectiveness of societal control strategies, that is, social distancing, quarantine, and isolation in preventing spread of COVID-19.
+Reporting trouble sleeping was found to slightly increase the odds of a positive COVID-19 test (OR = 1.134 (1.067, 1.206), p < 0.0001), which slightly disagrees with results of the SLEEP vairbale described earlier (Tab.
+3).
+
 Individuals who reported recent trouble sleeping were occured at a higher prevalence for those who tested postive for COVID-19 (0.374 (0.36, 0.39)) than those who did not recieve a positive COVID-19 test  (0.345 (0.34, 0.35)).
-Individuals who did not report recent trouble sleeping showed a difference by COVID-19 test status, with less testing positive for COVID-19 (0.626 (0.61, 0.64)) and more not testing positive for COVID-19 (0.655 (0.65, 0.66)).
+Individuals who did not report recent trouble sleeping showed a difference by COVID-19 test status, with less testing positive for COVID-19 (0.626 (0.61, 0.64)) and more not testing positive for COVID-19 (0.655 (0.65, 0.66)) (Tab.
+2).
 
-Model preformance was evaluated via the AUC matric.
+Model preformance was evaluated via the AUC metric.
 The logistic regression model had a pooled AUC of 0.5031.
@@
 The random forest model showed the highest pooled AUC of 0.5474.
-AUC values for all models are fairly close to 0.50, with only small improvement with the randome forest model.
+AUC values for all models are fairly close to 0.50, with only small improvement with the randome forest model (Fig.
+2).
 This indicates the models have little ability to discriminate between a postive or negative COVID-19 test based on mental health factors, with the random forest model being the best.
@@
 expected values provide a secondary assessment of model preformance, which show how closesly the training data matched the testing data.
-Plots for well calibrated models should show the points for imputations falling along the 45 degree dotted line, which is not present in the logistic regression or LASSO regression models.
-The random forest model shows some closeness to the 45 degree line, confirming it to be the best model.
+Plots for well calibrated models should show the points for imputations falling along the 45 degree dotted line, which is not present in the logistic regression (Fig.
+3) or LASSO regression models (Fig.
+4).
+The random forest model shows some closeness to the 45 degree line, confirming it to be the best model (Fig.
+5).
 However, no model shows good predicitive power, given there is minimal closeness to the line.
@@
 ## Conclusions
-We found that indications of poor mental health were common in the population, as ######%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%INSERT HERE.
+We found that indications of poor mental health were common in the population.
+Most common was  ######%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%INSERT HERE.
 YOU NEED TO MAKE THIS TABLE FOR NON SPLIT BY COVID TEST AND WRITE RESULTS IN.

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 updates

@@
 1.) What is the prevalence of poor mental health?
-2.) Is poor mental health associated with a positive covid test?
+2.) Is poor mental health associated with a positive COVID-19 test?
+3.) Can mental health facotrs be used to predict if an individual will recieve a positive COVID-19 test?
 
@@
 The exploratory analysis revelaed that feeling depressed occurs a prevalence of 0.095 (0.09, 0.10).
-meeting exercise requiremnts occurs at a prevalence of  0.526 (0.52, 0.53), getting enough sleep occurs at a prevalence of 0.494 (0.49, 0.50), recieving less social support occurs at a prevalence of 0.097 (0.09, 0.10), and having trouble sleeping occurs at a prevalence of 0.630 (0.62, 0.64) (Tab.
+meeting exercise requirements occurs at a prevalence of  0.526 (0.52, 0.53), getting enough sleep occurs at a prevalence of 0.494 (0.49, 0.50), recieving less social support occurs at a prevalence of 0.097 (0.09, 0.10), and having trouble sleeping occurs at a prevalence of 0.630 (0.62, 0.64) (Tab.
 2).
@@
 ## Conclusions
-We found that indications of poor mental health were common in the population.
-Most common was  ######%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%INSERT HERE.
-YOU NEED TO MAKE THIS TABLE FOR NON SPLIT BY COVID TEST AND WRITE RESULTS IN.
-ALSO YOU NEED TO PUT IN CITAIONS Differences between individuals who tested positive for COVID-19 and who didn't test positive for COVID-19 were found among those who indicated recieveing less social support and those who reported trouble sleeping.
-However, these differences were relatively small.
+We found that indications of poor mental health were common in the population United States population, with trouble sleeping, exercise, and getting enough sleep, while depressed and social isolation occured in at a lower prevalence (Tab.
+2).
+Sleep difficulties were commonly reported during the SARS-CoV-2 pandemic, which supports these results.
+CITE ME %%%%%%%%
 
-Predictive modeling with logistic regression, LASSO regression, and random forest models showed weak preformance and low predicitve power, indicating mental health factors show little indication of COVID-19 test status.
+Social support was the only mental health variable to vary among the population by COVID-19 test status (Tab.
+2).
+Interestingly, recieving less social support was found to be protective against SARS-CoV-2 (Tab.
+3).
+This is likely due to isolation and quarantining practices that occured during thee years 2020-2022.
+CITE ME %%%%%
 
-Despite the lack of strong results, these findings show the complications in understanding how mental health and physical health are related.
-Little consideration was given to the impacts of societal measures for disease control at the start of the SARS-CoV-2 pandemic, and it's proven difficult to find such connections in the data.
-Future work may explore these topics further by using longitudinal data to track changes in mental health over the duration of COVID-19 infection.
-Future surveys may differentiate the experinces of various disease control measures (quarantining, isolation, masking, etc.) or mental health conditions (feelings of anxiety, educational impacts) to further specify the potential impacts on mental health.
-Finally, including complete survey design in the creation of the full models would offer more generalizable results towards the United States population.
-
-1.) What is the prevalence of poor mental health?
-2.) Is poor mental health associated with a positive covid test?
+Predictive models showed weak preformancce and poor predicitve power (Figs. 2, 3). The relationship between SARS-CoV-2 and mental health continues to be challenging to understand and interpret.
 

### 2025-05-02 products/manuscript/Manuscript.qmd
fix citations

@@
 Social distancing, quantining, isolation, and lockdowns were prioritized control of transmission, but societal impacts were not initially considered.
-Disruptions to daily routine and social interactions quickly became a widespread concerns for their potential impacts on mental health, with widespread increases in anxiety, depression, loneliness, and sleep disturbances (CITATION NEEDED).
+Disruptions to daily routine and social interactions quickly became a widespread concerns for their potential impacts on mental health, with widespread increases in anxiety, depression, loneliness, and sleep disturbances$^1$.
 
 Pandemic control strategies have become commonly criticized for the impacts on individuals' psychological well-being.
-Research documents the mental health consequences of pandemic control strageies (CITATION SAME AS BEFORE), but the potential relationships between mental health factors and vulnerability to SARS-CoV-2 infection is less understood.
+Research documents the mental health consequences of pandemic control strageies$^1$, but the potential relationships between mental health factors and vulnerability to SARS-CoV-2 infection is less understood.
 
 This study seeks to add to the limited understanding of mental health and COVID-19 vulnerability by expliorung associations between mental health factors and COVID-19 test positivity status.
-Self reported bouts of depressed mood, perceived social support, amount of exercise, and sleep quality (via hours of sleep and reporting difficulty falling asleep) are examined.
-(CITATION HERE) Simple and multiple regression models including Lasso and Random Forest are used to explore the predicitve power of mental health factors to a positive COVID-19 test.
-
-Cite this properly later:
-1.) https://www.tandfonline.com/doi/full/10.1080/09638237.2020.1757052
+Self reported bouts of depressed mood, perceived social support, amount of exercise, and sleep quality (via hours of sleep and reporting difficulty falling asleep) are examined$^2$.Simple and multiple regression models including Lasso and Random Forest are used to explore the predicitve power of mental health factors to a positive COVID-19 test.
 
@@
 Data for this project is survey data collected from the National Health Interview Survey, produced by IPUMS through the University of Minnesota.
-This data was collected among United States residents of all ages and genders from 2019 - 2022 in the United States.
+This data was collected among United States residents of all ages and genders from 2020 - 2022 in the United States$^2$.
 
@@
 
-EXERCISE is a combination of number of minutes of passive or aerboic exercise, descirbing whether weekly amount meets the American Heart Association's guidelines for exercise needed to stay healthy.
+EXERCISE is a combination of number of minutes of passive or aerboic exercise, descirbing whether weekly amount meets the American Heart Association's guidelines (based on Office of Disease Prevention and Health Promotion recommendations)$^3$ for exercise needed to stay healthy.
 For adults, this includes 150 minutes per week of moderate intensity aerobic activity, 75 minutes per week of vigorous aerobic activity, or a combination of both.Children ages 6-17 should get 60 minutes of moderate to vigorous intenisty physical activity.
@@
 
-SLEEP is based on the NIH guidelines for recommended hours of sleep per day, which differs by age.
+SLEEP is based on the NIH guidelines for recommended hours of sleep per day, which differs by age$^4$.
 Individuals under age 1 are suggested to get 12-16 hours of sleep per day, children ages 1 to 2 are suggested to get 11-14 hours of sleep per day, children ages 3 to 5 years are suggested to get 10-13 hours of sleep per day, children ages 6 to 12 years are suggested to get 9-12 hours of sleep per day, teens ages 13-18 are suggested to get 8-10 hours of sleep per day, and adults (over the age of 18) are suggested to get 7-9 hours of sleep per day.
@@
 2).
-Sleep difficulties were commonly reported during the SARS-CoV-2 pandemic, which supports these results.
-CITE ME %%%%%%%%
+Sleep difficulties were commonly reported during the SARS-CoV-2 pandemic, which supports these results^$5$.
 
@@
 3).
-This is likely due to isolation and quarantining practices that occured during thee years 2020-2022.
-CITE ME %%%%%
+This is likely due to isolation and quarantining practices that occured during thee years 2020-2022$^6$.
 
-Predictive models showed weak preformancce and poor predicitve power (Figs. 2, 3). The relationship between SARS-CoV-2 and mental health continues to be challenging to understand and interpret.
-
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
+Predictive models showed weak preformancce and poor predictive power (Figs. 2, 3). The relationship between SARS-CoV-2 and mental health continues to be challenging to understand and interpret$^7$.
 
@@
 # References
+
+1.	Kumar A, Nayar KR. COVID-19 and its mental health consequences. J Ment Health. 2020;30(1):1–2. https://doi.org/10.1080/09638237.2020.1757052
+2.	Blewett LA, Rivera Drew JA, King ML, Williams KCW, Backman D, Chen A, et al.
+IPUMS Health Surveys: National Health Interview Survey, Version 7.4 [dataset].
+Minneapolis, MN: IPUMS; 2024.
+https://doi.org/10.18128/D070.V7.4
+3.	2018 Physical Activity Guidelines Advisory Committee. 2018 Physical Activity Guidelines Advisory Committee Scientific Report. Washington, DC: U.S. Department of Health and Human Services; 2018.
+4.	U.S. Department of Health and Human Services. 2005 Guide to Healthy Sleep. Maryland: U.S. Department of Health and Human Services; 2011.
+5.	Gupta R, Pandi-Perumal SR. COVID-Somnia: How the Pandemic Affects Sleep/Wake Regulation and How to Deal with it?. Sleep Vigil. 2020;4:51–53. https://doi.org/10.1007/s41782-020-00118-0
+6.	Szkody E, Stearns M, Stanhope L, McKinney C. Stress-buffering role of social support during COVID-19. Fam Process. 2021;60(3):1002–1015. https://doi.org/10.1111/famp.12618
+7.	Rajkumar RP. COVID-19 and mental health: A review of the existing literature. Asian J Psychiatr. 2020;52:102066. https://doi.org/10.1016/j.ajp.2020.102066

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 final

@@
 Infected individuals and healthy individuals were affected, but little is known about the potential relationships between mental health and physical health.
-In this study, we utilized Nataional Health Information Surveys from 2019-2021 to examine the relationships between receieving a positive COVID-19 test and self reporting mental health factors such as meeting recommended guidelines for exercise, meeting the recommended number of hours of sleep per night, reporting feeling depressed recently, reporting having trouble sleeping recently, and reporting receiving less social support.
+In this study, we utilized Nataional Health Information Surveys from 2020-2022 to examine the relationships between receieving a positive COVID-19 test and self reporting mental health factors such as meeting recommended guidelines for exercise, meeting the recommended number of hours of sleep per night, reporting feeling depressed recently, reporting having trouble sleeping recently, and reporting receiving less social support.
 
-We find
+Analysis is completed using complex survey design to find prevalence of mental health factors.
+Single models are used to determine associations, and logistic regression, LASSO regression, and a random forest model are explored to determine predictive power.
+
+We find touble sleeping, meeting exercise requirements, and getting enough sleep were common mental health struggles in the United States.
+Social support was found to vary by COVID-19 test status, and found to be protective againt receiveing a postive COVID-19 test.
+Trouble sleeping and getting enough sleep were found to very slightly increase chances of recieving a positive COVID-19 test.
 
@@
 National Health Information Survey data was obtained from IPUMS, used with permission.
-Data was subsetted to years 2019-2022, to cover the brunt of pandemic years.
+Data was subsetted to years 2020-2022, to cover the brunt of pandemic years.
 Despite low prevelance of COVID-19 cases in 2019 worldwide, the appearance of COVID-19 in the news (and consequential potential impact on mental health factors) led us to include this data.
+Additionally, United States citizens were not being widely tested for COVID-19 before 2020.
 The final dataset contains 151406 observations.
@@
 This may indicate effectiveness of societal control strategies, that is, social distancing, quarantine, and isolation in preventing spread of COVID-19.
-Reporting trouble sleeping was found to slightly increase the odds of a positive COVID-19 test (OR = 1.134 (1.067, 1.206), p < 0.0001), which slightly disagrees with results of the SLEEP vairbale described earlier (Tab.
+Reporting trouble sleeping was found to slightly increase the odds of a positive COVID-19 test (OR = 1.134 (1.067, 1.206), p < 0.0001), which slightly disagrees with results of the SLEEP variable described earlier (Tab.
 3).

### 2025-05-02 products/manuscript/Manuscript.qmd
Part 6 final

@@
 library(gt)
-here()
 ```
@@
 
-* Annalise Cramer$^{1,2 \land}$ (ORCID: 0009-0002-8718-3593)
+* Annalise Cramer$^{1,2}$ (ORCID: 0009-0002-8718-3593)
 
@@
 #| label: fig-schematic
-#| fig-cap: "Figure 1. Flowchart depicting methodology of this study."
+#| fig-cap: "Flowchart depicting methodology of this study."
 #| echo: FALSE
@@
 ```{r}
-#| label: tab-1
-#| fig-cap: "Table 1. Table depicting porportion of responses among all observations, incliding missingness."
+#| label: tbl-1
+#| tbl-cap: "Table depicting porportion of responses among all observations, incliding missingness."
 #| echo: FALSE
@@
 ```{r}
-#| label: tab-2
-#| fig-cap: "Table 2.
-Table depicting prevalence estimates with 95% confidence intervals among the United State population.
-Results are split by COVID-19 test status, and shown for overall population."
+#| label: tbl-2
+#| tbl-cap: "Table depicting prevalence estimates with 95% confidence intervals among the United State population. Results are split by COVID-19 test status, and shown for overall population."
 #| echo: FALSE
@@
 #| label: tbl-summarytable
-#| tbl-cap: "Table 3. Assocation of mental health factors with positive COVID-19 test results among the United State population."
+#| tbl-cap: "Assocation of mental health factors with positive COVID-19 test results among the United State population."
 #| echo: FALSE
-here()
 knitr::include_graphics(here("results","figures", "single_model_table.png"))
@@
 Higher AUC indicates better model preformance.
-Desspite the Logistic Regression model and the Lasso Regression model have the same mean AUC, they are not the same, shown by the differing values for the imputations.
+Despite the Logistic Regression model and the Lasso Regression model have the same mean AUC, they are not the same, shown by the differing values for the imputations.
 Code was checked to ensure the models were not the same.
@@
 #| label: fig-aucplot
-#| fig-cap: "Figure 2.
-Plot comparing area under the curve (AUC) metrics among imputations between logistic regression model, LASSO regression model, and random forest model.
+#| fig-cap: "Plot comparing area under the curve (AUC) metrics among imputations between logistic regression model, LASSO regression model, and random forest model.
 Higher AUC indicates better preformance.
@@
 #| echo: FALSE
-here()
 knitr::include_graphics(here("results","figures", "auc_comparison_plot.png"))
@@
 #| label: fig-log-plot
-#| fig-cap: "Figure 3. Logistic Regression calibration plot."
+#| fig-cap: "Logistic Regression calibration plot."
 #| echo: FALSE
-here()
 knitr::include_graphics(here("results","figures", "log_calibration_plot.png"))
@@
 #| label: fig-lasso-plot
-#| fig-cap: "Figure 4. LASSO Regression calibration plot."
+#| fig-cap: "LASSO Regression calibration plot."
 #| echo: FALSE
-here()
 knitr::include_graphics(here("results","figures", "lasso_calibration_plot.png"))
@@
 #| label: fig-rf-plot
-#| fig-cap: "Figure 5. Random Forest calibration plot. "
+#| fig-cap: "Random Forest calibration plot. "
 #| echo: FALSE
-here()
 knitr::include_graphics(here("results","figures", "rf_calibration_all_imputations.png"))
@@
 2).
-Sleep difficulties were commonly reported during the SARS-CoV-2 pandemic, which supports these results^$5$.
+Sleep difficulties were commonly reported during the SARS-CoV-2 pandemic, which supports these results$^5$.
 
